### 2017-05-18 paperDistance/paper.tex
splitting directories

% This project is part of the stellarTwins project.
% Copyright 2017 the authors.

%@archiver{PUT FIGURE NAME HERE}

% # To-do list
% - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
% - deal with all HOGG or Anderson or LMA in the text.
% - deal with everything in the google doc.
% - are there any github issues?

\documentclass[modern]{aastex61}
\usepackage{amsmath}
\setlength{\parindent}{\baselineskip} % trust me!
\renewcommand{\twocolumngrid}{} % trust me!

% text macros
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\tgas}{\project{\acronym{TGAS}}}
\newcommand{\tmass}{\project{\acronym{2MASS}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\panstarrs}{\project{Pan\acronym{STARRS}}}
\newcommand{\xd}{\acronym{XD}}
\newcommand{\cmd}{\acronym{CMD}}

% math macros
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\given}{\,|\,}
\newcommand{\true}{\mathrm{true}}

% start and title
\shorttitle{improving gaia parallaxes}
\shortauthors{anderson et al.}
\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me
\title{Improving \textsl{Gaia} parallaxes
       with a data-driven model \\
       of the color--magnitude diagram}

\author[0000-0001-5725-9329]{Lauren Anderson}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}

\author[0000-0003-2866-9403]{David W. Hogg}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}

%\author{Jo Bovy}
%\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
%\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}

%\author{Boris Leistedt}
%\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}

%\author{Adrian Price-Whelan}
%\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}

\begin{abstract}\noindent % trust me

Conversion of a noisy parallax measurement into a posterior belief over distances requires inference with a prior.
Usually this prior represents some kind of belief about the Milky Way.
However, there is multi-band photometry for a large fraction of the \gaia\ \tgas\ catalog from imaging surveys;
this imaging is incredibly informative about stellar distances.
Here we use color information of $1.2$ million \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
We use the Extreme Deconvolution (\xd) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
Our posterior photometric parallax estimates are more precise than the \tgas\ catalog entries by a median factor of [XXX]; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
%The precision can be attributed to the statistics concept of shrinkage.
We independently validate our distances by looking at members of a Milky Way star cluster M67, which is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly as a cluster in our posterior parallax estimates.
All our results, including a posterior parallax and distance sampling for [ZZZ] \tgas\ stars, are available in companion electronic tables.
\end{abstract}

\keywords{
  catalogs
  ---
  Hertzsprung--Russell and C--M diagrams
  ---
  methods: statistical
  ---
  parallaxes
  ---
  stars: distances
  ---
  stars: statistics
}

\section{Introduction}

The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar parallax measurements.
Only a small fraction (but large number) of these
measurements will be both precise and purely astrometric.
\gaia\ will use astrometric parallax to determine the distances of the more precise
stars, and then calibrate spectrophotometric models.
These spectrophotometric models, in turn, along with \gaia's on-board
low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
parallax estimates to less precise stars.
The full stack required for these parallax inferences is complex.
It involves modeling the stars, as well as the dust in the Milky Way,
and the response of the telescope itself.
It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.

Projects like \project{The Cannon} (\citealt{ness15}; \citealt{casey17}; \citealt{ho17})
and \project{Avast} (Bedell et al.\ in preparation),
explore the extent to which our predictive models of
stars could be purely data-driven or statistical.
That is, under what circumstances could the data themselves deliver
more precise or more accurate information than any theoretical or
physics-based model?
The answer to this question is extremely context-dependent: It depends
on what data are available, how much data are available, and what
questions are being asked.
In general, data-driven models contain fewer assumptions than
physics-driven models, but they are also usually less interpretable.
However, the \gaia\ data set is ideal for thinking about purely statistical
models of stars:
The distance to a star is a latent variable, and the nearby
stars constitute a high signal-to-noise ``training set''.
These nearby stars can be used to anchor empirical models that can
then be applied to the more distant stars.

Here we deliver a demonstration-of-concept that shows that a
data-driven model of the \gaia\ data could in principle deliver very
precise parallaxes, without any input of the physics or numerical models of stellar
evolution, interiors, or photospheres.
That is, this is a data-driven or purely empirical \emph{photometric
  parallax} method.
A photometric parallax is a parallax (or distance) estimate based on
the photometry (colors and apparent magnitude or magnitudes) of a
star.
Fundamentally, photometric parallaxes capitalize on the fact that the
apparent magnitude of a star is a strong function of distance, and
that the absolute magnitude is a strong function of color (or surface
temperature) and evolutionary stage.

Most stellar models used to generate photometric
parallaxes have been \emph{physical} models, based on gravity, fluid mechanics,
radiative transfer, nuclear reactions, and atomic and molecular transitions.
Because there are small issues with each of these components, the models are inaccurate in detail, and
the physical models therefore produce parallax (and hence distance) estimates that are biased.
In addition, they build a long list of physical assumptions (about
nuclear reactions, convection, and thermal timescales, for example)
into the parallax estimates.
When using physical models, it is impossible to deliver photometric parallax estimates that involve minimal assumptions.

However, the use of physical models for distance estimation is not necessary.
It is possible to build a stellar photometric model from the data themselves,
because there are stars at all luminosities with useful parallax information,.
One challenge is that different stars are observed at different levels of parallax precision,
so it requires relatively sophisticated technology to build this data-driven model
using all the data available, fairly and responsibly.
Luckily, we have this technology!

Here we build and use just such a data-driven model to infer more precise parallaxes for each star.
In particular, we use all of the \gaia\ \tgas\ data (\citealt{tgas};
these data were generated by the methodology of \citealt{michalik15}), that match to the
\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of
the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
We build the \cmd\ using an Empirical-Bayes approach known as Extreme Deconvolution (XD;
\citealt{bovy11}).
This method deconvolves heteroskedastic data to derive an
estimate of the noise-free or low-noise distribution that \emph{would
  have} been observed with far better data.
This method has been used in astrophysics
previously to model the velocity distribution of stars in the Solar
Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
perform quasar target selection
in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
Its advantage over other deconvolution methods is that it takes as input,
and handles in a principled way, heteroskedastic data.
Its disadvantage, for the present purposes, is that it requires
or implies a strictly Gaussian noise model.
As we show below, this requires us to transform the \cmd\ to a space in which
the noise is approximately Gaussian.

We then use the novel \xd\ output---the deconvolved \cmd---as
a prior for use in inference of individual-star parallaxes.
These inferences, one per star, provide much more precise parallax, distance,
or absolute magnitude estimates than we get from each star's primary
\tgas\ parallax measurement alone.
That is, we are using the \xd\ model to de-noise the \gaia\ parallax
measurements.
Technically, since \xd\ produces a maximum-marginalized-likelihood estimator
of the deconvolved distribution,
its use as a prior is technically using the data twice and is therefore only approximately valid.
However, since the data set is large, the \xd\ approximation is not bad; it is
sometimes known as the ``empirical Bayes'' method, and is well studied\footnote{For a start, try \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}.

It wouldn't be using the data twice---that is, it would be valid from a
probabilistic perspective---if we performed a full hierarchical Bayesian
inference.
Exploration of that possibility, and its computational tractability,
is among our long-term goals and motivations.
Along those lines, this paper can be seen as a companion to a related
project (Leistedt et al. in preparation), in which we use a slightly less appropriate
model for the \cmd, but in which we take a much more principled
approach to the inference.

Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
about the Milky Way, and similarly weak assumptions about the properties of
stars.
To date, there has been some discussion in the literature
about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
These papers use the \gaia\ measurement to construct a parallax likelihood,
and combine this with a distance prior.
While this is sensible and correct, all present methods proposed along
these lines build in informative (and known-to-be-wrong) assumptions
about the line-of-sight distance distribution to \gaia\ stars.
That is, they build in strong and informative assumptions about the Milky
Way.
The strongest new assumption made in this work is that for every star
in \tgas, there are other, photometrically (and bolometrically)
similar stars.
That is a much weaker assumption than has been made in most other Bayesian
distance estimations.

This project is fundamentally a demonstration of concept:
We are only using a subset of the ``small'' (relative to the expected final \gaia\ data set)
\tgas\ Catalog.
We are performing only an approximation to full Bayesian hierarchical inference.
We have to use photometry that is ground-based, rather than the full-precision,
space-based photometry that the \gaia\ Mission will deliver.
However, as we show below, we get very good results in terms of parallax precision.
It is promising that we will be able to obtain even better results
in later \gaia\ data releases, and eventually put distances on $>10^9$
\gaia\ stars using photometric parallax methodologies but without any
commitment to---or even use of---physical models of stars or the Milky Way.

\section{Why does this work?}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{toy.png}
\caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior. The upper left panel visualizes the simplified toy distribution. The red lines are the $1\sigma$ contours of the distribution, and the red points are 1024 samples from the distribution. These same $1\sigma$ contours are drawn in the bottom left panel. In the upper right and lower left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them. The error bars show the $1\sigma$ measurement uncertainties. The black points are 5 randomly chosen data to highlight the method. In the lower left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution. The blue points represent the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior. The red points represent the actual true values. The lower right panel shows the posterior belief of the true value for all the points. They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior. In contrast, more certain measurements tend to lie closer to their measured value. The blue points are better than the black for doing stuff.}
\label{fig:toy}
\end{figure}

In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
Understanding this method is challenging because we are using the data both for the prior and the likelihood in our inference.
To ease understanding, here we will demonstrate this method with a simpler toy model.
Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
We will do this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$.
Here the thickness of the toy distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
Similarly, noisy samples from this distribution are shown in the upper right panel.
Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.

Setting up the problem in more detail, we have the true samples from the underlying distribution and the noisy samples

\begin{eqnarray}
y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
y_n \, &=& \, y_{\true,n} + \epsilon_n
\label{eq:ytrue}
\end{eqnarray}

where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
The upper left shows 1024 realizations of the true values, and the upper right, as well as lower left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points. Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.

To infer a pdf of the true y value for each measured y value we need a prior. For our prior we will use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy y data themselves.
We learn the noise-deconvolved function using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data. Here our likelihood for each individual datum is

\begin{equation}
p(y_n \given m, b, t^2, \sigma_n^2) = \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)
\label{eq:toyLike}
\end{equation}

It is a Gaussian because the noise model is a Gaussian. The likelihood of the full data set is then the product of the individual likelihoods

\begin{equation}
\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2).
\label{eq:toyLikeFull}
\end{equation}

Maximizing this joint likelihood of all the data is the same as minimizing

\begin{equation}
\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,ln(2\pi\,[\sigma_n^2 + t^2])
\label{eq:chisq}
\end{equation}

This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$. In Figure \ref{fig:toy}, lower left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines. Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.

Here is a good point to step back and acknowledge that our method is a
bit strange. We are using the set of data to determine the
noise-deconvolved distribution, which is an estimate of the true
distribution, and then using that noise-deconvolved distribution to
make inferences on the data. We are technically using the data
twice! To be fully, technically correct, we should \emph{either} learn the true
distribution while leaving one datum out, the one datum we want to
infer the true y value for, or else go to full hierarchical Bayesian inference,
in which we treat the \cmd\ parameters probabilistically along with everything else.
With large data sets, the MML or Empirical Bayes approximation is safe,
in some sense because no individual data point makes a large difference to
the inference.

Returning the inference:
We use Bayes theorem to turn the likelihood of the noisy data into a posterior believe over the true y value.
\begin{eqnarray}
p(y_{\true,n} \given y_n, \tilde{\sigma}_n) &=& p(y_n \given y_{\true,n}, \sigma_n)\,p(y_{\true,n}) \\
\mathrm{where}\;p(y_n \given y_{\true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{\true,n}, \sigma_n) \\
\mathrm{and}\;p(y_{\true,n}) &=& \mathcal{N}(y_{\true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
\label{eq:toyBayes}
\end{eqnarray}
Since everything is Gaussian, this can be done analytically by multiplying Gaussians together. In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points. The red points are the associated true y values.
In Figure \ref{fig:toy}, lower right panel, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance, $t^2$, and the measurement variance, $\sigma_n^2$.
The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior. In contrast, more certain measurements remain closer to their measured values, $y_n$. We now have better estimates of each true y value by inferring each true y value from the measured y value and using an estimate of the true distribution the measured y value was drawn from as the prior.

In detail, the posterior points shown in the lower-right panel of
Figure~\ref{fig:toy} look more concentrated (vertically) than the true
values shown in the upper-left panel.
This is because the points in the lower-right (posterior) panel show
not typical values or typical draws from the posterior; they show the
modes (and, with error bars, the variances) of the posterior pdfs.
When an original data point has a large variance, the posterior ends
up close to the prior, which has a mode at the center line; the
posterior modes therefore concentrate more towards the center line
than any representative sampling.
A plot of posterior samplings (as opposed to a plot of posterior modes) would
show a scatter like that in the upper-left (truth) panel.

Bringing the toy model back to our original problem:
We can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.

\section{Assumptions and methods}\label{sec:method}

To infer the true parallax to a star using the noise-deconvolved \cmd\ as a prior, we make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the discussion section below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.  In detail, this assumption is wrong, (for example, distant halo stars are different than local
  disk stars), but the stationarity assumption is fairly weak; as long
  as there is support in the high signal-to-noise stars for the types
  of stars seen at low signal-to-noise, the results will not be
  strongly biased. Although a weak assumption, it is deep and fundamental to this
  work.
\item[big data] We assume that we have large numbers of stars, large enough that
  empirical Bayes, or maximum-marginalized likelihood, is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case, if a portion of the \cmd\ has only one (statistically distinct)
  exemplar, for example, a white dwarf, inference for these cases will be biased.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the [INSERT
    CORRECT TERMINOLOGY FROM GAIA HERE] parallax uncertainty estimates
  are correct. Here we are making use of the fact that, for
  de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
  dust correction at its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for neither
  color nor absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128 Gaussians.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

\subsection{Setting up the Inference}
 \emph{Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff. It's confusing right now jumping between inference and prior building !!}

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
The strength of our inference lies in how we generate our prior information of $\varpi_{\true}$.
Supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
We include this distance information in the inference of t$\varpi_{\true}$ by using the \cmd\ as our prior.
Specifically, we use the \cmd\ generated from the data themselves as our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the parallax of that star.

Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars. For inferring $\varpi_{\true}$, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia. Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$. The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\mathbf{y}$. Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude

\begin{equation}
y_0 = \varpi\,10^{0.2J^c_n} = 10^{0.2\,M^c_{J,n} + 2}
\label{eq:transform}
\end{equation}
The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain. Gaussian uncertainties are required to generate our prior using \xd, but more about this and our dust corrections below.
Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map. So our full terminology is

\begin{equation}
\begin{aligned}
\mathrm{true \, vector} \quad &\mathbf{Y}_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
\mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
\mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
\mathrm{assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
\end{aligned}
\label{eq:data}
\end{equation}
where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.

If life were simple, our inference would be a 1D problem using Bayes theorem and look something like:
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the
parallax, and $Z$ is a normalization constant.
But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the data is
\begin{equation}
p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
\end{equation}
with \\
$V_n = $
\[
\begin{bmatrix}
\sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
0 & (\sigma_{\varpi}\,10^{0.2\,J_n^C})^2
\end{bmatrix}
\]
where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude. So we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty. Again, this is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.

This likelihood is the probability of the data vector given the true vector, but we
want to flip it around and infer the true vector and therefore the true parallax, or distance. We do this
analogous to Equation \ref{eq:bayes}.

\begin{equation}
p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n)
\end{equation}
where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.

\subsection{Generating the Prior}

To generate our prior, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude. We fit this transfromed \cmd\ using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties. It generates an estimate of the underlying distribution from which the uncertain data were drawn from. \xd\ generates a represention of the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalizing over the true data.

\xd\ generates the gaussian parameters

\begin{eqnarray}
\alpha &=& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
       &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
\label{eq:xdmml}
\end{eqnarray}
where $\alpha$ encapsulates the parameters of each Gaussian in the mixture of Gaussians that defines our \cmd\ prior.

\subsection{Back to the Inference}

To infer $\varpi_{\true}$, we first do the inference in the 2D space of the \cmd. Using Bayes theorem,

\begin{equation}
p(\mathbf{Y} \given \mathbf{y}, \alpha) = \frac{1}{Z} \, p(\mathbf{y} \given \mathbf{Y}) \, p(\mathbf{Y} \given \alpha)
\label{eq:posterior}
\end{equation}
We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$

\begin{eqnarray}
p(\mathbf{Y}_0 \given \mathbf{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
\mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
              1 \quad \mathrm{if} \quad -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
              0 \quad \mathrm{otherwise}
              \end{cases}
\label{eq:parallaxPost}
\end{eqnarray}
with $f(\varpi_{true})$ being a window function to insure $\varpi_{\true}$ is positive and to put $\varpi_{\true}$ on a similar grid for all the posteriors.

Crossvalidation?

\section{Data and results}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{data.png}
\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K$ color, the $J$ band apparent magnitude, and the \gaia\ parallax. The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution. To give a sense of the uncertainties, in the right panel we subsample the dataset and include the $1\sigma$ uncertainties. The uncertainties are dominated by the parallax noise, with many stars diverging to infinitely far away and therefore very intrinsically bright.}
\label{fig:data}
\end{figure}

We use stars crossmatched in \tgas\ and \tmass\, that also lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model. We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.

\subsection{Dust}

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{dust.png}
\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates. This shows both the footprint of our analysis as well as the dust values we used. Each point represents a single star in our cross matched catalog. The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
\label{fig:dust}
\end{figure}

We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust. With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way. The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference. Dust corrections are particularly non trivial for stars with poor parallax signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion). Instead, we correct for dust using a point estimate from the 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$. In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior, we infer more precise parallaxes to all the stars in our sample. We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star. And repeat. This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd and bias the parallax inference. For each distance, we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD). We then regenerate our prior using the lightly dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution. We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?

\subsection{Output Data}
Using a data driven model of the observed \cmd\ we inferred a posterior belief of the true parallax of each star. Given our method, we call our posterior parallax the photometric parallax.
HOGG: Placeholder for the output format for our data products.
RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{prior_ngauss128.png}
\caption{Prior: the XD generated mixture of gaussians we use as our prior. The left panel shows a sampling of the prior, which is a mixture of gaussians shown in the right hand panel. The blue lines are the $1\sigma$ and $2\sigma$ contours. The gaussians have a funny shape due to the transformed space in which we fit the data.}
\label{fig:prior}
\end{figure}

%\begin{figure}
%\centering
%  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
%\caption{Prior: the XD generated mixture of gaussians we use as our prior. To determine the number of gaussians we used cross validation, using 70\% of our %sample as a training set and 30\% as the test set. Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, %showing that we are no over fitting. Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the %features of the CMD.}
%\label{fig:priorCompare}
%\end{figure}

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior.png}
\caption{5 example posteriors showing the priors and likelihoods as well}
\label{fig:posterior}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posteriorCMD.png}
\caption{the expectation values and subsampled error bars. Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd. The visual of the prior is a sampling of the Gaussians, but this is the expectation value so this lies much closer to the center of the distribution. Similar to the toy model, where the pdf maximum of the noisier points lies in the center of the distribution.}
\label{fig:posteriorCMD}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{delta.png}
\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color. The coloring of the points in both panels is by the fractional change in variance to guide the coloring of the right panel. The blue lines are the $1\sigma$ and $2\sigma$ contours. The median fractional change is XXX. Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^C$ about XXX. The right panel shows the posterior point estimates on the CMD colored by fraction change in variance. There is significant changes for the red giant branch stars, as well as some main sequence stars.}
\label{fig:delta}
\end{figure}

\subsection{Distances to M67}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{distancesM67.png}
\caption{Proof of concept: parallax inference to the open cluster M67}
\label{fig:m67}
\end{figure}

As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67. It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away. Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5. With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies. So it is a great test case for comparing parallax inferences with or without a prior. We made a window of 1 square degree on the sky with the cluster M67 lying in the middle. The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky. With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67. Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors. The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster. The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all. The posterior pdf, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods. Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.

\section{Discussion}

We have shown that it is possible to obtain photometric parallaxes for
distant stars in the \gaia\ \tgas\ data without any use of physical stellar
models, nor models of the Milky Way.
We used the geometric parallaxes to calibrate a photometric
model that is purely statistical, which is a model of the data rather than
a model of stars \foreign{per se}.
This opens up the possibility of completing the goals of the \gaia\ Mission
without building in unnecessary assumptions about the mechanical properties
of stars or the Galaxy.

Transition clause/sentence. We obtained the photometric parallaxes in this project
by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
and using it as a prior pdf for Bayesian inference.
The posterior pdfs for distance that we obtain are, in general, much
narrower than the likelihood functions delivered by the
\gaia\ Mission, and therefore the distance estimates (or,
equivalently, parallax estimates) are much more precise.
It is not surprising that a Bayesian inference provides more
precise inferences than the likelihood function alone; Bayesian
inferences bring in new information that decrease variance (but
often introduce bias).
We have shown that, in addition to pure precision improvements, at
least some aspects of \emph{accuracy} have been improved as well, by
showing that the posterior distance estimates to stars in (at least one) stellar
cluster are much more clustered than likelihood-based distance
estimates.

Transition clause/sentence. Although we have not performed principled hierarchical Bayesian inference in this
project, we built our prior for each star's parallax by
performing a statistically responsible deconvolution
of the \gaia\ \tgas\ data that is justifiable under a clear set of
assumptions.
This deconvolution, using a reasonable noise model and an
assumption of stationarity, shrinks the parallax uncertainties, most dramatically for the stars measured at lower signal-to-noise.
Because the method we use, \xd, which is
an Empirical Bayesian maximum-marginalized-likelihood estimato, accounts for
heteroskedastic noise, we were able to build our photometric parallax
model---which is a model of the \cmd---using
all the data, not just the data with the highest signal-to-noise.
So, the model for the \cmd\ we built is representative for
our subset of the \tgas\ Catalog.
Because our model is representitive of our full subset catalog, we expect any (sensible) parallax estimates
we generate from our parallax posterior pdfs to be only weakly biased.

Transition clause/sentence. The empirical Bayesian methodology is a good approximation to full
hierarchical inference in the limit of large numbers of data points, such that no individual star is carrying a lot of weight in the
model of the \cmd.
With its millions of stars, the \tgas\ Catalog appears to safely
live in this limit.
However, because there are parts of the \cmd\ that are poorly populated,
even in \tgas, the empirical Bayes approximation may be bad for some
portions of the \cmd.
In particular, there appear to be essentially no white dwarf stars,
and other classes (like bright red giant branch stars) have few to no exemplars
in the data set at high signal-to-noise (in parallax).
For all these reasons, we do think that this approximation is not perfectly
safe, and it is a goal to explore full hierarchical modeling in the coming
years (and see, for example, Leistedt et al. in preparation).

Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
function or completeness estimate or inverse-selection-volume corrections.
This model is a model of the \tgas\ Catalog
(or really the \tgas--\tmass\ intersection), not of the stars in
the Milky Way, nor any volume-limited subsample of the Milky Way.
Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
intersection, the model \emph{is} representive of the full intersection,
even the distant parts, where all the parallaxes are measured at low
signal-to-noise.
This is in contrast to techniques that might build the model from only
the high signal-to-noise sources; such a model would be biased towards
stellar types and compositions found locally, as well as lower-luminosity stars,
and stars which happened to get good parallaxes.
However, our use of the entire \tgas--\tmass\ intersection without
any selection function restricts the use of our \cmd\ model as a prior
to inferences within that same \tgas--\tmass\ intersection.
It would be biasing to use this same prior to infer
distances for the full billion-star catalog released in \gaia\ DR1
alongside \tgas, or for stars fainter or brighter or bluer than those
included in the \tgas--\tmass\ intersection.

Transition clause/statement. The outputs of this project include posterior pdfs for stellar parallaxes.
It is tempting to treat these outputs as equivalent to some kind of catalog of
measurements, as we treated the original parallax measurements from \tgas.
That is, it is tempting to treat these as simply ``better'' measurements.
Taken one star at a time, this is permitted and true, and the basis for our
claim that we have de-noised the \tgas\ Catalog.
However, there are important differences, as there are with all
probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
measurements and the de-noised measurements.
For users of our output, the most important difference between our
output and the \tgas\ data input is that the \tgas\ data is the representation
of a likelihood function, whereas our ouptut is a
representation of a posterior pdf, one per star.
Posterior information used naively can lead to serious statistical
errors, because each data point has had a prior multiplied in.
Multiplicative uses of the data will effectively take the prior
pdf to a large power.
We have (in previous work)
given examples of correct uses of posterior samplings or pdfs
for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
encourage power-users to consult those methodological
contributions before engaging with these outputs.

\subsection{Comparison with other priors}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{deltaSimple.png}
\caption{Other example prior: The same visualization as Figure \ref{fig:delta} of the change in the posterior, but here, if instead of prior information containing the \cmd\, the prior information is an exponentially declining spatial density of stars.  }
\label{fig:otherPrior}
\end{figure}

%How this compares to other Bayesian attempts at inferring distances,
%like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
%assumptions, in fact! What form should this comparison be in?
There are other options for prior information when inferring distances to stars. One option is the exponentially declining spatial density of stars (cite: Tri, CBJ). It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth. It is a fairly weak prior. It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax. It is a fairly broad prior so the median \gaia\ star has an increased posterior variance of XXX. It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes. Given this prior, what does the observed \cmd\ look like?  The \cmd\ is broad and the red clump highly stretched.

\subsection{What's That Feature?}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{whatsThatFeature.pdf}
\caption{Features in the \cmd: The expectation values from the posterior pdfs of the true parallax}
\label{fig:wtf}
\end{figure}

The denoised expectation values of the parallax projected into the \cmd\ show interest features.
\begin{itemize}
\item (1) Narrow Lower Main Sequence
\item (2) Wide Upper Main Sequence
\item (3) Plausible Binary Sequence
\item (4) Prominent Red clump
\item (5) Narrow Red Giant Branch
\item (6) Noticeable turn off
\end{itemize}
The lower main sequence is very narrow while the upper main sequence is much wider. This may be an age variation on the main sequence mixed with a malquist bias of seeing a much larger volume of upper main sequence stars. There is a plausible binary sequence of stars with about twice the brightness as the main sequence. The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is fairly straight and narrow. The main sequence turn off is noticeable and also surprisingly narrow, possibly a reflection of the star formation history of the local volume also mixed with a malmquist bias.

\subsection{Critical discussion of assumptions}

In Section~\ref{sec:method} we listed a set of assumptions upon which
the photometric parallax method in this project is based.
Here we return to these assumptions, and discuss them critically in more
detail.
\begin{description}
\item[stationarity] (HOGG needs to clean this paragraph)
  We have assumed that (locally) the morphology of
  the \cmd\ is independent of signal-to-noise; that is, we have assumed that
  both the low and high
  signal-to-noise objects are being drawn from the same distribution.
  This isn't precisely true because nearby stars in \tgas\ are
  mostly main sequence stars, while we can see luminous giants to much larger
  distances. So the red giant branch is a population of
  relatively lower signal to noise objects compared with the main
  sequence. This does not violate the stationarity assumption in
  detail, because the low and high signal to noise objects have to be
  drawn from the same distribution within some feature or patch of the
  space. This is all assuming the Gaia noise model is correct. If the
  uncertainties are over estimated then we could be over deconvolving
  (generating a tighter distribution than the truth) And in some
  extreme over estimate of uncertainties the lower signal to noise
  giants could be pulled to the main sequence. But with the quoted
  uncertainties this seems unlikely. As you move up the giant branch,
  and probe farther giants, the signal to noise will continue to
  diminish, but because the color uncertainties are not large, the S/N
  spread within a patch of the \cmd\ is well mixed. Possibly a
  metallicity trend with distance, placing more nearby (higher S/N)
  objects at slightly different positions on the CMD pulling lower S/N
  objects to that distribution for our prior. But if something has
  small error bars it is good data is good data both for inference and
  generating the prior.

  For very rare kinds of stars. Inference with our prior is bad for
  underrepresented populations in the \cmd. For example, it appears
  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
  underrepresented in the prior, the amplitudes of the gaussians
  representing them is very small, so the WDs are more likely to be
  pulled to mimic a much farther away MS star.
\item[big data] HOGG 
\item[noise model] HOGG
\item[dust] Within our probabilistic framework, we improperly handle
  dust. We are taking a point estimate of the probabilistic dust model
  at a quantile of our distance posterior, correcting our photometry
  using this point estimate, and then re-deriving the prior. A full,
  proper account would sample the dust model at distances sampled from
  the posterior, and propagate those samples back into the prior
  during the iterative process. The reason we don't do a full
  probabilistic treatment, is that this proper dust distribution would
  not be a Gaussian distribution, which \xd\ requires. We could
  calculate this more complex dust distribution and estimate it as a
  Gaussian distribution to feed to \xd, however, this is beyond the
  scope of this paper.
\item[mixture of Gaussians] HOGG
\item[no physics] HOGG
\end{description}

\subsection{Cross Validation}
For generating our prior, we use a mixture of 128 Gaussians. This number, 128, was chosen heuristically. It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product. We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.

Anderson: Compare with literature stellar models?? This is an enhanced goal; strike this.

\acknowledgments It is a pleasure to thank
  Ana Bonaca (Harvard),
  Andy Casey (Monash),
  Dustin Lang (Toronto) for cross matched 2MASS-TGAS catalogue
  Jo Bovy (Toronto)
  Boris Leistedt (NYU)
  Adrian Price-Whelan (Princeton)
and the attendees at the Stars Group Meeting at the Flatiron Institute
Center for Computational Astrophysics for comments and input.
This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
by the Center for Computational Astrophysics at the Simons Foundation
in New York City.

This work has made use of data from the European Space Agency (\acronym{ESA})
mission \gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the Gaia
Data Processing and Analysis Consortium\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}} (\acronym{DPAC}). Funding for the
\acronym{DPAC} has been provided by national institutions, in particular the
institutions participating in the Gaia Multilateral Agreement.

This publication makes use of data products from the Two Micron All
Sky Survey, which is a joint project of the
University of Massachusetts and the Infrared Processing and Analysis
Center/California Institute of Technology, funded by the National
Aeronautics and Space Administration and the National Science
Foundation.

This project was partially supported by This research was partially supported by
  the \acronym{NSF} (\acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  and the Moore-Sloan Data Science Environment at \acronym{NYU}.
It made use of the \acronym{NASA} Astrophysics Data System.
All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license.

%% \appendix
%% \section{Appendix material}

\bibliography{gaia}

\clearpage

\end{document}

%%
%% End of file `sample.tex'.

### 2017-05-18 comparePrior.py

@@ -1,0 +1,521 @@
+import matplotlib
+matplotlib.use('pdf')
+import numpy as np
+from xdgmm import XDGMM
+import drawEllipse
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import testXD
+import sys
+import demo_plots as dp
+import os
+import stellarTwins as st
+import scipy.integrate
+import corner
+from astroML.plotting import setup_text_plots
+from astropy.io import ascii
+
+def prior(xdgmm, ax):
+    for gg in range(xdgmm.n_components):
+        points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
+        ax[0].plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=1, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
+
+
+def dustFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+
+def priorSample(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz', Nsamples=1.2e6, xdgmmFilename='xdgmm.fit', xlabel='X', ylabel='Y', contourColor='k'):
+
+    setup_text_plots(fontsize=16, usetex=True)
+
+    xdgmm = XDGMM(filename=xdgmmFilename)
+    figPrior = plt.figure(figsize=(12, 5.5))
+    figPrior.subplots_adjust(left=0.1, right=0.95,
+                            bottom=0.15, top=0.95,
+                            wspace=0.1, hspace=0.1)
+    sample = xdgmm.sample(Nsamples)
+    negParallax = sample[:,1] < 0
+    nNegP = np.sum(negParallax)
+    while nNegP > 0:
+        sampleNew = xdgmm.sample(nNegP)
+        sample[negParallax] = sampleNew
+        negParallax = sample[:,1] < 0
+        nNegP = np.sum(negParallax)
+
+    samplex = sample[:,0]
+    sampley = testXD.absMagKinda2absMag(sample[:,1])
+    ax3 = figPrior.add_subplot(121)
+    alpha = 0.1
+    xlim = [-0.25, 1.25]
+    ylim = [6, -6]
+
[…]

### 2017-05-18 demo_plots.py

@@ -1,0 +1,280 @@
+import numpy as np
+import matplotlib
+matplotlib.use('pdf')
+#from matplotlib.backends.backend_pdf import PdfPages
+from matplotlib import pyplot as plt
+from astroML.plotting.tools import draw_ellipse
+from astroML.plotting import setup_text_plots
+from sklearn.mixture import GMM as skl_GMM
+import drawEllipse
+import matplotlib as mpl
+plt.style.use('seaborn-talk')
+mpl.rcParams['xtick.labelsize'] = 18
+mpl.rcParams['ytick.labelsize'] = 18
+from scipy.stats import gaussian_kde
+import corner
+
+def plot_bic(param_range,bics,lowest_comp):
+    plt.clf()
+    setup_text_plots(fontsize=16, usetex=True)
+    fig = plt.figure(figsize=(12, 6))
+    plt.bar(param_range-0.25,bics,color='blue',width=0.5)
+    plt.text(lowest_comp, bics.min() * 0.97 + .03 * bics.max(), '*',
+             fontsize=14, ha='center')
+
+    plt.xticks(param_range)
+    plt.ylim(bics.min() - 0.01 * (bics.max() - bics.min()),
+             bics.max() + 0.01 * (bics.max() - bics.min()))
+    plt.xlim(param_range.min() - 1, param_range.max() + 1)
+
+    plt.xticks(param_range,fontsize=14)
+    plt.yticks(fontsize=14)
+
+
+    plt.xlabel('Number of components',fontsize=18)
+    plt.ylabel('BIC score',fontsize=18)
+
+    plt.show()
+
+def plot_val_curve(param_range, train_mean, train_std, test_mean,
+                   test_std, log=False):
+    plt.clf()
+    setup_text_plots(fontsize=16, usetex=True)
+    fig=plt.figure(figsize=(12,8))
+    plt.plot(param_range, train_mean, label="Training score",
+             color="red")
+    plt.fill_between(param_range, train_mean - train_std,
+                     train_mean + train_std, alpha=0.2, color="red")
+    plt.plot(param_range, test_mean,label="Cross-validation score",
+             color="green")
+    plt.fill_between(param_range, test_mean - test_std,
[…]

### 2017-05-18 drawEllipse.py

@@ -1,0 +1,34 @@
+import numpy as	np
+
+def fixAbsMag(x):
+    return 5.*np.log10(10.*x)
+
+def plotvector(mean, var, step=0.001):
+    """
+    mean, var should be *projected* to the 2-d space in which plotting is about to occur
+    """
+    assert mean.shape == (2,)
+    assert var.shape ==	(2, 2)
+    ts = np.arange(0, 2. * np.pi, step) #magic
+    w, v = np.linalg.eigh(var)
+    ps = np.sqrt(w[0]) * (v[:, 0])[:,None] * (np.cos(ts))[None, :] + \
+         np.sqrt(w[1]) * (v[:, 1])[:,None] * (np.sin(ts))[None, :] + \
+	     mean[:, None]
+    return ps
+
+if __name__ == "__main__":
+    from xdgmm import XDGMM
+    import pylab as plt
+
+    xdgmm = XDGMM(filename='xdgmm.1028gauss.1.2M.fit')
+    amps = xdgmm.weights
+    mus = xdgmm.mu
+    Vs = xdgmm.V
+
+    plt.clf()
+    for	amp, mean, var in zip(amps, mus, Vs):
+        ps = plotvector(mean, var)
+        plt.plot(ps[0,:], fixAbsMag(ps[1,:]), "k-", alpha=amp/np.max(amps))
+    plt.xlim(-2, 3)
+    plt.ylim(10, -6)
+    plt.savefig("drawEllipse.png")

### 2017-05-18 dustDiff.py

@@ -1,0 +1,166 @@
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+from dustmaps.bayestar import BayestarQuery
+from astropy.coordinates import SkyCoord
+import astropy.units as units
+import testXD
+
+def distanceFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'distanceQuantiles.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+def dustFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+
+ngauss = 128
+thresholdSN = 0.001
+survey = '2MASS'
+dataFilename = 'All.npz'
+survey = '2MASS'
+quantile = 0.5
+
+tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+
+
+dustEBVnew = None
+if quantile == 0.5:
+    dustFile = dustFilename(ngauss, 0.05, '5th', survey, dataFilename)
+    dust = np.load(dustFile)
+    dustEBV = dust['ebv']
+else:
+    dustFile = dustFilename(ngauss, 0.05, '1st', survey, dataFilename)
+    dust = np.load(dustFile)
+    dustEBV = np.zeros(len(dust['ebv']))
+
+deltaDustMax = 2
+
+mainFig, mainAx = plt.subplots(1, 2, figsize=(15, 6))
+dustDiffFig, dustDiffAx = plt.subplots(1, 2, figsize=(15, 6))
+dustDistParFig = plt.figure()
+dustDiffMeanFig, dustDiffMeanAx = plt.subplots(2,2, figsize=(15,6))
+dustDiffMeanAx = dustDiffMeanAx.flatten()
+
+iterNum = [1, 2, 3, 4, 5, 6, 7, 8]#, 9]
+iteration = ['1st', '2nd', '3rd', '4th', '5th', '6th', '7th', '8th']#, '9th']#, '10th']
[…]

### 2017-05-18 testXD.py

@@ -1,0 +1,1102 @@
+import matplotlib
+matplotlib.use('pdf')
+import matplotlib.pyplot as plt
+import os
+import sys
+import stellarTwins as st
+from extreme_deconvolution import extreme_deconvolution as ed
+import astropy.units as units
+import numpy as np
+import pdb
+import time
+from sklearn.model_selection import train_test_split
+from astropy.io import fits
+from xdgmm import XDGMM
+from sklearn.learning_curve import validation_curve
+from sklearn.cross_validation import ShuffleSplit
+import demo_plots as dp
+import drawEllipse
+from astropy.coordinates import SkyCoord
+import astropy.units as units
+import scipy.integrate
+import time
+from matplotlib.colors import LogNorm
+from dustmaps.bayestar import BayestarQuery
+
+def convert2gal(ra, dec):
+    """
+    convert ra and dec to l and b, sky coordinates to galactic coordinates
+    """
+    return SkyCoord([ra, dec], unit=(units.hourangle, units.deg))
+
+def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963'):
+    """
+    return the indices of tgas which fall within dl and db of l and b.
+    default is M67
+    """
+
+    index = (tgas['b'] < np.float(b) + db) & \
+            (tgas['b'] > np.float(b) - db) & \
+            (tgas['l'] < np.float(l) + dl) & \
+            (tgas['l'] > np.float(l) - dl)
+    if plot:
+        fig, ax = plt.subplots()
+        ax.scatter(tgas['l'][index], tgas['b'][index], alpha=0.5, lw=0)
+        fig.savefig('clusterOnSky.png')
+    return index
+
+def absMagKinda2Parallax(absMagKinda, apparentMag):
+    """
+    convert my funny units of parallax[mas]*10**(0.2*apparent magnitude[mag]) to parallax[mas]
+    """
+    return absMagKinda/10.**(0.2*apparentMag)
+
+
+def parallax2absMagKinda(parallaxMAS, apparentMag):
+    """
+    convert parallax to my funny units of parallax[mas]*10**(0.2*apparent magnitude[mag])
+    """
+    return parallaxMAS*10.**(0.2*apparentMag)
+
+
[…]

### 2017-05-18 tightPosterior.py

@@ -1,0 +1,146 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import comparePrior
+import testXD
+from xdgmm import XDGMM
+import testXD
+import stellarTwins as st
+np.random.seed(42)
+
+def distanceFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'distanceQuantiles.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+def dustFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+
+ngauss = 128
+survey = '2MASS'
+quantile = 0.05
+dataFilename = 'All.npz'
+norm = mpl.colors.Normalize(vmin=-1, vmax=5)
+iter = '8th'
+
+if survey == 'APASS':
+    mag1 = 'B'
+    mag2 = 'V'
+    absmag = 'G'
+    xlabel='B-V'
+    ylabel = r'M$_\mathrm{G}$'
+    xlim = [-0.2, 2]
+    ylim = [9, -2]
+
+if survey == '2MASS':
+    mag1 = 'J'
+    mag2 = 'K'
+    absmag = 'J'
+    xlabel = 'J-K$_s$'
+    ylabel = r'M$_\mathrm{J}$'
+    xlim = [-0.25, 1.25]
+    ylim = [6, -6]
+
+
+
+tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+
+dustFile = dustFilename(ngauss, quantile, iter, survey, dataFilename)
+data = np.load(dustFile)
+dustEBV = data['ebv']
+
+color = testXD.colorArray(mag1, mag2, dustEBV, bandDictionary)
+absMagKinda, apparentMagnitude = testXD.absMagKindaArray(absmag, dustEBV, bandDictionary, tgas['parallax'])
+
+color_err = np.sqrt(bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']]**2. + bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']]**2.)
[…]

### 2017-05-18 toy/toy.py

@@ -1,0 +1,212 @@
+import numpy as np
+import scipy.optimize as op
+import sys
+
+
+def make_fake_data(parsTrue, N=512):
+    np.random.seed(42)
+    mtrue, btrue, ttrue = parsTrue
+    xns = np.random.uniform(size=N) * 2. - 1.
+    tns = np.random.normal(scale=ttrue, size=N)
+    #tns = np.random.normal(size=N) * ttrue
+    yntrues = mtrue * xns + btrue + tns
+    sigmans = (np.random.uniform(size=N)*1.5) ** 3.
+    #yns = yntrues + np.random.normal(size=N) * sigmans
+    yns = yntrues + np.random.normal(scale=sigmans)
+    fig, ax = plt.subplots(1,3)
+    ax[0].scatter(xns, tns, s=2)
+    ax[0].set_ylabel('sampled noise from ttrue gaussian')
+    ax[1].scatter(xns, sigmans, s=2)
+    ax[1].set_ylabel('sigmans')
+    ax[2].scatter(xns, np.random.normal(scale=sigmans), s=2)
+    ax[2].set_ylabel('sampled noise from sigmans gaussian')
+    plt.tight_layout()
+    fig.savefig('laurenUnderstanding.png')
+    return xns, yns, sigmans, yntrues
+
+def objective(pars, xns, yns, sigmas):
+    m, b, t = pars
+    resids = yns - (m * xns + b)
+    return np.sum(resids * resids / (sigmas * sigmas + t * t)
+                  + np.log(sigmas * sigmas + t * t))
+
+def get_best_fit_pars(xns, yns, sigmans):
+    pars0 = np.array([-1.5, 0., 1.])
+    result = op.minimize(objective, pars0, args=(xns, yns, sigmans))
+    return result.x
+
+def denoise_one_datum(xn, yn, sigman, m, b, t):
+    s2inv = 1. / (sigman * sigman)
+    t2inv = 1. / (t * t)
+    return (yn * s2inv + (m * xn + b) * t2inv) / (s2inv + t2inv), \
+        np.sqrt(1. / (s2inv + t2inv))
+
+def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5):
+
+    alpha_all = 0.05
[…]

### 2017-05-18 comparePrior.py
tweaks

@@ -14,6 +14,7 @@ import scipy.integrate
 import corner
 from astroML.plotting import setup_text_plots
 from astropy.io import ascii
+import corner
 
 def prior(xdgmm, ax):
     for gg in range(xdgmm.n_components):
@@ -128,7 +129,7 @@ def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='
     cb.set_label(r'E($B-V$ )')
     fig.savefig('dustViz.dQ' + str(quantile) + '.png')
 
-def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz', sdss5=False):
+def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz', sdss5=False, whatsThatFeature=False):
 
     if survey == 'APASS':
         mag1 = 'B'
@@ -191,6 +192,7 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
     testax[2].set_ylabel('absMag Plus')
     plt.show()
     """
+
     dp.plot_sample(color[positive], testXD.absMagKinda2absMag(y), sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                 xdgmm, xerr=color_err[positive], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim, ylim=ylim, errSubsample=2.4e3, thresholdScatter=2., binsScatter=200, contourColor=contourColor)
     dataFile = 'data_noDust.pdf'
@@ -205,7 +207,7 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
 
     cNorm  = plt.matplotlib.colors.Normalize(vmin=-6, vmax=2)
     posteriorFile = 'posteriorParallax.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
-    for file in [posteriorFile, 'posteriorSimple.npz']:
+    for file in [posteriorFile]:#, 'posteriorSimple.npz']:
         data = np.load(file)
         parallax = data['mean']
         parallax_err = np.sqrt(data['var'])
[…]

### 2017-05-18 demo_plots.py

@@ -105,7 +105,7 @@ def kdeDensity(ax, x, y, threshold=0.01, bins=100, s=1, lw=0, alpha=1):
     return ax
 
 
-def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None, plot_contours=True, sdss5=False):
+def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None, plot_contours=True, sdss5=False, whatsThatFeature=False, rasterized=False):
 
     prng = np.random.RandomState(1234567890)
     setup_text_plots(fontsize=16, usetex=True)
@@ -203,7 +203,6 @@ def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None
         cb.set_label(r'$ln \, \tilde{\sigma}_{\varpi}^2 - ln \, \sigma_{\varpi}^2$', fontsize=20)
         cb.set_clim(-7, 2)
     """
-
     figData.savefig('plot_sample.data.pdf', format='pdf')
     figPrior.savefig('plot_sample.prior.pdf', format='pdf')
 

### 2017-05-18 testXD.py

@@ -229,7 +229,7 @@ def posterior2d(means, amps, covs, xbins, ybins, nperGauss=100000., plot=False):
     Z, xedges, yedges = np.histogram2d(samplesX, absMagKinda2absMag(samplesY), bins=[xbins, ybins], normed=True)
     return xedges, yedges, Z, samplesY
 
-def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike, covLike, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=None, Y=None, Z=None, plot2D=False, samplesY=None, axAll=None):
+def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike, covLike, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=None, Y=None, Z=None, plot2D=False, samplesY=None, axAll=None, dataColor='black', posteriorColor='blue', truthColor='red', likeLabel=None, postLabel=None):
     """
     plot the posterior of each star into example.png and feed it axAll to put it on another plot
     """
@@ -251,8 +251,11 @@ def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike,
         else:
             #plot each individual gaussian of the posterior weighted by its amplitdue
             points = drawEllipse.plotvector(meanPost[gg], covPost[gg])
-            ax[0].plot(points[0, :], absMagKinda2absMag(points[1,:]), 'k', alpha=ampPost[gg]/np.max(ampPost), lw=0.5)
-            if axAll is not None: axAll[3].plot(points[0, :], absMagKinda2absMag(points[1,:]), 'k', alpha=ampPost[gg]/np.max(ampPost), lw=0.5)
+            ax[0].plot(points[0, :], absMagKinda2absMag(points[1,:]), posteriorColor, alpha=ampPost[gg]/np.max(ampPost), lw=0.5)
+            if axAll is not None:
+                if ampPost[gg] == np.max(ampPost): label=postLabel
+                else: label=None
[…]

### 2017-05-18 tightPosterior.py

@@ -42,6 +42,11 @@ if survey == '2MASS':
     ylim = [6, -6]
 
 
+trueColor = 'darkred'
+priorColor = 'darkgreen'
+posteriorColor='royalblue'
+#posteriorMap = mpl.cm.get_cmap('Blues')
+dataColor = 'black'
 
 tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
 
@@ -70,12 +75,17 @@ xparallaxMAS = np.linspace(0, 10, nPosteriorPoints)
 y = absMagKinda
 yplus  = y + absMagKinda_err
 yminus = y - absMagKinda_err
-parallaxErrGoesNegative = yminus < 0
+#parallaxErrGoesNegative = yminus < 0
 absMagYMinus = testXD.absMagKinda2absMag(yminus)
-absMagYMinus[parallaxErrGoesNegative] = -50.
-yerr_minus = testXD.absMagKinda2absMag(y) - absMagYMinus
-yerr_plus = testXD.absMagKinda2absMag(yplus) - testXD.absMagKinda2absMag(y)
+absMagYPlus = testXD.absMagKinda2absMag(yplus)
+absMag = testXD.absMagKinda2absMag(y)
+yerr_minus = absMag - absMagYMinus
+yerr_plus = absMagYPlus - absMag
 
+yerr_minus[y < 0] = 100.
+yerr_plus[y < 0] = 100.
+yerr_minus[np.isnan(yerr_minus)] = 100.
+import pdb; pdb.set_trace()
 
 #plot likelihood and posterior in each axes
 for iteration in np.arange(20, 40):
@@ -88,7 +98,7 @@ for iteration in np.arange(20, 40):
     #plot prior in upper left
     xdgmmFilename = 'xdgmm.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.2MASS.All.npz.fit'
     xdgmm = XDGMM(filename=xdgmmFilename)
-    testXD.plotPrior(xdgmm, ax[0], c='k', lw=1)
+    testXD.plotPrior(xdgmm, ax[0], c=priorColor, lw=1)
     ax[0].set_xlim(xlim)
     ax[0].set_ylim(ylim)
     ax[0].set_xlabel('$(J-K)^C$', fontsize=18)
@@ -99,8 +109,8 @@ for iteration in np.arange(20, 40):
         index = currentInd[np.random.randint(0, high=len(currentInd))]
 
[…]

### 2017-05-18 toy/toy.py

@@ -1,7 +1,7 @@
 import numpy as np
 import scipy.optimize as op
 import sys
-
+import matplotlib as mpl
 
 def make_fake_data(parsTrue, N=512):
     np.random.seed(42)
@@ -41,42 +41,38 @@ def denoise_one_datum(xn, yn, sigman, m, b, t):
     return (yn * s2inv + (m * xn + b) * t2inv) / (s2inv + t2inv), \
         np.sqrt(1. / (s2inv + t2inv))
 
-def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5):
+def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5, dataColor='k', priorColor='g', posteriorColor='b', trueColor='r', posteriorMap = 'Blues'):
 
     alpha_all = 0.05
     alpha_chosen = 1.0
     xlim = (-1, 1)
     ylim = (-5, 5)
 
-    dataMap = mpl.cm.get_cmap('Blues')
-    dataColor = dataMap(0.75)
-    trueMap = mpl.cm.get_cmap('Reds')
-    trueColor = trueMap(0.75)
-
     for ax in axes[1:-1]:
-        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color="k", alpha=alpha_all ,mew=0)
-        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color="k", zorder=37, alpha=alpha_chosen, mew=0)
+        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color=dataColor, alpha=alpha_all ,mew=0)
+        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
 
     xp = np.array(xlim)
-    axes[0].plot(xp, mtrue*xp + btrue + ttrue, color='red', linewidth=2, alpha=0.75, label=r'$y=m_{true}\,x+b_{true}\pm t$')
-    axes[0].plot(xp, mtrue*xp + btrue - ttrue, color='red', linewidth=2, alpha=0.75)
[…]

### 2017-05-23 paperDistance/paper.tex
various edits till sec 3.2

@@
 \usepackage{amsmath}
+\usepackage{color}
 \setlength{\parindent}{\baselineskip} % trust me!
 \renewcommand{\twocolumngrid}{} % trust me!
+
+\newcommand\BLedit[1]{{\color{blue}{#1}}}
+\newcommand\BLcomment[1]{{\color{cyan}{[BL: #1]}}}
 
@@
 These nearby stars can be used to anchor empirical models that can
-then be applied to the more distant stars.
+then be applied to the more distant stars\BLedit{, thanks to the simple physical relations connecting the measurements (apparent magnitudes, parallaxes) to some of the physical parameters of interest (absolute magnitudes, distances).}
 
@@
 Along those lines, this paper can be seen as a companion to a related
-project (Leistedt et al. in preparation), in which we use a slightly less appropriate
+project \citep{leistedtHogg2017}, in which we use a slightly less appropriate
 model for the \cmd, but in which we take a much more principled
@@
 
-\begin{equation}
-p(y_n \given m, b, t^2, \sigma_n^2) = \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)
+\BLedit{
+\begin{eqnarray}
+p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \mathrm{d}y_{\true, n} \\
+&=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)
 \label{eq:toyLike}
-\end{equation}
+\end{eqnarray}
 
-It is a Gaussian because the noise model is a Gaussian. The likelihood of the full data set is then the product of the individual likelihoods
+where we have marginalized over the true sample $y_{\true, n}$.} The resulting likelihood is a Gaussian because the noise model is a Gaussian.
+\BLcomment{should all these equations have $x_n$ on the right hand side?
+I don't have a strong opinion about that.} The likelihood of the full data set is then the product of the individual likelihoods
 
@@
 \begin{equation}
-\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,ln(2\pi\,[\sigma_n^2 + t^2])
+\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,\ln(2\pi\,[\sigma_n^2 + t^2])
 \label{eq:chisq}
@@
   work.
+\BLcomment{Isn't that assumption made conditionally on color and magnitude in the CMD?
+In other words, isn't the assumption that there are low- and high-SNR stars (and that those are statistically drawn from the same distribution of physical parameters) in any color--magnitude cell?}
 \item[big data] We assume that we have large numbers of stars, large enough that
@@
   de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
+\BLedit{In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.} \BLcomment{please rephrase if this is unclear.
+I think it is worth clarifying this "under-deconvolution".
+I noticed people don't typically understand this when they hear/read it, and it is an important aspect of this paper.}
 \item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
@@
 
+\BLcomment{Is it the right place to comment that this whole approach also assumes that 1) the selection function is totally known (and neglected here), 2) all measurements and their errors are correct (unbiased and statistically representative)?
+This could also go in the discussion, but I personally see those as fundamental assumptions.}
+
 \subsection{Setting up the Inference}
@@
 
+\BLcomment{This whole section is highlighted because I have moved paragraphs around and also added small clauses here and there.
+I've tried to improve the logical motivation and presentation using the (great) paragraphs you wrote.
+Feel free to revert if you don't like it.}\BLedit{
+
 Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
+A first application of Bayes theorem leads to
+\begin{equation}
+p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
+\label{eq:bayes}
+\end{equation}
+where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the
+parallax, and $Z$ is a normalization constant \BLedit{(the model evidence of marginalized likelihood)}.
 The strength of our inference lies in how we generate our prior information of $\varpi_{\true}$.
-Supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
+Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
+This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
 The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
-The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
-We include this distance information in the inference of t$\varpi_{\true}$ by using the \cmd\ as our prior.
+The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
+We include this distance information in the inference of $\varpi_{\true}$ by using the \cmd\ as our prior.
 Specifically, we use the \cmd\ generated from the data themselves as our prior.
-Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
-This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the parallax of that star.
 
+In this context, equation~\ref{eq:bayes} is significantly complexified.
 Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
@@
 The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\mathbf{y}$.
+
+We now describe a particularity of our method, which will allow us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
 Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude
-
 \begin{equation}
@@
 \mathrm{true \, vector} \quad &\mathbf{Y}_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
-\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
+\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
 \mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
@@
 
-If life were simple, our inference would be a 1D problem using Bayes theorem and look something like:
-\begin{equation}
-p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
-\label{eq:bayes}
-\end{equation}
-where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the
-parallax, and $Z$ is a normalization constant.
-But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
 We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the data is
@@
 
-This likelihood is the probability of the data vector given the true vector, but we
-want to flip it around and infer the true vector and therefore the true parallax, or distance. We do this
-analogous to Equation \ref{eq:bayes}.
-
+We can now write the multidimensional analogous to Equation \ref{eq:bayes},
 \begin{equation}
-p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n)
+p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
 \end{equation}
 where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
+Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior.
+}
 
@@
 
-To generate our prior, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude.
-We fit this transfromed \cmd\ using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
+To generate our prior $p(\mathbf{Y}_n)$, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude.
+We fit this transformed \cmd\ using \xd\ which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are heterogeneous measurement uncertainties.
 It generates an estimate of the underlying distribution from which the uncertain data were drawn from.
-\xd\ generates a represention of the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalizing over the true data.
-
-\xd\ generates the gaussian parameters
-
+\xd\ generates a representation of the distribution as a $K$-mixture of Gaussians.
+In other words, the prior reads
+\BLedit{
+\begin{equation}
+	p\bigl(\mathbf{Y} \given  \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K \mathrm{amp}_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
+\end{equation}
+where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$ component, which is a Gaussian in our transformed \cmd\ space.
+The relative amplitudes between the components are described by the scalars $\{ \mathrm{amp}_k \}$.
+}
+Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data, \textit{i.e.}, finds the coefficients
 \begin{eqnarray}
-\alpha &=& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
+\alpha &=& \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
        &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
@@
 \end{eqnarray}
-where $\alpha$ encapsulates the parameters of each Gaussian in the mixture of Gaussians that defines our \cmd\ prior.
 
@@
 
-To infer $\varpi_{\true}$, we first do the inference in the 2D space of the \cmd. Using Bayes theorem,
-
-\begin{equation}
-p(\mathbf{Y} \given \mathbf{y}, \alpha) = \frac{1}{Z} \, p(\mathbf{y} \given \mathbf{Y}) \, p(\mathbf{Y} \given \alpha)
-\label{eq:posterior}
-\end{equation}
-We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
+\BLedit{To infer $\varpi_{\true}$, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.} We then need to project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
 
@@
 In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance.
-We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd and bias the parallax inference.
+We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference.
 For each distance, we sample the median of the probabilistic dust model.
@@
 safe, and it is a goal to explore full hierarchical modeling in the coming
-years (and see, for example, Leistedt et al. in preparation).
+years (and see, for example, \citealt{leistedtHogg2017}).
 

### 2017-05-24 paperDistance/paper.tex
minor edits and comments

@@
 Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion).
-Instead, we correct for dust using a point estimate from the 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$.
+Instead, we correct for dust using a point estimate from the 3D dust map \citep{green15} sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$.
 In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
@@
 We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution.
-We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
+\BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
 We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
@@
 Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
+
+\BLcomment{Also add the statistics about the parallax uncertainty shrinkage. Or put in a separate subsection? In fact, I now realize some of the figures are not discussed in the text. }
 
@@
 included in the \tgas--\tmass\ intersection.
+\BLcomment{I would specifically mention again the assumption of stationarity, and explain how it is broken by the selection function (in particular fluctuations in the detection probability of objects as a function of position, parallax, photometry, etc).
+Here, or in the critical discussion of assumptions.}
 
@@
 
+\BLcomment{it's weird to end with these subsections.
+Cross-validation should go in the methods part (I think there is a flag there already), and the discussion of assumptions should go in the discussion.
+There should be a conclusion and look towards the future.}
+
 \acknowledgments It is a pleasure to thank
@@
 It made use of the \acronym{NASA} Astrophysics Data System.
-All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license.
+All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license. \BLcomment{wrong link.}
 

### 2017-05-30 paperDistance/paper.tex

@@
 \end{equation}
-where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the
-parallax, and $Z$ is a normalization constant \BLedit{(the model evidence of marginalized likelihood)}.
-The strength of our inference lies in how we generate our prior information of $\varpi_{\true}$.
+where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
+parallax, and $Z$ is a normalization constant \BLedit{(the model evidence, or marginalized likelihood)}.
+The strength of our inference lies in how we generate our prior.
 Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
@@
 The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
-We include this distance information in the inference of $\varpi_{\true}$ by using the \cmd\ as our prior.
-Specifically, we use the \cmd\ generated from the data themselves as our prior.
+We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
+Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.
 
-In this context, equation~\ref{eq:bayes} is significantly complexified.
+Equation~\ref{eq:bayes} is too simplistic to describe this procedure.
 Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
-For inferring $\varpi_{\true}$, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
+Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
 Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$.
@@
 
-We now describe a particularity of our method, which will allow us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
+We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
 Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude
@@
 Gaussian uncertainties are required to generate our prior using \xd, but more about this and our dust corrections below.
-Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map. So our full terminology is
+Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
 
@@
 \mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
-\mathrm{assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
+\mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
 \end{aligned}
@@
 \end{equation}
-where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
+where $E(B-V)_n$ is the output of the dust model of \cite{green15}, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
 
-We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the data is
+We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
 \begin{equation}
@@
 with \\
-$V_n = $
 \[
-\begin{bmatrix}
+V_n = \begin{bmatrix}
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
@@
 So we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
-Again, this is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
 
@@
 where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
-Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior.
+Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves.
+We will use XD, which will take advantage of the Gaussian likelihood.
 }
@@
 
-To generate our prior $p(\mathbf{Y}_n)$, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude.
-We fit this transformed \cmd\ using \xd\ which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are heterogeneous measurement uncertainties.
-It generates an estimate of the underlying distribution from which the uncertain data were drawn from.
-\xd\ generates a representation of the distribution as a $K$-mixture of Gaussians.
-In other words, the prior reads
+To generate our prior $p(\mathbf{Y}_n)$, we fit our data in the slightly transformed \cmd\ space using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are Gaussian, heterogeneous measurement uncertainties.
+It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a $K$-mixture of Gaussians:
 \BLedit{
 \begin{equation}
-	p\bigl(\mathbf{Y} \given  \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K \mathrm{amp}_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
+	p\bigl(\ \mathbf{Y} \given  \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K \mathrm{amp}_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
 \end{equation}
-where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$ component, which is a Gaussian in our transformed \cmd\ space.
+where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
 The relative amplitudes between the components are described by the scalars $\{ \mathrm{amp}_k \}$.
 }
-Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data, \textit{i.e.}, finds the coefficients
+Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data. In equations, it finds the coefficients
 \begin{eqnarray}
-\alpha &=& \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
+\alpha &=& \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \nonumber\\
        &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
@@
 
+which is made analytically tractable with the Gaussian likelihood and prior \citep{bovy11}.
+
 \subsection{Back to the Inference}
 
-\BLedit{To infer $\varpi_{\true}$, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.} We then need to project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
+\BLedit{To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.} We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
 
@@
 
-Crossvalidation?
+TODO: talk about crossvalidation?
 
@@
 
+\BLcomment{Needs to say much more - unfortunately I cannot complete this part since I don't know what you've done in enough detail.}
+
 \subsection{Dust}
@@
 
+\BLcomment{Isn't this repeating what was described in the previous paragraph?}
 We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust.
@@
 
-\BLcomment{Also add the statistics about the parallax uncertainty shrinkage. Or put in a separate subsection? In fact, I now realize some of the figures are not discussed in the text. }
+\BLcomment{Figure \ref{fig:prior} needs to be discussed here or in a separate subsection.
+I also find that it is the right place to discuss its features.
+I find it strange to discuss the posterior parallaxes without describing the prior.}
+
+\BLedit{
+\subsection{Shrinkage of parallax uncertainties}
+
+Figure~\ref{fig:posterior} shows the end result of our inference for a few randomly-chosen objects: the posterior pdf on the true parallaxes. 
+We also show the likelihood and the prior. 
+In all cases, the prior information brought by the \cmd\ is significant, and leads to a compelling decrease in the parallax uncertainty.
+The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact strongly depend on the location of each object in \cmd\ space.
+
+Figure~\ref{fig:posteriorCMD} shows the distribution of the parallaxes derived with our \cmd\ prior.
+This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
+Those typically are the regions with the largest photometric and parallax errors, for the faintest objects. 
+This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
+
+}
+\BLcomment{Need to discuss the implications of that.}
 
@@
   \includegraphics[width=\textwidth]{prior_ngauss128.png}
-\caption{Prior: the XD generated mixture of gaussians we use as our prior.
-The left panel shows a sampling of the prior, which is a mixture of gaussians shown in the right hand panel.
-The blue lines are the $1\sigma$ and $2\sigma$ contours.
-The gaussians have a funny shape due to the transformed space in which we fit the data.}
+\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
+The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
+The right panel shows the 1-$\sigma$ contours of the individual components.
+The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
+Thus, they appear non-Gaussian in the physical color--magnitude space.}}
 \label{fig:prior}
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{the expectation values and subsampled error bars.
+\caption{The expectation values and subsampled error bars.
 Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd.

### 2017-05-30 comparePrior.py
added stellar twins

@@ -561,9 +561,56 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
 
         figVarDiff.savefig('denoisedVarianceSamples_' + file.split('.')[0] + '.png')
 
+def paperComparePrior(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', dataFilename='All.npz', contourColor='k', posteriorColor='royalblue'):
+    setup_text_plots(fontsize=16, usetex=True)
+    tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+    xdgmm = XDGMM(filename=xdgmmFilename)
+    absmag = 'J'
+    mag1 = 'J'
+    mag2 = 'K'
+    xlabel = '$(J-K)^C$'
+    ylabel = r'$M_J^C$'
+    xlim = [-0.25, 1.25]
+    ylim = [6, -6]
+
+    ndim = 2
+    data = np.load(dustFile)
+    dustEBV = data['ebv']
+    absMagKinda, apparentMagnitude = testXD.absMagKindaArray(absmag, dustEBV, bandDictionary, tgas['parallax'])
+
+    color = testXD.colorArray(mag1, mag2, dustEBV, bandDictionary)
+    color_err = np.sqrt(bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']]**2. + bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']]**2.)
+
+    postFile = 'posteriorParallax.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+    yim = (-1, 5)
+
+    indices = np.random.randint(0, high=len(color), size=1024)
+
+    fig, ax = plt.subplots(1, 2)
+    for i, file in enumerate(['posteriorSimple.npz', postFile]):
+        data = np.load(file)
+        posterior = data['posterior']
+        sigma = np.sqrt(data['var'])
+        mean = data['mean']
+        absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))
+        absMagSigma = testXD.absMagKinda2absMag(sigma*10.**(0.2*apparentMagnitude))
+        ax[0].scatter(color[indices], absMag[indices])
[…]

### 2017-05-30 paper/paper.tex
two BL comments

--- 2017-05-30 paperDistance/paper.tex
@@
 models of stars:
-The distance to a star is a latent variable, and the nearby
-stars constitute a high signal-to-noise ``training set''.
-These nearby stars can be used to anchor empirical models that can
-then be applied to the more distant stars\BLedit{, thanks to the simple physical relations connecting the measurements (apparent magnitudes, parallaxes) to some of the physical parameters of interest (absolute magnitudes, distances).}
+The different stars observed by \gaia\ are measured at very different
+signal-to-noise ratios.
+A data-driven model of the stellar color--magnitude distribution can capitalize
+on the precise information coming from the high signal-to-noise stars
+and deliver valuable information about the low signal-to-noise stars,
+which in turn will deliver precise parallax (or distance or absolute magnitude) inferences.
 
@@
 data-driven model of the \gaia\ data could in principle deliver very
-precise parallaxes, without any input of the physics or numerical models of stellar
+precise parallax measurements, without any input of the physics or numerical models of stellar
 evolution, interiors, or photospheres.
@@
 Setting up the problem in more detail, we have the true samples from the underlying distribution and the noisy samples
-
 \begin{eqnarray}
 y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
-y_n \, &=& \, y_{\true,n} + \epsilon_n
+y_n \, &=& \, y_{\true,n} + \epsilon_n \quad,
 \label{eq:ytrue}
 \end{eqnarray}
-
 where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
@@
 Here our likelihood for each individual datum is
-
-\BLedit{
 \begin{eqnarray}
 p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \mathrm{d}y_{\true, n} \\
-&=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)
+&=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2) \quad,
 \label{eq:toyLike}
 \end{eqnarray}
-
-where we have marginalized over the true sample $y_{\true, n}$.} The resulting likelihood is a Gaussian because the noise model is a Gaussian.
+where we have marginalized over the true sample $y_{\true, n}$.
+The resulting likelihood is a Gaussian because the noise model is a Gaussian.
 \BLcomment{should all these equations have $x_n$ on the right hand side?
 I don't have a strong opinion about that.} The likelihood of the full data set is then the product of the individual likelihoods
-
 \begin{equation}
-\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2).
+\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
 \label{eq:toyLikeFull}
 \end{equation}
-
 Maximizing this joint likelihood of all the data is the same as minimizing
-
 \begin{equation}
 \chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,\ln(2\pi\,[\sigma_n^2 + t^2])
+\quad .
 \label{eq:chisq}

### 2017-05-30 paper/paper.tex
three BLs
bls

@@
 
-%@archiver{PUT FIGURE NAME HERE}
+%@archiver{posteriorCMD.png}
 
@@
 The resulting likelihood is a Gaussian because the noise model is a Gaussian.
-\BLcomment{should all these equations have $x_n$ on the right hand side?
-I don't have a strong opinion about that.} The likelihood of the full data set is then the product of the individual likelihoods
+The likelihood of the full data set is then the product of the individual likelihoods
 \begin{equation}
@@
   stars measured at lower signal-to-noise have analogs measured at
-  higher signal-to-noise.  In detail, this assumption is wrong, (for example, distant halo stars are different than local
-  disk stars), but the stationarity assumption is fairly weak; as long
-  as there is support in the high signal-to-noise stars for the types
-  of stars seen at low signal-to-noise, the results will not be
-  strongly biased. Although a weak assumption, it is deep and fundamental to this
+  higher signal-to-noise.
+  In detail, this assumption is wrong, (for example, distant halo stars are different than local
+  disk stars), but the stationarity assumption is fairly weak:
+  The requirement is that, locally in the \cmd, there is
+  support among the high signal-to-noise stars for the types
+  of stars also seen at low signal-to-noise.
+  If this requirement is met, the results will not be
+  strongly biased. Although this is a weak assumption, it is deep and fundamental to this
   work.
-\BLcomment{Isn't that assumption made conditionally on color and magnitude in the CMD?
-In other words, isn't the assumption that there are low- and high-SNR stars (and that those are statistically drawn from the same distribution of physical parameters) in any color--magnitude cell?}
 \item[big data] We assume that we have large numbers of stars, large enough that
@@
   de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
-\BLedit{In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.} \BLcomment{please rephrase if this is unclear.
-I think it is worth clarifying this "under-deconvolution".
-I noticed people don't typically understand this when they hear/read it, and it is an important aspect of this paper.}
+In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
 \item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
@@
   \cmd\ is somehow smooth and compact).
-\end{description}
+\enyd{description}
 

### 2017-05-30 paper/paper.tex
median factor better

@@
 The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
-Our posterior photometric parallax estimates are more precise than the \tgas\ catalog entries by a median factor of [XXX]; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+Our posterior photometric parallax estimates are more precise than the \tgas\ catalog entries by a median factor of 1.2, and a 75, 90, and 99 percentile factor of 1.6, 2.3, and 4.9, respectively; and more precise than previous examples of Bayesian distance estimates by a median factor of 1.2, and a 75, 90 and 99 percentile factor of 1.5, 1.6, and 2.1, respectively.
 %The precision can be attributed to the statistics concept of shrinkage.

### 2017-05-30 paper/paper.tex
fixed all XXXs and YYYs

@@
 We independently validate our distances by looking at members of a Milky Way star cluster M67, which is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly as a cluster in our posterior parallax estimates.
-All our results, including a posterior parallax and distance sampling for [ZZZ] \tgas\ stars, are available in companion electronic tables.
+All our results, including a posterior parallax and distance sampling for 1.4 million \tgas\ stars, are available in companion electronic tables.
 \end{abstract}
@@
 The blue lines are the $1\sigma$ and $2\sigma$ contours.
-The median fractional change is XXX.
-Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^C$ about XXX.
+The median fractional change is 1.2.
+Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^C$ of about 0.6.
 The right panel shows the posterior point estimates on the CMD colored by fraction change in variance.
@@
 It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
-It is a fairly broad prior so the median \gaia\ star has an increased posterior variance of XXX.
+It is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
 It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.

### 2017-06-01 paper/paper.tex
typo
Boris

@@
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
 \affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
 \affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
+
+\author{Boris Leistedt}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
 
@@
 %\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
-
-%\author{Boris Leistedt}
-%\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 
@@
   \cmd\ is somehow smooth and compact).
-\enyd{description}
+\end{description}
 
@@
 
-Anderson: Compare with literature stellar models?? This is an enhanced goal; strike this.
-
 \BLcomment{it's weird to end with these subsections.
@@
 It made use of the \acronym{NASA} Astrophysics Data System.
-All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license. \BLcomment{wrong link.}
+All the code used in this project is available online\footnote{\url{https://github.com/andersdot/photoParallax}} under an open-source license.
 

### 2017-06-02 paper/paper.tex
boris
prior comparison plot

@@
 
-\author{Boris Leistedt}
+\author[0000-0002-3962-9274]{Boris Leistedt}
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
@@
 \centering
-  \includegraphics[width=\textwidth]{deltaSimple.png}
-\caption{Other example prior: The same visualization as Figure \ref{fig:delta} of the change in the posterior, but here, if instead of prior information containing the \cmd\, the prior information is an exponentially declining spatial density of stars.
-}
-\label{fig:otherPrior}
+  \includegraphics[width=\textwidth]{comparePrior.png}
+\caption{Comparison with common prior: A visualization of the posterior pdf for a subsection of TGAS stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
+On the left is the posterior pdf using our \cmd\ prior, and on the right is the posterior pdf when using the exponentially declining spatial density, a common prior used when inferring parallaxes.
+The scale of the exponentially declining spatial density prior is XXX, the optimal value found in YYY.
+Using the \cmd\ as a prior generates a posterior pdf that has smaller variance.}
+\label{fig:comparePrior}
 \end{figure}
@@
 
+\section{Conclusion}
+Data driven prior of \cmd\ in parallax inference is a good idea
+
 \acknowledgments It is a pleasure to thank

### 2017-06-02 paper/paper.tex
conclusion

@@
 On the left is the posterior pdf using our \cmd\ prior, and on the right is the posterior pdf when using the exponentially declining spatial density, a common prior used when inferring parallaxes.
-The scale of the exponentially declining spatial density prior is XXX, the optimal value found in YYY.
+The scale of the exponentially declining spatial density prior is 1.35 kpc, the optimal value found in \cite{astraatmadja16a}.
 Using the \cmd\ as a prior generates a posterior pdf that has smaller variance.}
@@
 \section{Conclusion}
-Data driven prior of \cmd\ in parallax inference is a good idea
+If you want a conclusion, re-read the abstract
 

### 2017-06-02 paper/paper.tex
parallax error
tweaks ;
dust figure

@@
   the parallax and color uncertainties are Gaussian in form with
-  correctly known variances. In detail, we assume that the [INSERT
-    CORRECT TERMINOLOGY FROM GAIA HERE] parallax uncertainty estimates
-  are correct. Here we are making use of the fact that, for
+  correctly known variances. In detail, we assume that the parallax uncertainty estimates, the parallax{\_}error entry in the \gaia\ catalogue, are correct.
+  Here we are making use of the fact that, for
   de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
@@
 We then regenerate our prior using the lightly dust corrected photometry.
-We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution.
+The dust values we obtain in our final iteration, and which we apply to our photometry, are visualized in galactic coordinates in Figure \ref{fig:dust}.
+After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution.
 \BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.

### 2017-06-02 paper/paper.tex
data figure

@@
 We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
+The full data set is visualized in the \cmd\ in Figure \ref{fig:data}.
+The left panel shows the point estimates of the color and absolute magnitude, using the point estimate of the parallax from \gaia, and the $1\sigma$ and $2\sigma$ contours of the distribution.
+The right panel shows a subset of the data with it's associated error bars.
+The uncertainties in the colors are fairly well behaved, but the large uncertainties in the absolute magnitude are due to the large uncertainties in the parallax measurements.
 

### 2017-06-02 paper/paper.tex
prior figure

@@
 
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{prior_ngauss128.png}
+\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
+The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
+The right panel shows the 1-$\sigma$ contours of the individual components.
+The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
+Thus, they appear non-Gaussian in the physical color--magnitude space.}}
+\label{fig:prior}
+\end{figure}
+
+\subsection{prior}
+Using \xd\ and the MML method described in the methods section, we generate our prior shown in Figure \ref{fig:prior}.
+This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution.
+The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from.
+The main sequence and red giant branch are much tighter, as well as the red clump.
+We'll come back to more features in the data below, especially those seen in the posterior distribution.
+
 \BLcomment{Figure \ref{fig:prior} needs to be discussed here or in a separate subsection.
@@
 RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
-
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{prior_ngauss128.png}
-\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
-The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
-The right panel shows the 1-$\sigma$ contours of the individual components.
-The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
-Thus, they appear non-Gaussian in the physical color--magnitude space.}}
-\label{fig:prior}
-\end{figure}
 

### 2017-06-02 paper/paper.tex
prior comparison figure

@@
 %assumptions, in fact! What form should this comparison be in?
-There are other options for prior information when inferring distances to stars.
-One option is the exponentially declining spatial density of stars (cite: Tri, CBJ).
+There are other options for prior information to include when inferring distances to stars.
+One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
 It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
@@
 It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
-It is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
+A comparison of the posteriors using the two different priors, our \cmd\ prior and the EDSD, is shown in Figure \ref{fig:comparePrior}.
+Here the parallax posterior pdf expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
+The left panel visualizes results using our \cmd\ prior, and the right panel visualizes results using the EDSD prior. 
+The variance for the \cmd\ prior is significantly smaller than the variance for the EDSD prior.
+In general, it is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
 It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
-Given this prior, what does the observed \cmd\ look like?
-The \cmd\ is broad and the red clump highly stretched.
+%LMA: chat about bias?
 

### 2017-06-02 paper/paper.tex
whats that feature figure

@@
 The denoised expectation values of the parallax projected into the \cmd\ show interest features.
+We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
 \begin{itemize}
@@
 The lower main sequence is very narrow while the upper main sequence is much wider.
-This may be an age variation on the main sequence mixed with a malquist bias of seeing a much larger volume of upper main sequence stars.
+This may be an age variation on the main sequence mixed with a malmquist bias of seeing a much larger volume of upper main sequence stars.
 There is a plausible binary sequence of stars with about twice the brightness as the main sequence.

### 2017-06-02 paper/paper.tex
abstract and acks

--- 2017-06-02 paper/paper.tex
@@
 % - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
-% - deal with all HOGG or Anderson or LMA in the text.
+% - deal with all DWH or LMA in the text.
 % - deal with everything in the google doc.
@@
 
+\author{Adrian Price-Whelan}
+\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
+
 %\author{Jo Bovy}
@@
 
-%\author{Adrian Price-Whelan}
-%\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
-
 \begin{abstract}\noindent % trust me
-
-Conversion of a noisy parallax measurement into a posterior belief over distances requires inference with a prior.
+Conversion of a noisy parallax measurement into a posterior
+belief over distances requires inference with a prior.
 Usually this prior represents some kind of belief about the Milky Way.
@@
 The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
-Our posterior photometric parallax estimates are more precise than the \tgas\ catalog entries by a median factor of 1.2, and a 75, 90, and 99 percentile factor of 1.6, 2.3, and 4.9, respectively; and more precise than previous examples of Bayesian distance estimates by a median factor of 1.2, and a 75, 90 and 99 percentile factor of 1.5, 1.6, and 2.1, respectively.
-%The precision can be attributed to the statistics concept of shrinkage.
+Our posterior photometric parallax estimates are more precise than the
+\tgas\ catalog entries by a median factor of 1.2, and 20 percent (LMA check)
+are more precise by a factor of two or more; they are also more precise than
+previous examples of Bayesian distance estimates using a spatial prior.
 We independently validate our distances by looking at members of a Milky Way star cluster M67, which is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly as a cluster in our posterior parallax estimates.
@@
 We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
-The full data set is visualized in the \cmd\ in Figure \ref{fig:data}.
-The left panel shows the point estimates of the color and absolute magnitude, using the point estimate of the parallax from \gaia, and the $1\sigma$ and $2\sigma$ contours of the distribution.
-The right panel shows a subset of the data with it's associated error bars.
-The uncertainties in the colors are fairly well behaved, but the large uncertainties in the absolute magnitude are due to the large uncertainties in the parallax measurements.
 
@@
 We then regenerate our prior using the lightly dust corrected photometry.
-The dust values we obtain in our final iteration, and which we apply to our photometry, are visualized in galactic coordinates in Figure \ref{fig:dust}.
-After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution.
+We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution.
 \BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
@@
 Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
-
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{prior_ngauss128.png}
-\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
-The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
-The right panel shows the 1-$\sigma$ contours of the individual components.
-The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
-Thus, they appear non-Gaussian in the physical color--magnitude space.}}
-\label{fig:prior}
-\end{figure}
-
-\subsection{prior}
-Using \xd\ and the MML method described in the methods section, we generate our prior shown in Figure \ref{fig:prior}.
-This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution.
-The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from.
-The main sequence and red giant branch are much tighter, as well as the red clump.
-We'll come back to more features in the data below, especially those seen in the posterior distribution.
 
@@
 Using a data driven model of the observed \cmd\ we inferred a posterior belief of the true parallax of each star. Given our method, we call our posterior parallax the photometric parallax.
-HOGG: Placeholder for the output format for our data products.
+DWH: Placeholder for the output format for our data products.
 RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{prior_ngauss128.png}
+\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
+The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
+The right panel shows the 1-$\sigma$ contours of the individual components.
+The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
+Thus, they appear non-Gaussian in the physical color--magnitude space.}}
+\label{fig:prior}
+\end{figure}
 
@@
 \begin{description}
-\item[stationarity] (HOGG needs to clean this paragraph)
+\item[stationarity] (DWH needs to clean this paragraph)
   We have assumed that (locally) the morphology of
@@
   pulled to mimic a much farther away MS star.
-\item[big data] HOGG
-\item[noise model] HOGG
+\item[big data] DWH
+\item[noise model] DWH
 \item[dust] Within our probabilistic framework, we improperly handle
@@
   scope of this paper.
-\item[mixture of Gaussians] HOGG
-\item[no physics] HOGG
+\item[mixture of Gaussians] DWH
+\item[no physics] DWH
 \end{description}
@@
 \section{Conclusion}
-If you want a conclusion, re-read the abstract
+If you want to read a conclusion, re-read the abstract!
 
@@
   Andy Casey (Monash),
-  Dustin Lang (Toronto) for cross matched 2MASS-TGAS catalogue
+  Stephen Feeney (Flatiron),
+  Dustin Lang (Toronto),
   Jo Bovy (Toronto)
-  Boris Leistedt (NYU)
-  Adrian Price-Whelan (Princeton)
 and the attendees at the Stars Group Meeting at the Flatiron Institute

### 2017-06-02 paper/paper.tex

--- 2017-06-02 paper/paper.tex
@@
 Using a data driven model of the observed \cmd\ we inferred a posterior belief of the true parallax of each star. Given our method, we call our posterior parallax the photometric parallax.
-HOGG: Placeholder for the output format for our data products.
+DWH: Placeholder for the output format for our data products.
 RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)

### 2017-06-02 paper/paper.tex
orcid
cross validation moved
fraction better by factor of 2

@@
 
-\author{Adrian Price-Whelan}
+\author[0000-0003-0872-7098]{Adrian Price-Whelan}
 \affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
@@
 Our posterior photometric parallax estimates are more precise than the
-\tgas\ catalog entries by a median factor of 1.2, and 20 percent (LMA check)
+\tgas\ catalog entries by a median factor of 1.2, and 14 percent
 are more precise by a factor of two or more; they are also more precise than
@@
 
-TODO: talk about crossvalidation?
+\subsection{Cross Validation}
+For generating our prior, we use a mixture of 128 Gaussians.
+This number, 128, was chosen heuristically.
+It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product.
+We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.
 
@@
 
-\subsection{Cross Validation}
-For generating our prior, we use a mixture of 128 Gaussians.
-This number, 128, was chosen heuristically.
-It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product.
-We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.
-
-\BLcomment{it's weird to end with these subsections.
-Cross-validation should go in the methods part (I think there is a flag there already), and the discussion of assumptions should go in the discussion.
-There should be a conclusion and look towards the future.}
-
 \section{Conclusion}

### 2017-06-02 paper/paper.tex
assumptions

--- 2017-06-02 paper/paper.tex
@@
 
-To infer the true parallax to a star using the noise-deconvolved \cmd\ as a prior, we make the following assumptions. The method we present is correct
+To infer the true parallax to a star using the noise-deconvolved \cmd\ as
+a prior, we make the following assumptions. The method we present is correct
 and justifiable under these assumptions, each of which is questionable
@@
   work.
+\item[selection] We have no model for the selection function of \tgas,
+  nor selection volumes for any kinds of stars.
+  For this reason, we are building a model of \emph{the contents of \tgas}, not the
+  properties of the stars in the Milky Way (nor any volume-limited component or region
+  thereof).
 \item[big data] We assume that we have large numbers of stars, large enough that
@@
   the parallax and color uncertainties are Gaussian in form with
-  correctly known variances. In detail, we assume that the parallax uncertainty estimates, the parallax{\_}error entry in the \gaia\ catalogue, are correct.
+  correctly known variances. In detail, we assume that the parallax uncertainty estimates,
+  (the \texttt{parallax{\_}error} entry in the \gaia\ catalog), are correct.
   Here we are making use of the fact that, for
@@
 \end{description}
-
-\BLcomment{Is it the right place to comment that this whole approach also assumes that 1) the selection function is totally known (and neglected here), 2) all measurements and their errors are correct (unbiased and statistically representative)?
-This could also go in the discussion, but I personally see those as fundamental assumptions.}
 
@@
   pulled to mimic a much farther away MS star.
+\item[selection] DWH
 \item[big data] DWH
@@
 \end{description}
-
-\subsection{Cross Validation}
-For generating our prior, we use a mixture of 128 Gaussians.
-This number, 128, was chosen heuristically.
-It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product.
-We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.
-
-\BLcomment{it's weird to end with these subsections.
-Cross-validation should go in the methods part (I think there is a flag there already), and the discussion of assumptions should go in the discussion.
-There should be a conclusion and look towards the future.}
 

### 2017-06-02 paper/paper.tex
discussion rearrange

@@
 
-\BLcomment{This whole section is highlighted because I have moved paragraphs around and also added small clauses here and there.
-I've tried to improve the logical motivation and presentation using the (great) paragraphs you wrote.
-Feel free to revert if you don't like it.}\BLedit{
-
 Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
@@
 where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
-parallax, and $Z$ is a normalization constant \BLedit{(the model evidence, or marginalized likelihood)}.
+parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
 The strength of our inference lies in how we generate our prior.
@@
 We will use XD, which will take advantage of the Gaussian likelihood.
-}
 
@@
 
-\BLedit{To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.} We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
+To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.
+We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
 
@@
 
-\BLcomment{Needs to say much more - unfortunately I cannot complete this part since I don't know what you've done in enough detail.}
-
 \subsection{Dust}
@@
 For each distance, we sample the median of the probabilistic dust model.
-We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD).
+We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (LMA REF TABLE SFD).
 We then regenerate our prior using the lightly dust corrected photometry.
@@
 
-\BLedit{
 \subsection{Shrinkage of parallax uncertainties}
@@
 
-}
 \BLcomment{Need to discuss the implications of that.}
@@
 
-Transition clause/sentence. We obtained the photometric parallaxes in this project
+We obtained the photometric parallaxes in this project
 by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
@@
 inferences bring in new information that decrease variance (but
-often introduce bias).
+can introduce bias).
 We have shown that, in addition to pure precision improvements, at
-least some aspects of \emph{accuracy} have been improved as well, by
-showing that the posterior distance estimates to stars in (at least one) stellar
+least some aspects of \emph{accuracy} have been improved as well:
+The posterior distance estimates to stars in (at least one) stellar
 cluster are much more clustered than likelihood-based distance
@@
 
-Transition clause/sentence. Although we have not performed principled hierarchical Bayesian inference in this
+Although we have not performed principled hierarchical Bayesian inference in this
 project, we built our prior for each star's parallax by
@@
 Because the method we use, \xd, which is
-an Empirical Bayesian maximum-marginalized-likelihood estimato, accounts for
+an Empirical Bayesian maximum-marginalized-likelihood estimator, accounts for
 heteroskedastic noise, we were able to build our photometric parallax
@@
 we generate from our parallax posterior pdfs to be only weakly biased.
-
-Transition clause/sentence. The empirical Bayesian methodology is a good approximation to full
-hierarchical inference in the limit of large numbers of data points, such that no individual star is carrying a lot of weight in the
-model of the \cmd.
-With its millions of stars, the \tgas\ Catalog appears to safely
-live in this limit.
-However, because there are parts of the \cmd\ that are poorly populated,
-even in \tgas, the empirical Bayes approximation may be bad for some
-portions of the \cmd.
-In particular, there appear to be essentially no white dwarf stars,
-and other classes (like bright red giant branch stars) have few to no exemplars
-in the data set at high signal-to-noise (in parallax).
-For all these reasons, we do think that this approximation is not perfectly
-safe, and it is a goal to explore full hierarchical modeling in the coming
-years (and see, for example, \citealt{leistedtHogg2017}).
-
-Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
-function or completeness estimate or inverse-selection-volume corrections.
-This model is a model of the \tgas\ Catalog
-(or really the \tgas--\tmass\ intersection), not of the stars in
-the Milky Way, nor any volume-limited subsample of the Milky Way.
-Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
-intersection, the model \emph{is} representive of the full intersection,
-even the distant parts, where all the parallaxes are measured at low
-signal-to-noise.
-This is in contrast to techniques that might build the model from only
-the high signal-to-noise sources; such a model would be biased towards
-stellar types and compositions found locally, as well as lower-luminosity stars,
-and stars which happened to get good parallaxes.
-However, our use of the entire \tgas--\tmass\ intersection without
-any selection function restricts the use of our \cmd\ model as a prior
-to inferences within that same \tgas--\tmass\ intersection.
-It would be biasing to use this same prior to infer
-distances for the full billion-star catalog released in \gaia\ DR1
-alongside \tgas, or for stars fainter or brighter or bluer than those
-included in the \tgas--\tmass\ intersection.
-\BLcomment{I would specifically mention again the assumption of stationarity, and explain how it is broken by the selection function (in particular fluctuations in the detection probability of objects as a function of position, parallax, photometry, etc).
-Here, or in the critical discussion of assumptions.}
-
-Transition clause/statement. The outputs of this project include posterior pdfs for stellar parallaxes.
-It is tempting to treat these outputs as equivalent to some kind of catalog of
-measurements, as we treated the original parallax measurements from \tgas.
-That is, it is tempting to treat these as simply ``better'' measurements.
-Taken one star at a time, this is permitted and true, and the basis for our
-claim that we have de-noised the \tgas\ Catalog.
-However, there are important differences, as there are with all
-probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
-measurements and the de-noised measurements.
-For users of our output, the most important difference between our
-output and the \tgas\ data input is that the \tgas\ data is the representation
-of a likelihood function, whereas our ouptut is a
-representation of a posterior pdf, one per star.
-Posterior information used naively can lead to serious statistical
-errors, because each data point has had a prior multiplied in.
-Multiplicative uses of the data will effectively take the prior
-pdf to a large power.
-We have (in previous work)
-given examples of correct uses of posterior samplings or pdfs
-for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
-encourage power-users to consult those methodological
-contributions before engaging with these outputs.
 
@@
 \subsection{What's That Feature?}
+The denoised expectation values of the parallax projected into the \cmd\ show interest features.
+We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
 \begin{figure}
@@
 
-The denoised expectation values of the parallax projected into the \cmd\ show interest features.
-We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
-\begin{itemize}
-\item (1) Narrow Lower Main Sequence
-\item (2) Wide Upper Main Sequence
-\item (3) Plausible Binary Sequence
-\item (4) Prominent Red clump
-\item (5) Narrow Red Giant Branch
-\item (6) Noticeable turn off
-\end{itemize}
-The lower main sequence is very narrow while the upper main sequence is much wider.
-This may be an age variation on the main sequence mixed with a malmquist bias of seeing a much larger volume of upper main sequence stars.
-There is a plausible binary sequence of stars with about twice the brightness as the main sequence.
-The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is fairly straight and narrow.
-The main sequence turn off is noticeable and also surprisingly narrow, possibly a reflection of the star formation history of the local volume also mixed with a malmquist bias.
+The lower main sequence is very narrow while the upper main sequence
+is much wider. This may be an age variation on the main sequence mixed
+with a malmquist bias of seeing a much larger volume of upper main
+sequence stars. There is a plausible binary sequence of stars with
+about twice the brightness as the main sequence. The helium burning
+red clump is prominent for the volume DR1 probed, and the red giant
+branch is fairly straight and narrow. The main sequence turn off is
+noticeable and also surprisingly narrow, possibly a reflection of the
+star formation history of the local volume also mixed with a malmquist
+bias.
 
@@
 Here we return to these assumptions, and discuss them critically in more
-detail.
+detail, under the same set of assumption labels:
 \begin{description}
@@
   drawn from the same distribution within some feature or patch of the
-  space. This is all assuming the Gaia noise model is correct. If the
-  uncertainties are over estimated then we could be over deconvolving
-  (generating a tighter distribution than the truth) And in some
-  extreme over estimate of uncertainties the lower signal to noise
-  giants could be pulled to the main sequence. But with the quoted
-  uncertainties this seems unlikely. As you move up the giant branch,
-  and probe farther giants, the signal to noise will continue to
-  diminish, but because the color uncertainties are not large, the S/N
-  spread within a patch of the \cmd\ is well mixed. Possibly a
-  metallicity trend with distance, placing more nearby (higher S/N)
-  objects at slightly different positions on the CMD pulling lower S/N
-  objects to that distribution for our prior. But if something has
-  small error bars it is good data is good data both for inference and
-  generating the prior.
-
-  For very rare kinds of stars. Inference with our prior is bad for
-  underrepresented populations in the \cmd. For example, it appears
-  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
-  underrepresented in the prior, the amplitudes of the gaussians
-  representing them is very small, so the WDs are more likely to be
-  pulled to mimic a much farther away MS star.
-\item[selection] DWH
-\item[big data] DWH
-\item[noise model] DWH
+  \cmd\ space.
+\item[selection] The biggest limitation of our method for generating
+  the \cmd\ from the data is that we made no use of any kind of
+  \tgas\ selection function or completeness estimate or
+  inverse-selection-volume corrections.  This model is a model of the
+  \tgas\ Catalog (or really the \tgas--\tmass\ intersection), not of
+  the stars in the Milky Way, nor any volume-limited subsample of the
+  Milky Way.  Importantly, because we used \emph{all} of the stars in
+  the \tgas--\tmass\ intersection, the model \emph{is} representive of
+  the full intersection, even the distant parts, where all the
+  parallaxes are measured at low signal-to-noise.  This is in contrast
+  to techniques that might build the model from only the high
+  signal-to-noise sources; such a model would be biased towards
+  stellar types and compositions found locally, as well as
+  lower-luminosity stars, and stars which happened to get good
+  parallaxes.  However, our use of the entire
+  \tgas--\tmass\ intersection without any selection function restricts
+  the use of our \cmd\ model as a prior to inferences within that same
+  \tgas--\tmass\ intersection.  It would be biasing to use this same
+  prior to infer distances for the full billion-star catalog released
+  in \gaia\ DR1 alongside \tgas, or for stars fainter or brighter or
+  bluer than those included in the \tgas--\tmass\ intersection.
+\item[big data] The empirical Bayesian methodology is a good
+  approximation to full hierarchical inference in the limit of large
+  numbers of data points, such that no individual star is carrying a
+  lot of weight in the model of the \cmd.  With its millions of stars,
+  the \tgas\ Catalog appears to safely live in this limit.  However,
+  because there are parts of the \cmd\ that are poorly populated, even
+  in \tgas, the empirical Bayes approximation may be bad for some
+  portions of the \cmd.  In particular, there appear to be essentially
+  no white dwarf stars, and other classes (like bright red giant
+  branch stars) have few to no exemplars in the data set at high
+  signal-to-noise (in parallax).  For all these reasons, we do think
+  that this approximation is not perfectly safe, and it is a goal to
+  explore full hierarchical modeling in the coming years (and see, for
+  example, \citealt{leistedtHogg2017}).
+\item[noise model] Everything in both the establishment of the data-driven
+  prior and the use of it for parallax inference has been done under the deep
+  assumption that the \gaia\ noise model is correct. If the
+  uncertainties are over-estimated then we could be over-deconvolving.
+  That is, the data-driven model could be producing a tighter or more
+  informative distribution than what would truly be observed in a much
+  higher signal-to-noise experiment or survey. That is, incorrect uncertainty
+  estimates will bias the empirical-Bayes prior we obtain. On the inference
+  side, wrong uncertainty estimates lead to further biases in the derived photometric parallaxes.
 \item[dust] Within our probabilistic framework, we improperly handle
@@
   scope of this paper.
-\item[mixture of Gaussians] DWH
+\item[mixture of Gaussians] In everything in this project DWH
 \item[no physics] DWH
 \end{description}
+
+The outputs of this project include posterior pdfs for stellar parallaxes.
+It is tempting to treat these outputs as equivalent to some kind of catalog of
+measurements, as we treated the original parallax measurements from \tgas.
+That is, it is tempting to treat these as simply ``better'' measurements.
+Taken one star at a time, this is permitted and true, and the basis for our
+claim that we have de-noised the \tgas\ Catalog.
+However, there are important differences, as there are with all
+probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
+original \tgas\ measurements and the de-noised measurements.
+For users of our output, the most important difference between our
+output and the \tgas\ data input is that the \tgas\ data is the representation
+of a likelihood function, whereas our ouptut is a
+representation of a posterior pdf, one per star.
+Posterior information used naively can lead to serious statistical
+errors, because each data point has had a prior multiplied in.
+Multiplicative uses of the data will effectively take the prior
+pdf to a large power.
+We have (in previous work)
+given examples of correct uses of posterior samplings or pdfs
+for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
+encourage power-users to consult those methodological
+contributions before engaging with these outputs.
 
@@
   Ana Bonaca (Harvard),
+  Jo Bovy (Toronto),
   Andy Casey (Monash),
@@
   Dustin Lang (Toronto),
-  Jo Bovy (Toronto)
+  David Spergel (Flatiron),
 and the attendees at the Stars Group Meeting at the Flatiron Institute

### 2017-06-02 paper/paper.tex
discussion

@@
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
-
 \begin{equation}
@@
 
-We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
+We assume (as \gaia\ suggests LMA REFERENCE HERE) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
 \begin{equation}
@@
 To generate our prior $p(\mathbf{Y}_n)$, we fit our data in the slightly transformed \cmd\ space using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are Gaussian, heterogeneous measurement uncertainties.
-It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a $K$-mixture of Gaussians:
+It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a mixture of $K$ Gaussians:
 \BLedit{
@@
 \end{eqnarray}
-
-which is made analytically tractable with the Gaussian likelihood and prior \citep{bovy11}.
+which is made analytically tractable with the Gaussian likelihood and mixture-of-Gaussians prior \citep{bovy11}.
 
@@
 with $f(\varpi_{true})$ being a window function to insure $\varpi_{\true}$ is positive and to put $\varpi_{\true}$ on a similar grid for all the posteriors.
-
-\subsection{Cross Validation}
-For generating our prior, we use a mixture of 128 Gaussians.
-This number, 128, was chosen heuristically.
-It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product.
-We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.
 
@@
 
-\BLcomment{Isn't this repeating what was described in the previous paragraph?}
 We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust.
@@
 \subsection{prior}
-Using \xd\ and the MML method described in the methods section, we generate our prior shown in Figure \ref{fig:prior}.
+Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
 This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution.
@@
 We'll come back to more features in the data below, especially those seen in the posterior distribution.
-
-\BLcomment{Figure \ref{fig:prior} needs to be discussed here or in a separate subsection.
-I also find that it is the right place to discuss its features.
-I find it strange to discuss the posterior parallaxes without describing the prior.}
 
@@
   scope of this paper.
-\item[mixture of Gaussians] In everything in this project DWH
-\item[no physics] DWH
+\item[mixture of Gaussians] The \cmd\ model is a mixture of Gaussians,
+  and furthermore fixed at $K=128$ components. This setting was chosen
+  after some experimentation as being able to capture features but still
+  be easily optimized. We performed some experiments with
+  that suggest that larger $K$ would lead to better models, even in the
+  conservative sense of cross-validation score. Choosing an optimized or
+  optimal number for this mixture is a natural extension of this work.
+
+  Beyond this, the choice of the Gaussian form itself for the mixture is
+  questionable. The Gaussian has great properties, that we capitalize on
+  both at the prior-generation stage (with \xd), and at the photometric-parallax
+  stage (capitalizing on Gaussian product rules). We could move to other
+  forms, but they would be extremely expensive, computationally.
+\item[no physics] Although it is presented as an advantage of our
+  method that we never make any use of physical models, it is also
+  brings some significant disadvantages: For example, the DWH
 \end{description}

### 2017-06-02 paper/paper.tex
input delta var cdf plot
discussion

@@
 \renewcommand{\twocolumngrid}{} % trust me!
-
-\newcommand\BLedit[1]{{\color{blue}{#1}}}
-\newcommand\BLcomment[1]{{\color{cyan}{[BL: #1]}}}
 
@@
 It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a mixture of $K$ Gaussians:
-\BLedit{
 \begin{equation}
@@
 The relative amplitudes between the components are described by the scalars $\{ \mathrm{amp}_k \}$.
-}
+
 Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data. In equations, it finds the coefficients
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{dust.png}
-\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates.
-This shows both the footprint of our analysis as well as the dust values we used.
-Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
-\label{fig:dust}
-\end{figure}
-
 We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust.
@@
 After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution.
-\BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
+In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).
+We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
 We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
 The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.
-
-Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
-
 \begin{figure}
 \centering
-  \includegraphics[width=\textwidth]{prior_ngauss128.png}
-\caption{\BLedit{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
-The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
-The right panel shows the 1-$\sigma$ contours of the individual components.
-The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
-Thus, they appear non-Gaussian in the physical color--magnitude space.}}
-\label{fig:prior}
+  \includegraphics[width=\textwidth]{dust.png}
+\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates.
+This shows both the footprint of our analysis as well as the dust values we used.
+Each point represents a single star in our cross matched catalog.
+The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
+\label{fig:dust}
 \end{figure}
+
+LMA: Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
 
@@
 We'll come back to more features in the data below, especially those seen in the posterior distribution.
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{prior_ngauss128.png}
+\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
+The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
+The right panel shows the 1-$\sigma$ contours of the individual components.
+The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
+Thus, they appear non-Gaussian in the physical color--magnitude space.}
+\label{fig:prior}
+\end{figure}
 
@@
 
-Figure~\ref{fig:posteriorCMD} shows the distribution of the parallaxes derived with our \cmd\ prior.
+Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax inferences derived with our \cmd\ prior.
 This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
@@
 This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
-
-\BLcomment{Need to discuss the implications of that.}
-
-\subsection{Output Data}
-Using a data driven model of the observed \cmd\ we inferred a posterior belief of the true parallax of each star. Given our method, we call our posterior parallax the photometric parallax.
-DWH: Placeholder for the output format for our data products.
-RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
 
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{The expectation values and subsampled error bars.
+\caption{The expectation values of the posterior inferences (and error bars for a sub-sample).
 Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd.
@@
 \subsection{Distances to M67}
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{distancesM67.png}
-\caption{Proof of concept: parallax inference to the open cluster M67}
-\label{fig:m67}
-\end{figure}
-
-As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67.
-It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
+As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67.It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
 Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
@@
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{distancesM67.png}
+\caption{Proof of concept: parallax inference to the open cluster M67}
+\label{fig:m67}
+\end{figure}
 
@@
 \begin{description}
-\item[stationarity] (DWH needs to clean this paragraph)
-  We have assumed that (locally) the morphology of
+\item[stationarity] We have assumed that (locally) the morphology of
   the \cmd\ is independent of signal-to-noise; that is, we have assumed that
@@
   method that we never make any use of physical models, it is also
-  brings some significant disadvantages: For example, the DWH
+  brings some significant disadvantages: For example, the \cmd\ contains
+  features that are obvious products of stellar evolution, but doesn't in
+  itself inform the theory or models of stellar evolution, because our
+  mixture-of-Gaussian \cmd\ model is divorced from any theoretical ideas
+  about stars. For another, stellar evolution models are in fact very
+  predictive and successful; our model does not capitalize in any way on
+  those successes. A next-generation model might be designed to be data-driven
+  but nonetheless capitalize on physics-driven successes. One framework
+  for that would be to fit not the \cmd\ but rather deviations of the \cmd\ away
+  from a good physics-based model prediction.
 \end{description}

### 2017-06-02 paper/paper.tex
test

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{deltaCDF.png}
+\caption{Fractional Change in Variance: The cumulative distribution function of the fractional variance change when comparing our posterior to the \gaia\ catalogue.}
+\label{fig:deltaCDF}
+\end{figure}
+
 \subsection{Distances to M67}
-As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67.It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
-Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
+Previous work (LMA cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
 With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.

### 2017-06-02 paper/paper.tex
captioned cdf

--- 2017-06-02 paper/paper.tex
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{deltaCDF.png}
+\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.
+Negative numbers represent stars which have a posterior that is narrower than the \gaia\ catalogue, which is true for a large fraction of the stars.
+We've drawn two dashed lines.
+One represents the median star $(y=0.5)$ which has a fractional variation of $x=-0.3=ln(1/1.2^2)$, corresponding to an increase in precision by a factor of 1.2.
+Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.}
+\label{fig:deltaCDF}
+\end{figure}
+
 \subsection{Distances to M67}

### 2017-06-02 paper/paper.tex
abstract-tweaks

--- 2017-06-02 paper/paper.tex
@@
 Usually this prior represents some kind of belief about the Milky Way.
-However, there is multi-band photometry for a large fraction of the \gaia\ \tgas\ catalog from imaging surveys;
+However, there is multi-band photometry for a large fraction of the
+\gaia\ \tgas\ Catalog from imaging surveys;
 this imaging is incredibly informative about stellar distances.
-Here we use color information of $1.2$ million \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
+Here we use color information for $1.2$ million \tgas\ stars
+from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
 This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
 We use the Extreme Deconvolution (\xd) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
-The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
+The algorithm is run not in absolute-magnitude space but in a
+transformed space where the measurement uncertainty is close to Gaussian.
 The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
@@
 \tgas\ catalog entries by a median factor of 1.2, and 14 percent
-are more precise by a factor of two or more; they are also more precise than
+are more precise by a factor $>2$. They are also more precise than
 previous examples of Bayesian distance estimates using a spatial prior.
-We independently validate our distances by looking at members of a Milky Way star cluster M67, which is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly as a cluster in our posterior parallax estimates.
+We independently validate our parallax inferences distances by looking
+at members of Milky Way star cluster M67, which is not visible as a
+cluster in the \tgas\ parallax estimates, but appears clearly as a
+cluster in our posterior parallax estimates.
 All our results, including a posterior parallax and distance sampling for 1.4 million \tgas\ stars, are available in companion electronic tables.

### 2017-06-02 paper/paper.tex

--- 2017-06-02 paper/paper.tex
@@
 
-We assume (as \gaia\ suggests LMA REFERENCE HERE) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
+We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
 \begin{equation}
@@
 For each distance, we sample the median of the probabilistic dust model.
-We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (LMA REF TABLE SFD).
+We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from Table 6 of cite{schlafly11}.
 We then regenerate our prior using the lightly dust corrected photometry.
@@
 \subsection{Distances to M67}
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{distancesM67.png}
-\caption{Proof of concept: parallax inference to the open cluster M67}
-\label{fig:m67}
-\end{figure}
-
-As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67.
+As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
 It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
-Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+Previous work \citep{cbj15} has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
 With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.
@@
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{distancesM67.png}
+\caption{Proof of concept: parallax inference to the open cluster M67}
+\label{fig:m67}
+\end{figure}
 

### 2017-06-02 paper/paper.tex
removed typo
datarelease

--- 2017-06-02 paper/paper.tex
@@
 
-\subsection{prior}
+\subsection{XD-generated Prior}
 Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
@@
 
-%\begin{figure}
-%\centering
-%  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
-%\caption{Prior: the XD generated mixture of gaussians we use as our prior.
-To determine the number of gaussians we used cross validation, using 70\% of our %sample as a training set and 30\% as the test set.
-Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, %showing that we are no over fitting.
-Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the %features of the CMD.}
-%\label{fig:priorCompare}
-%\end{figure}
+\subsection{Data products}
+
+The method produces a mixture-of-Gaussian parallax posterior pdf for every
+star in the intersection of \tgas, \tmass, and the dust map.
+Associated with this paper is an electronic table containing our results, and everything
+that is necessary to use them responsibly.
+It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
+photometry.
+It contains posterior expectations and variances for the stellar parallax.
+It also contains the full posterior pdf for each star, in the form of a set of pdf values,
+evaluated on a common grid in parallax.
+Because all of the data are public, and all of the code used for this project is available
+publicly under an open-source license, power users can generate any of the figures in this
+paper or any other data outputs straightforwardly.
 

### 2017-06-02 paper/paper.tex

@@
 %@archiver{posteriorCMD.png}
-
-% # To-do list
-% - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
-% - deal with all DWH or LMA in the text.
-% - deal with everything in the google doc.
-% - are there any github issues?
 
@@
 
-We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust.
-With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way.
-The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference.
-Dust corrections are particularly non trivial for stars with poor parallax signal to noise.
-Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections.
-This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes.
-Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion).
-Instead, we correct for dust using a point estimate from the 3D dust map \citep{green15} sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$.
-In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
-Using this raw prior, we infer more precise parallaxes to all the stars in our sample.
-We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star.
-And repeat.
-This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone.
-In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance.
-We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference.
-For each distance, we sample the median of the probabilistic dust model.
-We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (LMA REF TABLE SFD).
-We then regenerate our prior using the lightly dust corrected photometry.
-The dust values we obtain in our final iteration, and which we apply to our photometry, are visualized in galactic coordinates in Figure \ref{fig:dust}.
-After applying a dust correction, we do not update the uncertainties, but keep them the same as before the dust correction, which is wrong and under representing our uncertainties but this leads, again, to a lighter deconvolution.
-In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).
-We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
-We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
-The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.
+We need to correct the photometry used in our method, both for the
+generation of the prior and the likelihood, for dust. With the
+emergence of 3D dust maps, it is now possible to estimate dust
+corrections to stars within the Milky Way. The challenge is getting a
+proper estimate of the distance to the star before inferring
+$\varpi_{\true}$, since we need a dust correction prior to doing our
+inference. Dust corrections are particularly non trivial for stars
+with poor parallax signal to noise. Stars with poor signal to noise
+have a likelihood that is consistent with infinite distance and can
+therefore have severe dust corrections. This is most obvious in the
+giant stars, which tend to have the lowest signal to noise, and can
+therefore have dust corrections $> 2$ magnitudes. Although the dust
+model is probabilistic, as well as our distance inference, to move
+forward, we break from a fully Bayesian framework, in which we might
+infer the distance and the dust simultaneously (more about this in the
+discussion). Instead, we correct for dust using a point estimate from
+the 3D dust map \citep{green15} sampled at a distance estimated from
+iteratively inferring and then sampling our posterior pdf over
+$\varpi_{\true}$. In more detail, to apply a dust correction to the
+\tmass\ photometry, we first generate our prior using the observed,
+attenuated photometry. Using this raw prior, we infer more precise
+parallaxes to all the stars in our sample. We then use this more
+precise parallax posterior to get a measurement of dust in the 3D dust
+map for each star. And repeat. This approach minimizes severe
+distances, and their associated severe dust corrections, which would
+occur if using the likelihoods alone. In detail, we take the $5\%$
+quantile of each distance posterior, so the closest bit of the
+posterior, and sample the 3D dust map at that distance. We take the
+$5\%$ quantile, as apposed to the median, for example, to do a light
+dust correction, because over correcting for dust can shoot the
+likelihood into an unphysical, different space of the \cmd\ and bias
+the parallax inference. For each distance, we sample the median of the
+probabilistic dust model. We apply each light dust correction to our
+\tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$
+= [0.709, 0.302] for bands [J, K] respectively, taken from Table 6 of
+cite{schlafly11}. We then regenerate our prior using the lightly dust
+corrected photometry. The dust values we obtain in our final
+iteration, and which we apply to our photometry, are visualized in
+galactic coordinates in Figure \ref{fig:dust}. After applying a dust
+correction, we do not update the uncertainties, but keep them the same
+as before the dust correction, which is wrong and under representing
+our uncertainties but this leads, again, to a lighter
+deconvolution. In other words, the features of our inferred
+  \cmd\ are conservatively broad (broader than if we over-estimated
+  the errors which would have lead to excessive deconvolution and
+  unrealistically narrow features). We could add the covariance of
+the dust to our likelihoods, but this is beyond the scope of this
+demonstration of concept. We iterate this process 10 times, however
+the dust values seem to converge after about 2-3 iterations. The
+iterative process allows for slight variations in distance and
+corresponding dust to become more consistent with the \cmd.
 \begin{figure}
@@
 
-LMA: Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
+Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
 
@@
   \includegraphics[width=\textwidth]{prior_ngauss128.png}
-\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by running XD on all of the data.
-The left panel shows a sampling of the prior,  with the blue lines showing the $1\sigma$ and $2\sigma$ contours.
-The right panel shows the 1-$\sigma$ contours of the individual components.
-The latter are Gaussian in the transformed magnitude space to make the XD inference tractable.
-Thus, they appear non-Gaussian in the physical color--magnitude space.}
+\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by
+  running XD on all of the data. The left panel shows a sampling of
+  the prior, with the blue lines showing the $1\sigma$ and $2\sigma$
+  contours. The right panel shows the 1-$\sigma$ contours of the
+  individual components. The latter are Gaussian in the transformed
+  magnitude space to make the XD inference tractable. Thus, they
+  appear non-Gaussian in the physical color--magnitude space.}
 \label{fig:prior}
@@
 This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
-
-\subsection{Data products}
-
-The method produces a mixture-of-Gaussian parallax posterior pdf for every
-star in the intersection of \tgas, \tmass, and the dust map.
-Associated with this paper is an electronic table containing our results, and everything
-that is necessary to use them responsibly.
-It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
-photometry.
-It contains posterior expectations and variances for the stellar parallax.
-It also contains the full posterior pdf for each star, in the form of a set of pdf values,
-evaluated on a common grid in parallax.
-Because all of the data are public, and all of the code used for this project is available
-publicly under an open-source license, power users can generate any of the figures in this
-paper or any other data outputs straightforwardly.
-
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{The expectation values of the posterior inferences (and error bars for a sub-sample).
-Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd.
-The visual of the prior is a sampling of the Gaussians, but this is the expectation value so this lies much closer to the center of the distribution.
-Similar to the toy model, where the pdf maximum of the noisier points lies in the center of the distribution.}
+\caption{The expectation values of the posterior inferences (and error
+  bars for a sub-sample). Notice that some parts of this \cmd\ are
+  tighter than the prior, or noise-deconvolved \cmd. The visual of the
+  prior is a sampling of the Gaussians, but this is the expectation
+  value so this lies much closer to the center of the
+  distribution. Similar to the toy model, where the pdf maximum of the
+  noisier points lies in the center of the distribution.}
 \label{fig:posteriorCMD}
@@
 \includegraphics[width=\textwidth]{deltaCDF.png}
-\caption{Fractional Change in Variance: The cumulative distribution function of the fractional variance change when comparing our posterior to the \gaia\ catalogue.}
+\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.
+Negative numbers represent stars which have a posterior that is narrower than the \gaia\ catalogue, which is true for a large fraction of the stars.
+We've drawn two dashed lines.
+One represents the median star $(y=0.5)$ which has a fractional variation of $x=-0.3=ln(1/1.2^2)$, corresponding to an increase in precision by a factor of 1.2.
+Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.}
 \label{fig:deltaCDF}
@@
 
+\subsection{Data products}
+
+The method produces a mixture-of-Gaussian parallax posterior pdf for every
+star in the intersection of \tgas, \tmass, and the dust map.
+Associated with this paper is an electronic table containing our results, and everything
+that is necessary to use them responsibly.
+It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
+photometry.
+It contains posterior expectations and variances for the stellar parallax.
+It also contains the full posterior pdf for each star, in the form of a set of pdf values,
+evaluated on a common grid in parallax.
+Because all of the data are public, and all of the code used for this project is available
+publicly under an open-source license, power users can generate any of the figures in this
+paper or any other data outputs straightforwardly.
+
 \subsection{Distances to M67}
-As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
-Previous work (LMA cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
+It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
+Previous work \citep{cbj15} has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
 With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.
@@
 \begin{description}
-\item[stationarity] We have assumed that (locally) the morphology of
+\item[stationarity]
+  We have assumed that (locally) the morphology of
   the \cmd\ is independent of signal-to-noise; that is, we have assumed that
@@
   method that we never make any use of physical models, it is also
-  brings some significant disadvantages: For example, the \cmd\ contains
-  features that are obvious products of stellar evolution, but doesn't in
-  itself inform the theory or models of stellar evolution, because our
-  mixture-of-Gaussian \cmd\ model is divorced from any theoretical ideas
-  about stars. For another, stellar evolution models are in fact very
-  predictive and successful; our model does not capitalize in any way on
-  those successes. A next-generation model might be designed to be data-driven
-  but nonetheless capitalize on physics-driven successes. One framework
-  for that would be to fit not the \cmd\ but rather deviations of the \cmd\ away
-  from a good physics-based model prediction.
+  brings some significant disadvantages: For example, the 
 \end{description}

### 2017-06-02 paper/paper.tex
fixed git issue

@@
   method that we never make any use of physical models, it is also
-  brings some significant disadvantages: For example, the 
+  brings some significant disadvantages: For example, the \cmd\ contains
+  features that are obvious products of stellar evolution, but doesn't in
+  itself inform the theory or models of stellar evolution, because our
+  mixture-of-Gaussian \cmd\ model is divorced from any theoretical ideas
+  about stars. For another, stellar evolution models are in fact very
+  predictive and successful; our model does not capitalize in any way on
+  those successes. A next-generation model might be designed to be data-driven
+  but nonetheless capitalize on physics-driven successes. One framework
+  for that would be to fit not the \cmd\ but rather deviations of the \cmd\ away
+  from a good physics-based model prediction.
 \end{description}

### 2017-06-04 paper/paper.tex
notes removal
fix apw name and affil
serious abstract changes

@@
 
-\author[0000-0003-0872-7098]{Adrian Price-Whelan}
-\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
+\author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
+\affil{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
 
@@
 \begin{abstract}\noindent % trust me
-Conversion of a noisy parallax measurement into a posterior
-belief over distances requires inference with a prior.
-Usually this prior represents some kind of belief about the Milky Way.
-However, there is multi-band photometry for a large fraction of the
-\gaia\ \tgas\ Catalog from imaging surveys;
-this imaging is incredibly informative about stellar distances.
-Here we use color information for $1.2$ million \tgas\ stars
-from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
-This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
-We use the Extreme Deconvolution (\xd) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
-The algorithm is run not in absolute-magnitude space but in a
-transformed space where the measurement uncertainty is close to Gaussian.
-The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
-Our posterior photometric parallax estimates are more precise than the
-\tgas\ catalog entries by a median factor of 1.2, and 14 percent
-are more precise by a factor $>2$. They are also more precise than
-previous examples of Bayesian distance estimates using a spatial prior.
-We independently validate our parallax inferences distances by looking
-at members of Milky Way star cluster M67, which is not visible as a
-cluster in the \tgas\ parallax estimates, but appears clearly as a
-cluster in our posterior parallax estimates.
-All our results, including a posterior parallax and distance sampling for 1.4 million \tgas\ stars, are available in companion electronic tables.
+Converting a noisy parallax measurement into a posterior
+belief over distance requires inference with a prior.
+Usually this prior represents beliefs about the stellar density distribution of
+the Milky Way.
+However, multi-band photometry exists for a large fraction of the
+\gaia\ \tgas\ Catalog and is incredibly informative about stellar distances.
+Here we use \tmass\ colors for $1.2$ million \tgas\ stars to build a
+noise-deconvolved empirical prior distribution for stars in color--magnitude
+space.
+This data-driven model contains no knowledge of stellar astrophysics or the
+Milky Way, but rather derives its precision from its generative model of noisy
+parallax measurements and an assumption of stationarity.
+We use the Extreme Deconvolution (\xd) algorithm---an approximation to a full
+hierarchical model of the true parallax and photometry of every star---to
+construct this prior.
+The \xd-optimized prior is used to infer a precise photometric parallax estimate
+and uncertainty (and full posterior) for every star.
+Our parallax estimates are more precise than the \tgas\ catalog entries by a
+median factor of 1.2 (14\% are more precise by a factor $>2$) and are more
+precise than previous Bayesian distance estimates that use spatial priors.
+We validate our parallax inferences using members of the Milky Way
+star cluster M67, which is not visible as a cluster in the \tgas\ parallax
+estimates, but appears as a cluster in our posterior parallax estimates.
+Our results, including a posterior parallax and distance sampling for 1.4
+million \tgas\ stars, are available in companion electronic tables.
 \end{abstract}
@@
   stars: distances
-  ---
-  stars: statistics
 }
@@
 
-%How this compares to other Bayesian attempts at inferring distances,
-%like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
-%assumptions, in fact! What form should this comparison be in?
 There are other options for prior information to include when inferring distances to stars.
@@
 It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
-%LMA: chat about bias?
 

### 2017-06-04 paper/paper.tex
tweaks to the introduction

@@
 
-The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar parallax measurements.
+The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
+stellar parallax measurements.
 Only a small fraction (but large number) of these
 measurements will be both precise and purely astrometric.
-\gaia\ will use astrometric parallax to determine the distances of the more precise
-stars, and then calibrate spectrophotometric models.
+\gaia\ will use astrometric parallaxes to determine the distances of the more
+precise stars, and then calibrate spectrophotometric models.
 These spectrophotometric models, in turn, along with \gaia's on-board
@@
 It is possible to build a stellar photometric model from the data themselves,
-because there are stars at all luminosities with useful parallax information,.
+because there are stars at all luminosities with useful parallax information.
 One challenge is that different stars are observed at different levels of parallax precision,
 so it requires relatively sophisticated technology to build this data-driven model
-using all the data available, fairly and responsibly.
-Luckily, we have this technology!
+using all of the data available, fairly and responsibly.
 
@@
 However, since the data set is large, the \xd\ approximation is not bad; it is
-sometimes known as the ``empirical Bayes'' method, and is well studied\footnote{For a start, try \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}.
+sometimes known as the ``empirical Bayes'' method, and is well studied.\footnote{For a primer on ``empirical Bayes'', see \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}
 
-It wouldn't be using the data twice---that is, it would be valid from a
-probabilistic perspective---if we performed a full hierarchical Bayesian
-inference.
+A probabilistically valid approach that wouldn't re-use the data is to perform
+a full hierarchical Bayesian inference.
 Exploration of that possibility, and its computational tractability,
@@
 It is promising that we will be able to obtain even better results
-in later \gaia\ data releases, and eventually put distances on $>10^9$
+using later \gaia\ data releases, and eventually infer distances for $>10^9$
 \gaia\ stars using photometric parallax methodologies but without any

### 2017-06-04 paper/paper.tex
changes to the why does this work section and switched Figure to \figurename

@@
 \caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior.
-The upper left panel visualizes the simplified toy distribution.
+The upper-left panel visualizes the simplified toy distribution.
 The red lines are the $1\sigma$ contours of the distribution, and the red points are 1024 samples from the distribution.
 These same $1\sigma$ contours are drawn in the bottom left panel.
-In the upper right and lower left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them.
+In the upper-right and lower-left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them.
 The error bars show the $1\sigma$ measurement uncertainties.
 The black points are 5 randomly chosen data to highlight the method.
-In the lower left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
+In the lower-left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
 The blue points represent the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
 The red points represent the actual true values.
-The lower right panel shows the posterior belief of the true value for all the points.
+The lower-right panel shows the posterior belief of the true value for all the points.
 They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior.
@@
 
-In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
-Understanding this method is challenging because we are using the data both for the prior and the likelihood in our inference.
-To ease understanding, here we will demonstrate this method with a simpler toy model.
-Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
-We will do this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$.
-Here the thickness of the toy distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
-The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
-Similarly, noisy samples from this distribution are shown in the upper right panel.
-Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
+In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (pdfs) from \gaia\ measurements.
+Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
+To ease understanding, here we demonstrate the generaly methodology using a
+simpler toy model with simulated data.
+Instead of inferring parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior pdfs of the value $Y$.
+We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
+Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
+The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel.
+Similarly, noisy samples from this distribution are shown in the upper-right panel.
+Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noise distribution.
 
@@
 \end{eqnarray}
-where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
-These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
-The upper left shows 1024 realizations of the true values, and the upper right, as well as lower left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
-Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
+where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$---which accounts for the intrinsic width---and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$---the noise distribution---so $\sigma_n$ is the measurement uncertainty.
+These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of \figurename~\ref{fig:toy}.
+The upper-left shows 1024 realizations of the true values, and the upper-right, as well as lower-left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points; we highlight 5 random points in black.
 
-To infer a pdf of the true y value for each measured y value we need a prior.
-For our prior we will use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy y data themselves.
-We learn the noise-deconvolved function using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
+To infer a posterior pdf of the true $y$ value for each measured $y$ value we need a prior.
+We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
+We learn the noise-deconvolved function using marginalized, maximum-likelihood estimation (MMLE), which marginalizes over the true data values and maximizes the likelihood of the data.
 Here our likelihood for each individual datum is
@@
 where we have marginalized over the true sample $y_{\true, n}$.
-The resulting likelihood is a Gaussian because the noise model is a Gaussian.
-The likelihood of the full data set is then the product of the individual likelihoods
+The resulting likelihood is a Gaussian because the noise model and our assumptions about the intrinsic width are both Gaussian.
+The likelihood of the full data set $\{y_n\}$ is then the product of the individual likelihoods
 \begin{equation}
-\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
+p(\{y_n\} \given m, b, t^2, \sigma_n^2) = \prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
 \label{eq:toyLikeFull}
@@
 
-This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In Figure \ref{fig:toy}, lower left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+The MMLE gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
+In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
@@
 Here is a good point to step back and acknowledge that our method is a
-bit strange. We are using the set of data to determine the
-noise-deconvolved distribution, which is an estimate of the true
-distribution, and then using that noise-deconvolved distribution to
-make inferences on the data. We are technically using the data
+bit strange. We are using the data to determine the
+noise-deconvolved distribution (an estimate of the true
+distribution), and then using that noise-deconvolved distribution to
+make inferences about the data. We are technically using the data
 twice! To be fully, technically correct, we should \emph{either} learn the true
 distribution while leaving one datum out, the one datum we want to
-infer the true y value for, or else go to full hierarchical Bayesian inference,
+infer the true y value for, or else perform a full hierarchical Bayesian inference,
 in which we treat the \cmd\ parameters probabilistically along with everything else.
-With large data sets, the MML or Empirical Bayes approximation is safe,
+With large data sets, the MMLE or Empirical Bayes approximation is safe,
 in some sense because no individual data point makes a large difference to
@@
 
-Returning the inference:
-We use Bayes theorem to turn the likelihood of the noisy data into a posterior believe over the true y value.
+Returning to the inference:
+We use Bayes' theorem to turn the likelihood of the noisy data into a posterior belief over the true $y$ value:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Since everything is Gaussian, this can be done analytically by multiplying Gaussians together.
-In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
-The red points are the associated true y values.
-In Figure \ref{fig:toy}, lower right panel, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
+In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs of the true values $y_{\true,n}$ for the 5 highlighted black points from the upper-right panel.
+The red points are the associated true $y$ values.
+In the lower-right panel of \figurename~\ref{fig:toy}, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance, $t^2$, and the measurement variance, $\sigma_n^2$.
-The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior.
+The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MMLE distribution, our prior.
 In contrast, more certain measurements remain closer to their measured values, $y_n$.
@@
 We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
-The full data set is visualized in the \cmd\ in Figure \ref{fig:data}.
+The full data set is visualized in the \cmd\ in \figurename~\ref{fig:data}.
 The left panel shows the point estimates of the color and absolute magnitude, using the point estimate of the parallax from \gaia, and the $1\sigma$ and $2\sigma$ contours of the distribution.
@@
 iteration, and which we apply to our photometry, are visualized in
-galactic coordinates in Figure \ref{fig:dust}. After applying a dust
+galactic coordinates in \figurename~\ref{fig:dust}. After applying a dust
 correction, we do not update the uncertainties, but keep them the same
@@
 \subsection{XD-generated Prior}
-Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
-This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution.
+Using \xd\ and the MMLE method described in the methods section, we generate the prior. This is shown in \figurename~\ref{fig:prior}.
+This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in \figurename~\ref{fig:data}, this is a tighter distribution.
 The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from.
@@
 With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
-Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors.
+\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
 The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
@@
 It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
-A comparison of the posteriors using the two different priors, our \cmd\ prior and the EDSD, is shown in Figure \ref{fig:comparePrior}.
+A comparison of the posteriors using the two different priors, our \cmd\ prior and the EDSD, is shown in \figurename~\ref{fig:comparePrior}.
 Here the parallax posterior pdf expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
@@
 The denoised expectation values of the parallax projected into the \cmd\ show interest features.
-We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
+We visualize this \cmd\ again in \figurename~\ref{fig:wtf} pointing out some interesting features.
 \begin{figure}

### 2017-06-04 paper/paper.tex
assumptions section

@@
 
-To infer the true parallax to a star using the noise-deconvolved \cmd\ as
+To infer the true parallax for a star using the noise-deconvolved \cmd\ as
 a prior, we make the following assumptions. The method we present is correct
@@
   case, if a portion of the \cmd\ has only one (statistically distinct)
-  exemplar, for example, a white dwarf, inference for these cases will be biased.
+  exemplar---for example, a white dwarf---inference for these cases will be biased.
 \item[noise model] We assume that the \tgas\ parallax uncertainties dominate the

### 2017-06-04 paper/paper.tex
methods

@@
 \subsection{Setting up the Inference}
- \emph{Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff.
-It's confusing right now jumping between inference and prior building !!}
 
@@
 
-Equation~\ref{eq:bayes} is too simplistic to describe this procedure.
-Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
+Equation~\ref{eq:bayes} does not fully specify this procedure.
+Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
 Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
@@
 \begin{equation}
-y_0 = \varpi\,10^{0.2J^c_n} = 10^{0.2\,M^c_{J,n} + 2}
+y_0 = \varpi\,10^{\frac{1}{5}\,J^c_n} = 10^{\frac{1}{5}\,M^c_{J,n} + 2}
 \label{eq:transform}
 \end{equation}
-The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain.
-Gaussian uncertainties are required to generate our prior using \xd, but more about this and our dust corrections below.
+where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
+The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain (Gaussian uncertainties are required to use \xd, see below).
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
@@
 
-We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
+We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood for the $n$th star is
 \begin{equation}
@@
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
-0 & (\sigma_{\varpi}\,10^{0.2\,J_n^C})^2
+0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^C}\right)^2
 \end{bmatrix}
 \]
-where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
-Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude.
-So we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+where $\mathcal{N}(\boldsymbol{\mu}, \mathbf{\Sigma})$ represents a normal distribution with mean vector $\boldsymbol{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
+Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
+We take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
 
-We can now write the multidimensional analogous to Equation \ref{eq:bayes},
+We can now write the multidimensional posterior expression analogous to Equation~\ref{eq:bayes},
 \begin{equation}
-p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
+p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) \propto p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
 \end{equation}
@@
 Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves.
-We will use XD, which will take advantage of the Gaussian likelihood.
+For this, we use XD, which takes advantage of the Gaussian likelihood.
 
@@
 
-To generate our prior $p(\mathbf{Y}_n)$, we fit our data in the slightly transformed \cmd\ space using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are Gaussian, heterogeneous measurement uncertainties.
-It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a mixture of $K$ Gaussians:
+To generate the empirical prior $p(\mathbf{Y}_n)$, we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with heterogeneous, Gaussian noise variances $\{ V_n\}$.
+\xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
 \begin{equation}
-	p\bigl(\ \mathbf{Y} \given  \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K \mathrm{amp}_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
+	p\bigl(\ \mathbf{Y} \given  \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
 \end{equation}
 where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
-The relative amplitudes between the components are described by the scalars $\{ \mathrm{amp}_k \}$.
+The relative amplitudes between the components are described by the scalars $\{ A_k \}$.
 
-Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data. In equations, it finds the coefficients
+Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
+Assuming $\alpha = \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K$, it finds the coefficients
 \begin{eqnarray}
-\alpha &=& \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \nonumber\\
-       &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
-       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
+\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
+             &=& \argmax_{\alpha} \, \prod_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
 \label{eq:xdmml}

### 2017-06-05 paper/paper.tex
data and results

@@
 
-We use stars crossmatched in \tgas\ and \tmass\, that also lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
-We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
+We use stars crossmatched in \tgas\ and \tmass\, that also lie within the
+observing footprint of \panstarrs\ to access the \cite{green15} 3D dust model.
+We require that the photometry have real values, and nonzero, real errors, and
+remove a small selection of \tmass\ stars that have zero color and zero $J$-band
+magnitude.
 The full data set is visualized in the \cmd\ in \figurename~\ref{fig:data}.
-The left panel shows the point estimates of the color and absolute magnitude, using the point estimate of the parallax from \gaia, and the $1\sigma$ and $2\sigma$ contours of the distribution.
+The left panel shows the point estimates of the color and absolute magnitude,
+using the point estimate of the parallax from \gaia, and the $1\sigma$ and
+$2\sigma$ contours of the distribution.
 The right panel shows a subset of the data with it's associated error bars.
-The uncertainties in the colors are fairly well behaved, but the large uncertainties in the absolute magnitude are due to the large uncertainties in the parallax measurements.
+The uncertainties in the colors are fairly well behaved, but the large
+uncertainties in some absolute magnitudes are due to the large uncertainties in
+the parallax measurements.
 
@@
 
-We need to correct the photometry used in our method, both for the
-generation of the prior and the likelihood, for dust. With the
-emergence of 3D dust maps, it is now possible to estimate dust
+For the generation of the prior and the evaluation of the likelihood, we need to
+correct the \tmass\ photometry for dust extinction.
+With the emergence of 3D dust maps, it is now possible to estimate dust
 corrections to stars within the Milky Way. The challenge is getting a
@@
 model is probabilistic, as well as our distance inference, to move
-forward, we break from a fully Bayesian framework, in which we might
+forward, we break from a fully Bayesian framework in which we might
 infer the distance and the dust simultaneously (more about this in the
@@
 precise parallax posterior to get a measurement of dust in the 3D dust
-map for each star. And repeat. This approach minimizes severe
+map for each star. We then iterate this procedure. This approach minimizes severe
 distances, and their associated severe dust corrections, which would
-occur if using the likelihoods alone. In detail, we take the $5\%$
-quantile of each distance posterior, so the closest bit of the
-posterior, and sample the 3D dust map at that distance. We take the
-$5\%$ quantile, as apposed to the median, for example, to do a light
-dust correction, because over correcting for dust can shoot the
-likelihood into an unphysical, different space of the \cmd\ and bias
-the parallax inference. For each distance, we sample the median of the
-probabilistic dust model. We apply each light dust correction to our
-\tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$
-= [0.709, 0.302] for bands [J, K] respectively, taken from Table 6 of
-cite{schlafly11}. We then regenerate our prior using the lightly dust
-corrected photometry. The dust values we obtain in our final
+occur if using the likelihoods alone.
+
+In detail, we take the $5\%$
+quantile of each distance posterior (the closest part of the
+distance posterior) and query the 3D dust map at that distance.
+We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
+correction because over-correcting for dust can push the peak of the likelihood
+function into an unphysical, different space of the \cmd\ and thus bias the
+parallax inference.
+For each distance, we sample the median of the
+probabilistic dust model. We apply each dust correction to our
+\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
+= [0.709, 0.302]$ for bands $[J, K]$ respectively (see Table 6 in
+\citealt{schlafly11}).
+We then regenerate our prior using the dust-corrected photometry.
+The dust values we obtain in our final
 iteration, and which we apply to our photometry, are visualized in
-galactic coordinates in \figurename~\ref{fig:dust}. After applying a dust
-correction, we do not update the uncertainties, but keep them the same
-as before the dust correction, which is wrong and under representing
-our uncertainties but this leads, again, to a lighter
-deconvolution. In other words, the features of our inferred
+galactic coordinates in \figurename~\ref{fig:dust}.
+After applying the dust correction, we do not update the uncertainties, which is
+technically wrong and under-represents the true uncertainties; as mentioned
+previously, this leads to a less severe deconvolution.
+In other words, the features of our inferred
   \cmd\ are conservatively broad (broader than if we over-estimated
@@
   \includegraphics[width=\textwidth]{dust.png}
-\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates.
-This shows both the footprint of our analysis as well as the dust values we used.
+\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates.
+This shows both the footprint of our analysis as well as the adopted extinction values.
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
+The large missing areas are due to the \panstarrs\ footprint ($>-30~{\rm deg}$ in declination)}
 \label{fig:dust}
@@
 
-Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
-
-\subsection{XD-generated Prior}
-Using \xd\ and the MMLE method described in the methods section, we generate the prior. This is shown in \figurename~\ref{fig:prior}.
-This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in \figurename~\ref{fig:data}, this is a tighter distribution.
-The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from.
-The main sequence and red giant branch are much tighter, as well as the red clump.
+\subsection{The empirical prior}
+Using \xd\ and the MMLE method described in the methods section, we generate the
+prior, shown in \figurename~\ref{fig:prior}.
+The left panel shows 1.3 million samples from the prior distribution, and the
+right panel shows $1\sigma$ contours for each component in the underlying
+Gaussian mixture model.
+Compared to the raw data in \figurename~\ref{fig:data}, the deconvolved
+color--magnitude diagram for the \gaia\ + \tmass\ stars is tighter, as expected.
 We'll come back to more features in the data below, especially those seen in the posterior distribution.
@@
   magnitude space to make the XD inference tractable. Thus, they
-  appear non-Gaussian in the physical color--magnitude space.}
+  appear as slightly deformed ellipses in color--magnitude space.}
 \label{fig:prior}
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
-Previous work \citep{cbj15} has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
-With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.
-So it is a great test case for comparing parallax inferences with or without a prior.
-We made a window of 1 square degree on the sky with the cluster M67 lying in the middle.
-The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky.
+M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of 3--$5~{\rm Gyr}$.
+Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
+With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
+M67 is therefore a great test case for comparing parallax inferences with or without a prior.
+
+We select sources from our catalog within a $1~{\rm deg}^2$ square window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
 With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
@@
 The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
-The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all.
-The posterior pdf, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods.
+The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
+The posterior pdfs, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster (vertical lines) for far more stars.
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.

### 2017-06-05 paper/paper.tex

@@
 We have shown that it is possible to obtain photometric parallaxes for
-distant stars in the \gaia\ \tgas\ data without any use of physical stellar
-models, nor models of the Milky Way.
+distant stars in the \gaia\ \tgas\ catalog without any use of physical stellar
+models, nor stellar density models of the Milky Way.
 We used the geometric parallaxes to calibrate a photometric
@@
 an Empirical Bayesian maximum-marginalized-likelihood estimator, accounts for
-heteroskedastic noise, we were able to build our photometric parallax
+heteroskedastic noise, we are able to build our photometric parallax
 model---which is a model of the \cmd---using
 all the data, not just the data with the highest signal-to-noise.
-So, the model for the \cmd\ we built is representative for
+So, the generated model for the \cmd\ is representative for
 our subset of the \tgas\ Catalog.
@@
   \includegraphics[width=\textwidth]{comparePrior.png}
-\caption{Comparison with common prior: A visualization of the posterior pdf for a subsection of TGAS stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
-On the left is the posterior pdf using our \cmd\ prior, and on the right is the posterior pdf when using the exponentially declining spatial density, a common prior used when inferring parallaxes.
-The scale of the exponentially declining spatial density prior is 1.35 kpc, the optimal value found in \cite{astraatmadja16a}.
+\caption{Comparison with a common spatial prior: A visualization of the posterior pdf for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
+Left panel shows the posterior pdf using our \cmd\ prior, right panel shows the posterior pdfs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
+The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}.
 Using the \cmd\ as a prior generates a posterior pdf that has smaller variance.}
@@
 One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
-It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
+It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so that the posteriors are also very smooth.
 It is a fairly weak prior.
 It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
-A comparison of the posteriors using the two different priors, our \cmd\ prior and the EDSD, is shown in \figurename~\ref{fig:comparePrior}.
+\figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
 Here the parallax posterior pdf expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
@@
 \subsection{What's That Feature?}
-The denoised expectation values of the parallax projected into the \cmd\ show interest features.
-We visualize this \cmd\ again in \figurename~\ref{fig:wtf} pointing out some interesting features.
+The denoised expectation values of the parallax projected into the \cmd\ show many interesting and familiar features; we visualize this \cmd\ again in \figurename~\ref{fig:wtf} and highlight some of these features.
 \begin{figure}
@@
 is much wider. This may be an age variation on the main sequence mixed
-with a malmquist bias of seeing a much larger volume of upper main
+with a Malmquist bias of seeing a much larger volume of upper main
 sequence stars. There is a plausible binary sequence of stars with
@@
 noticeable and also surprisingly narrow, possibly a reflection of the
-star formation history of the local volume also mixed with a malmquist
+star formation history of the local volume also mixed with a Malmquist
 bias.
@@
   distances. So the red giant branch is a population of
-  relatively lower signal to noise objects compared with the main
+  relatively lower signal-to-noise objects compared with the main
   sequence. This does not violate the stationarity assumption in
-  detail, because the low and high signal to noise objects have to be
+  detail, because the low and high signal-to-noise objects have to be
   drawn from the same distribution within some feature or patch of the
@@
 \item[selection] The biggest limitation of our method for generating
-  the \cmd\ from the data is that we made no use of any kind of
-  \tgas\ selection function or completeness estimate or
+  the \cmd\ from the data is that we have not used a
+  \tgas\ selection function, completeness estimate, or
   inverse-selection-volume corrections.  This model is a model of the
@@
   after some experimentation as being able to capture features but still
-  be easily optimized. We performed some experiments with
+  be easily optimized. We performed some experiments
   that suggest that larger $K$ would lead to better models, even in the

### 2017-06-05 paper/paper.tex
panstarrs reference

@@
 \newcommand{\gaia}{\project{Gaia}}
-\newcommand{\panstarrs}{\project{Pan\acronym{STARRS}}}
+\newcommand{\panstarrs}{\project{Pan\acronym{STARRS1}}}
+\newcommand{\ps1}{\project{\acronym{PS1}}}
 \newcommand{\xd}{\acronym{XD}}
@@
 these data were generated by the methodology of \citealt{michalik15}), that match to the
-\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of
-the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
+\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\ps1)
+to make use of the \cite{green15} dust map. We use this data to make a model of the
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
-\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \ps1-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
@@
 
-We use stars crossmatched in \tgas\ and \tmass\, that also lie within the
-observing footprint of \panstarrs\ to access the \cite{green15} 3D dust model.
+We use stars crossmatched in \tgas\ and \tmass\.
+The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
+\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}
+We also required that the stars lie within the
+observing footprint of \ps1\ to access the \cite{green15} 3D dust model.
 We require that the photometry have real values, and nonzero, real errors, and
@@
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the \panstarrs\ footprint ($>-30~{\rm deg}$ in declination)}
+The large missing areas are due to the \ps1\ footprint ($>-30~{\rm deg}$ in declination)}
 \label{fig:dust}

### 2017-06-05 paper/paper.tex
fixed bug

@@
 \newcommand{\tmass}{\project{\acronym{2MASS}}}
+\newcommand{\psone}{\project{\acronym{PS1}}}
 \newcommand{\gaia}{\project{Gaia}}
 \newcommand{\panstarrs}{\project{Pan\acronym{STARRS1}}}
-\newcommand{\ps1}{\project{\acronym{PS1}}}
 \newcommand{\xd}{\acronym{XD}}
@@
 these data were generated by the methodology of \citealt{michalik15}), that match to the
-\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\ps1)
+\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\psone; \citealt{ps1})
 to make use of the \cite{green15} dust map. We use this data to make a model of the
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
-\item[dust] We treat the \ps1-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
@@
 The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
-\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}
+\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
 We also required that the stars lie within the
-observing footprint of \ps1\ to access the \cite{green15} 3D dust model.
+observing footprint of \psone\ to access the \cite{green15} 3D dust model.
 We require that the photometry have real values, and nonzero, real errors, and
@@
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the \ps1\ footprint ($>-30~{\rm deg}$ in declination)}
+The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
 \label{fig:dust}

### 2017-06-05 paper/paper.tex
software

--- 2017-06-05 paper/paper.tex
@@
 
-\author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
-\affil{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
+\author[0000-0003-0872-7098]{Adrian Price-Whelan}
+\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
 
@@
 \begin{abstract}\noindent % trust me
-Converting a noisy parallax measurement into a posterior
-belief over distance requires inference with a prior.
-Usually this prior represents beliefs about the stellar density distribution of
-the Milky Way.
-However, multi-band photometry exists for a large fraction of the
-\gaia\ \tgas\ Catalog and is incredibly informative about stellar distances.
-Here we use \tmass\ colors for $1.2$ million \tgas\ stars to build a
-noise-deconvolved empirical prior distribution for stars in color--magnitude
-space.
-This data-driven model contains no knowledge of stellar astrophysics or the
-Milky Way, but rather derives its precision from its generative model of noisy
-parallax measurements and an assumption of stationarity.
-We use the Extreme Deconvolution (\xd) algorithm---an approximation to a full
-hierarchical model of the true parallax and photometry of every star---to
-construct this prior.
-The \xd-optimized prior is used to infer a precise photometric parallax estimate
-and uncertainty (and full posterior) for every star.
-Our parallax estimates are more precise than the \tgas\ catalog entries by a
-median factor of 1.2 (14\% are more precise by a factor $>2$) and are more
-precise than previous Bayesian distance estimates that use spatial priors.
-We validate our parallax inferences using members of the Milky Way
-star cluster M67, which is not visible as a cluster in the \tgas\ parallax
-estimates, but appears as a cluster in our posterior parallax estimates.
-Our results, including a posterior parallax and distance sampling for 1.4
-million \tgas\ stars, are available in companion electronic tables.
+Conversion of a noisy parallax measurement into a posterior
+belief over distances requires inference with a prior.
+Usually this prior represents some kind of belief about the Milky Way.
+However, there is multi-band photometry for a large fraction of the
+\gaia\ \tgas\ Catalog from imaging surveys;
+this imaging is incredibly informative about stellar distances.
+Here we use color information for $1.2$ million \tgas\ stars
+from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
+This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
+We use the Extreme Deconvolution (\xd) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
+The algorithm is run not in absolute-magnitude space but in a
+transformed space where the measurement uncertainty is close to Gaussian.
+The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
+Our posterior photometric parallax estimates are more precise than the
+\tgas\ catalog entries by a median factor of 1.2, and 14 percent
+are more precise by a factor $>2$. They are also more precise than
+previous examples of Bayesian distance estimates using a spatial prior.
+We independently validate our parallax inferences distances by looking
+at members of Milky Way star cluster M67, which is not visible as a
+cluster in the \tgas\ parallax estimates, but appears clearly as a
+cluster in our posterior parallax estimates.
+All our results, including a posterior parallax and distance sampling for 1.4 million \tgas\ stars, are available in companion electronic tables.
 \end{abstract}
@@
   stars: distances
+  ---
+  stars: statistics
 }
@@
 
-The \gaia\ Mission (\citealt{gaia16}) will soon deliver more than a billion
-stellar parallax measurements.
+The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar parallax measurements.
 Only a small fraction (but large number) of these
 measurements will be both precise and purely astrometric.
-\gaia\ will use astrometric parallaxes to determine the distances of the more
-precise stars, and then calibrate spectrophotometric models.
+\gaia\ will use astrometric parallax to determine the distances of the more precise
+stars, and then calibrate spectrophotometric models.
 These spectrophotometric models, in turn, along with \gaia's on-board
@@
 It is possible to build a stellar photometric model from the data themselves,
-because there are stars at all luminosities with useful parallax information.
+because there are stars at all luminosities with useful parallax information,.
 One challenge is that different stars are observed at different levels of parallax precision,
 so it requires relatively sophisticated technology to build this data-driven model
-using all of the data available, fairly and responsibly.
+using all the data available, fairly and responsibly.
+Luckily, we have this technology!
 
@@
 However, since the data set is large, the \xd\ approximation is not bad; it is
-sometimes known as the ``empirical Bayes'' method, and is well studied.\footnote{For a primer on ``empirical Bayes'', see \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}
+sometimes known as the ``empirical Bayes'' method, and is well studied\footnote{For a start, try \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}.
 
-A probabilistically valid approach that wouldn't re-use the data is to perform
-a full hierarchical Bayesian inference.
+It wouldn't be using the data twice---that is, it would be valid from a
+probabilistic perspective---if we performed a full hierarchical Bayesian
+inference.
 Exploration of that possibility, and its computational tractability,
@@
 It is promising that we will be able to obtain even better results
-using later \gaia\ data releases, and eventually infer distances for $>10^9$
+in later \gaia\ data releases, and eventually put distances on $>10^9$
 \gaia\ stars using photometric parallax methodologies but without any
@@
 \caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior.
-The upper-left panel visualizes the simplified toy distribution.
+The upper left panel visualizes the simplified toy distribution.
 The red lines are the $1\sigma$ contours of the distribution, and the red points are 1024 samples from the distribution.
 These same $1\sigma$ contours are drawn in the bottom left panel.
-In the upper-right and lower-left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them.
+In the upper right and lower left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them.
 The error bars show the $1\sigma$ measurement uncertainties.
 The black points are 5 randomly chosen data to highlight the method.
-In the lower-left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
+In the lower left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
 The blue points represent the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
 The red points represent the actual true values.
-The lower-right panel shows the posterior belief of the true value for all the points.
+The lower right panel shows the posterior belief of the true value for all the points.
 They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior.
@@
 
-In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (pdfs) from \gaia\ measurements.
-Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
-To ease understanding, here we demonstrate the generaly methodology using a
-simpler toy model with simulated data.
-Instead of inferring parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior pdfs of the value $Y$.
-We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
-Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
-The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel.
-Similarly, noisy samples from this distribution are shown in the upper-right panel.
-Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noise distribution.
+In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
+Understanding this method is challenging because we are using the data both for the prior and the likelihood in our inference.
+To ease understanding, here we will demonstrate this method with a simpler toy model.
+Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
+We will do this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$.
+Here the thickness of the toy distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
+The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
+Similarly, noisy samples from this distribution are shown in the upper right panel.
+Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
 
@@
 \end{eqnarray}
-where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$---which accounts for the intrinsic width---and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$---the noise distribution---so $\sigma_n$ is the measurement uncertainty.
-These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of \figurename~\ref{fig:toy}.
-The upper-left shows 1024 realizations of the true values, and the upper-right, as well as lower-left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points; we highlight 5 random points in black.
+where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
+These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
+The upper left shows 1024 realizations of the true values, and the upper right, as well as lower left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
+Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
 
-To infer a posterior pdf of the true $y$ value for each measured $y$ value we need a prior.
-We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
-We learn the noise-deconvolved function using marginalized, maximum-likelihood estimation (MMLE), which marginalizes over the true data values and maximizes the likelihood of the data.
+To infer a pdf of the true y value for each measured y value we need a prior.
+For our prior we will use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy y data themselves.
+We learn the noise-deconvolved function using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
 Here our likelihood for each individual datum is
@@
 where we have marginalized over the true sample $y_{\true, n}$.
-The resulting likelihood is a Gaussian because the noise model and our assumptions about the intrinsic width are both Gaussian.
-The likelihood of the full data set $\{y_n\}$ is then the product of the individual likelihoods
+The resulting likelihood is a Gaussian because the noise model is a Gaussian.
+The likelihood of the full data set is then the product of the individual likelihoods
 \begin{equation}
-p(\{y_n\} \given m, b, t^2, \sigma_n^2) = \prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
+\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
 \label{eq:toyLikeFull}
@@
 
-The MMLE gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
+In Figure \ref{fig:toy}, lower left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
@@
 Here is a good point to step back and acknowledge that our method is a
-bit strange. We are using the data to determine the
-noise-deconvolved distribution (an estimate of the true
-distribution), and then using that noise-deconvolved distribution to
-make inferences about the data. We are technically using the data
+bit strange. We are using the set of data to determine the
+noise-deconvolved distribution, which is an estimate of the true
+distribution, and then using that noise-deconvolved distribution to
+make inferences on the data. We are technically using the data
 twice! To be fully, technically correct, we should \emph{either} learn the true
 distribution while leaving one datum out, the one datum we want to
-infer the true y value for, or else perform a full hierarchical Bayesian inference,
+infer the true y value for, or else go to full hierarchical Bayesian inference,
 in which we treat the \cmd\ parameters probabilistically along with everything else.
-With large data sets, the MMLE or Empirical Bayes approximation is safe,
+With large data sets, the MML or Empirical Bayes approximation is safe,
 in some sense because no individual data point makes a large difference to
@@
 
-Returning to the inference:
-We use Bayes' theorem to turn the likelihood of the noisy data into a posterior belief over the true $y$ value:
+Returning the inference:
+We use Bayes theorem to turn the likelihood of the noisy data into a posterior believe over the true y value.
 \begin{eqnarray}
@@
 \end{eqnarray}
-In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs of the true values $y_{\true,n}$ for the 5 highlighted black points from the upper-right panel.
-The red points are the associated true $y$ values.
-In the lower-right panel of \figurename~\ref{fig:toy}, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
+Since everything is Gaussian, this can be done analytically by multiplying Gaussians together.
+In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
+The red points are the associated true y values.
+In Figure \ref{fig:toy}, lower right panel, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance, $t^2$, and the measurement variance, $\sigma_n^2$.
-The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MMLE distribution, our prior.
+The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior.
 In contrast, more certain measurements remain closer to their measured values, $y_n$.
@@
 
-To infer the true parallax for a star using the noise-deconvolved \cmd\ as
+To infer the true parallax to a star using the noise-deconvolved \cmd\ as
 a prior, we make the following assumptions. The method we present is correct
@@
   case, if a portion of the \cmd\ has only one (statistically distinct)
-  exemplar---for example, a white dwarf---inference for these cases will be biased.
+  exemplar, for example, a white dwarf, inference for these cases will be biased.
 \item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
@@
 \subsection{Setting up the Inference}
+ \emph{Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff.
+It's confusing right now jumping between inference and prior building !!}
 
@@
 
-Equation~\ref{eq:bayes} does not fully specify this procedure.
-Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
+Equation~\ref{eq:bayes} is too simplistic to describe this procedure.
+Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
 Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
@@
 \begin{equation}
-y_0 = \varpi\,10^{\frac{1}{5}\,J^c_n} = 10^{\frac{1}{5}\,M^c_{J,n} + 2}
+y_0 = \varpi\,10^{0.2J^c_n} = 10^{0.2\,M^c_{J,n} + 2}
 \label{eq:transform}
 \end{equation}
-where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
-The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain (Gaussian uncertainties are required to use \xd, see below).
+The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain.
+Gaussian uncertainties are required to generate our prior using \xd, but more about this and our dust corrections below.
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
@@
 
-We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood for the $n$th star is
+We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the $n$th star is
 \begin{equation}
@@
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
-0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^C}\right)^2
+0 & (\sigma_{\varpi}\,10^{0.2\,J_n^C})^2
 \end{bmatrix}
 \]
-where $\mathcal{N}(\boldsymbol{\mu}, \mathbf{\Sigma})$ represents a normal distribution with mean vector $\boldsymbol{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
-Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
-We take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
+Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude.
+So we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
 
-We can now write the multidimensional posterior expression analogous to Equation~\ref{eq:bayes},
+We can now write the multidimensional analogous to Equation \ref{eq:bayes},
 \begin{equation}
-p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) \propto p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
+p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
 \end{equation}
@@
 Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves.
-For this, we use XD, which takes advantage of the Gaussian likelihood.
+We will use XD, which will take advantage of the Gaussian likelihood.
 
@@
 
-To generate the empirical prior $p(\mathbf{Y}_n)$, we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with heterogeneous, Gaussian noise variances $\{ V_n\}$.
-\xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
+To generate our prior $p(\mathbf{Y}_n)$, we fit our data in the slightly transformed \cmd\ space using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with its uncertainties $\{ V_n\}$, which are Gaussian, heterogeneous measurement uncertainties.
+It generates an estimate of the underlying distribution from which the uncertain data were drawn from, modeled as a mixture of $K$ Gaussians:
 \begin{equation}
-	p\bigl(\ \mathbf{Y} \given  \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
+	p\bigl(\ \mathbf{Y} \given  \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K \mathrm{amp}_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
 \end{equation}
 where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
-The relative amplitudes between the components are described by the scalars $\{ A_k \}$.
+The relative amplitudes between the components are described by the scalars $\{ \mathrm{amp}_k \}$.
 
-Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
-Assuming $\alpha = \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K$, it finds the coefficients
+Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data. In equations, it finds the coefficients
 \begin{eqnarray}
-\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
-             &=& \argmax_{\alpha} \, \prod_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
+\alpha &=& \left\{\mathrm{amp}_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \nonumber\\
+       &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
+       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
 \label{eq:xdmml}
@@
 
-We use stars crossmatched in \tgas\ and \tmass\, that also lie within the
-observing footprint of \panstarrs\ to access the \cite{green15} 3D dust model.
-We require that the photometry have real values, and nonzero, real errors, and
-remove a small selection of \tmass\ stars that have zero color and zero $J$-band
-magnitude.
-The full data set is visualized in the \cmd\ in \figurename~\ref{fig:data}.
-The left panel shows the point estimates of the color and absolute magnitude,
-using the point estimate of the parallax from \gaia, and the $1\sigma$ and
-$2\sigma$ contours of the distribution.
+We use stars crossmatched in \tgas\ and \tmass\, that also lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
+We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
+The full data set is visualized in the \cmd\ in Figure \ref{fig:data}.
+The left panel shows the point estimates of the color and absolute magnitude, using the point estimate of the parallax from \gaia, and the $1\sigma$ and $2\sigma$ contours of the distribution.
 The right panel shows a subset of the data with it's associated error bars.
-The uncertainties in the colors are fairly well behaved, but the large
-uncertainties in some absolute magnitudes are due to the large uncertainties in
-the parallax measurements.
+The uncertainties in the colors are fairly well behaved, but the large uncertainties in the absolute magnitude are due to the large uncertainties in the parallax measurements.
 
@@
 
-For the generation of the prior and the evaluation of the likelihood, we need to
-correct the \tmass\ photometry for dust extinction.
-With the emergence of 3D dust maps, it is now possible to estimate dust
+We need to correct the photometry used in our method, both for the
+generation of the prior and the likelihood, for dust. With the
+emergence of 3D dust maps, it is now possible to estimate dust
 corrections to stars within the Milky Way. The challenge is getting a
@@
 model is probabilistic, as well as our distance inference, to move
-forward, we break from a fully Bayesian framework in which we might
+forward, we break from a fully Bayesian framework, in which we might
 infer the distance and the dust simultaneously (more about this in the
@@
 precise parallax posterior to get a measurement of dust in the 3D dust
-map for each star. We then iterate this procedure. This approach minimizes severe
+map for each star. And repeat. This approach minimizes severe
 distances, and their associated severe dust corrections, which would
-occur if using the likelihoods alone.
-
-In detail, we take the $5\%$
-quantile of each distance posterior (the closest part of the
-distance posterior) and query the 3D dust map at that distance.
-We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
-correction because over-correcting for dust can push the peak of the likelihood
-function into an unphysical, different space of the \cmd\ and thus bias the
-parallax inference.
-For each distance, we sample the median of the
-probabilistic dust model. We apply each dust correction to our
-\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
-= [0.709, 0.302]$ for bands $[J, K]$ respectively (see Table 6 in
-\citealt{schlafly11}).
-We then regenerate our prior using the dust-corrected photometry.
-The dust values we obtain in our final
+occur if using the likelihoods alone. In detail, we take the $5\%$
+quantile of each distance posterior, so the closest bit of the
+posterior, and sample the 3D dust map at that distance. We take the
+$5\%$ quantile, as apposed to the median, for example, to do a light
+dust correction, because over correcting for dust can shoot the
+likelihood into an unphysical, different space of the \cmd\ and bias
+the parallax inference. For each distance, we sample the median of the
+probabilistic dust model. We apply each light dust correction to our
+\tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$
+= [0.709, 0.302] for bands [J, K] respectively, taken from Table 6 of
+cite{schlafly11}. We then regenerate our prior using the lightly dust
+corrected photometry. The dust values we obtain in our final
 iteration, and which we apply to our photometry, are visualized in
-galactic coordinates in \figurename~\ref{fig:dust}.
-After applying the dust correction, we do not update the uncertainties, which is
-technically wrong and under-represents the true uncertainties; as mentioned
-previously, this leads to a less severe deconvolution.
-In other words, the features of our inferred
+galactic coordinates in Figure \ref{fig:dust}. After applying a dust
+correction, we do not update the uncertainties, but keep them the same
+as before the dust correction, which is wrong and under representing
+our uncertainties but this leads, again, to a lighter
+deconvolution. In other words, the features of our inferred
   \cmd\ are conservatively broad (broader than if we over-estimated
@@
   \includegraphics[width=\textwidth]{dust.png}
-\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates.
-This shows both the footprint of our analysis as well as the adopted extinction values.
+\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates.
+This shows both the footprint of our analysis as well as the dust values we used.
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the \panstarrs\ footprint ($>-30~{\rm deg}$ in declination)}
+The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
 \label{fig:dust}
@@
 
-\subsection{The empirical prior}
-Using \xd\ and the MMLE method described in the methods section, we generate the
-prior, shown in \figurename~\ref{fig:prior}.
-The left panel shows 1.3 million samples from the prior distribution, and the
-right panel shows $1\sigma$ contours for each component in the underlying
-Gaussian mixture model.
-Compared to the raw data in \figurename~\ref{fig:data}, the deconvolved
-color--magnitude diagram for the \gaia\ + \tmass\ stars is tighter, as expected.
+Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
+
+\subsection{XD-generated Prior}
+Using \xd\ and the MML method described in the methods section, we generate the prior. This is shown in Figure \ref{fig:prior}.
+This is the \gaia\ + \tmass\ data set deconvolved with it's noise, so compared with the raw data in Figure \ref{fig:data}, this is a tighter distribution.
+The left panel shows 1.3M samples from the prior distribution, and the right panel shows the underlying gaussian mixture model these samples are drawn from.
+The main sequence and red giant branch are much tighter, as well as the red clump.
 We'll come back to more features in the data below, especially those seen in the posterior distribution.
@@
   magnitude space to make the XD inference tractable. Thus, they
-  appear as slightly deformed ellipses in color--magnitude space.}
+  appear non-Gaussian in the physical color--magnitude space.}
 \label{fig:prior}
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of 3--$5~{\rm Gyr}$.
-Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
-With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
-M67 is therefore a great test case for comparing parallax inferences with or without a prior.
-
-We select sources from our catalog within a $1~{\rm deg}^2$ square window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
+It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
+Previous work \citep{cbj15} has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.
+So it is a great test case for comparing parallax inferences with or without a prior.
+We made a window of 1 square degree on the sky with the cluster M67 lying in the middle.
+The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky.
 With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
-\figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
+Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors.
 The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
-The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
-The posterior pdfs, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster (vertical lines) for far more stars.
+The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all.
+The posterior pdf, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods.
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
@@
 We have shown that it is possible to obtain photometric parallaxes for
-distant stars in the \gaia\ \tgas\ catalog without any use of physical stellar
-models, nor stellar density models of the Milky Way.
+distant stars in the \gaia\ \tgas\ data without any use of physical stellar
+models, nor models of the Milky Way.
 We used the geometric parallaxes to calibrate a photometric
@@
 an Empirical Bayesian maximum-marginalized-likelihood estimator, accounts for
-heteroskedastic noise, we are able to build our photometric parallax
+heteroskedastic noise, we were able to build our photometric parallax
 model---which is a model of the \cmd---using
 all the data, not just the data with the highest signal-to-noise.
-So, the generated model for the \cmd\ is representative for
+So, the model for the \cmd\ we built is representative for
 our subset of the \tgas\ Catalog.
@@
   \includegraphics[width=\textwidth]{comparePrior.png}
-\caption{Comparison with a common spatial prior: A visualization of the posterior pdf for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
-Left panel shows the posterior pdf using our \cmd\ prior, right panel shows the posterior pdfs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
-The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}.
+\caption{Comparison with common prior: A visualization of the posterior pdf for a subsection of TGAS stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
+On the left is the posterior pdf using our \cmd\ prior, and on the right is the posterior pdf when using the exponentially declining spatial density, a common prior used when inferring parallaxes.
+The scale of the exponentially declining spatial density prior is 1.35 kpc, the optimal value found in \cite{astraatmadja16a}.
 Using the \cmd\ as a prior generates a posterior pdf that has smaller variance.}
@@
 One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
-It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so that the posteriors are also very smooth.
+It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
 It is a fairly weak prior.
 It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
-\figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
+A comparison of the posteriors using the two different priors, our \cmd\ prior and the EDSD, is shown in Figure \ref{fig:comparePrior}.
 Here the parallax posterior pdf expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
@@
 \subsection{What's That Feature?}
-The denoised expectation values of the parallax projected into the \cmd\ show many interesting and familiar features; we visualize this \cmd\ again in \figurename~\ref{fig:wtf} and highlight some of these features.
+The denoised expectation values of the parallax projected into the \cmd\ show interest features.
+We visualize this \cmd\ again in Figure \ref{fig:wtf} pointing out some interesting features.
 \begin{figure}
@@
 is much wider. This may be an age variation on the main sequence mixed
-with a Malmquist bias of seeing a much larger volume of upper main
+with a malmquist bias of seeing a much larger volume of upper main
 sequence stars. There is a plausible binary sequence of stars with
@@
 noticeable and also surprisingly narrow, possibly a reflection of the
-star formation history of the local volume also mixed with a Malmquist
+star formation history of the local volume also mixed with a malmquist
 bias.
@@
   distances. So the red giant branch is a population of
-  relatively lower signal-to-noise objects compared with the main
+  relatively lower signal to noise objects compared with the main
   sequence. This does not violate the stationarity assumption in
-  detail, because the low and high signal-to-noise objects have to be
+  detail, because the low and high signal to noise objects have to be
   drawn from the same distribution within some feature or patch of the
@@
 \item[selection] The biggest limitation of our method for generating
-  the \cmd\ from the data is that we have not used a
-  \tgas\ selection function, completeness estimate, or
+  the \cmd\ from the data is that we made no use of any kind of
+  \tgas\ selection function or completeness estimate or
   inverse-selection-volume corrections.  This model is a model of the
@@
   after some experimentation as being able to capture features but still
-  be easily optimized. We performed some experiments
+  be easily optimized. We performed some experiments with
   that suggest that larger $K$ would lead to better models, even in the
@@
 Center for Computational Astrophysics for comments and input.
-This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
+This project was developed in part at the 2016 \acronym{NYC} \gaia\ Sprint, hosted
 by the Center for Computational Astrophysics at the Simons Foundation
@@
 This work has made use of data from the European Space Agency (\acronym{ESA})
-mission \gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the Gaia
-Data Processing and Analysis Consortium\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}} (\acronym{DPAC}). Funding for the
+mission \gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the \gaia\ Data
+Processing and Analysis Consortium\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}} (\acronym{DPAC}). Funding for the
 \acronym{DPAC} has been provided by national institutions, in particular the
-institutions participating in the Gaia Multilateral Agreement.
+institutions participating in the \gaia\ Multilateral Agreement.
 
@@
 
-%% \appendix
-%% \section{Appendix material}
+\software{
+The code used in this project is available from
+\url{https://github.com/andersdot/photoParallax}.
+This research utilized the following open-source \textsl{Python} packages:
+    \textsl{Astropy} (\citealt{Astropy-Collaboration:2013}),
+    \textsl{matplotlib} (\citealt{Hunter:2007}), and
+    \textsl{numpy} (\citealt{Van-der-Walt:2011}).
+This work additionally used the \gaia\ science archive
+(\url{https://gea.esac.esa.int/archive/}).
+}
 
@@
 
-\clearpage
-
 \end{document}
-
-%%
-%% End of file `sample.tex'.

### 2017-06-05 paper/paper.tex

--- 2017-06-05 paper/paper.tex
@@
 \newcommand{\gaia}{\project{Gaia}}
-\newcommand{\panstarrs}{\project{Pan\acronym{STARRS1}}}
-\newcommand{\ps1}{\project{\acronym{PS1}}}
+\newcommand{\panstarrs}{\project{Pan\acronym{STARRS}}}
 \newcommand{\xd}{\acronym{XD}}
@@
 these data were generated by the methodology of \citealt{michalik15}), that match to the
-\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\ps1)
-to make use of the \cite{green15} dust map. We use this data to make a model of the
+\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of
+the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
-\item[dust] We treat the \ps1-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
@@
 
-We use stars crossmatched in \tgas\ and \tmass\.
-The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
-\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}
-We also required that the stars lie within the
-observing footprint of \ps1\ to access the \cite{green15} 3D dust model.
+We use stars crossmatched in \tgas\ and \tmass\, that also lie within the
+observing footprint of \panstarrs\ to access the \cite{green15} 3D dust model.
 We require that the photometry have real values, and nonzero, real errors, and
@@
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the \ps1\ footprint ($>-30~{\rm deg}$ in declination)}
+The large missing areas are due to the \panstarrs\ footprint ($>-30~{\rm deg}$ in declination)}
 \label{fig:dust}

### 2017-06-05 paper/paper.tex
desnark
alpha hat

@@
 \newcommand{\tmass}{\project{\acronym{2MASS}}}
+\newcommand{\psone}{\project{\acronym{PS1}}}
 \newcommand{\gaia}{\project{Gaia}}
-\newcommand{\panstarrs}{\project{Pan\acronym{STARRS}}}
+\newcommand{\panstarrs}{\project{Pan\acronym{STARRS1}}}
 \newcommand{\xd}{\acronym{XD}}
@@
 these data were generated by the methodology of \citealt{michalik15}), that match to the
-\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of
-the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
+\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of \panstarrs\ (\psone; \citealt{ps1})
+to make use of the \cite{green15} dust map. We use this data to make a model of the
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
-\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
@@
 \begin{eqnarray}
-p(\mathbf{Y}_0 \given \mathbf{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
+p(\mathbf{Y}_0 \given \mathbf{y}, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{alpha}) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
@@
 
-We use stars crossmatched in \tgas\ and \tmass\, that also lie within the
-observing footprint of \panstarrs\ to access the \cite{green15} 3D dust model.
+We use stars crossmatched in \tgas\ and \tmass\.
+The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
+\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
+We also required that the stars lie within the
+observing footprint of \psone\ to access the \cite{green15} 3D dust model.
 We require that the photometry have real values, and nonzero, real errors, and
@@
 Each point represents a single star in our cross matched catalog.
-The large missing areas are due to the \panstarrs\ footprint ($>-30~{\rm deg}$ in declination)}
+The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
 \label{fig:dust}
@@
 \section{Conclusion}
-If you want to read a conclusion, re-read the abstract!
+We forgo an explicit conclusion; we attempt to summarize the conclusions in the Abstract above.
 

### 2017-06-05 paper/paper.tex
math stuff

--- 2017-06-05 paper/paper.tex
@@
 \newcommand{\given}{\,|\,}
+\newcommand{\dd}{\mathrm{d}}
 \newcommand{\true}{\mathrm{true}}
+\renewcommand{\vec}[1]{\boldsymbol{{#1}}}
+\newcommand{\truth}{\vec{Y}}
+\newcommand{\data}{\vec{y}}
+\newcommand{\mean}{\vec{\mu}}
+\newcommand{\mat}[1]{{\mathrm{{#1}}}}
+\newcommand{\var}{\mat{V}}
+\newcommand{\covar}{\mat{C}}
 
@@
 \begin{eqnarray}
-p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \mathrm{d}y_{\true, n} \\
+p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \dd y_{\true, n} \\
 &=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2) \quad,
@@
 Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
-Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$.
-The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\mathbf{y}$.
+Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
+The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\data$.
 
 We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
-Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude
+Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
 \begin{equation}
@@
 \begin{aligned}
-\mathrm{true \, vector} \quad &\mathbf{Y}_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
-\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
+\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
+\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
 \mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
@@
 \begin{equation}
-p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
+p(\data_n \given \truth_n, I_n) = \mathcal{N}(\data_n \given \truth_n, \covar_n)
 \end{equation}
@@
 \[
-V_n = \begin{bmatrix}
+\covar_n = \begin{bmatrix}
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
@@
 \]
-where $\mathcal{N}(\boldsymbol{\mu}, \mathbf{\Sigma})$ represents a normal distribution with mean vector $\boldsymbol{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
+where $\mathcal{N}(\mean, \mat{\Sigma})$ represents a normal distribution with
+mean vector $\mean$ and covariance matrix $\mat{\Sigma}$.
 Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
@@
 \begin{equation}
-p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) \propto p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
+p(\truth_n \given \data_n, I_n) \propto p(\data_n \given \truth_n, I_n) \, p(\truth_n) \label{eq:posterior}
 \end{equation}
-where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
+where $p(\data_n \given \truth_n, I_n)$ is the likelihood of the data vector, and $p(\truth_n)$ is our prior.
 Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves.
@@
 
-To generate the empirical prior $p(\mathbf{Y}_n)$, we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with heterogeneous, Gaussian noise variances $\{ V_n\}$.
-\xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
+To generate the empirical prior $p(\truth_n)$,
+we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd,
+which deconvolves the data $\{ \truth_n \}$ with heterogeneous,
+Gaussian noise variances $\{ \covar_n\}$. \xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
 \begin{equation}
-	p\bigl(\ \mathbf{Y} \given  \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
+	p\bigl(\ \truth \given  \left\{A_k, \mean_k, \var_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\truth \given \mean_k, \var_k )
 \end{equation}
-where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
-The relative amplitudes between the components are described by the scalars $\{ A_k \}$.
+where $\mean_k$ and $\var_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
+The relative amplitudes of the components are given by the scalars $\{ A_k \}$.
 
-Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
-Assuming $\alpha = \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K$, it finds the coefficients
+Given the data $\{ \truth_n, \covar_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
+Assuming $\alpha = \left\{A_k, \mean_k, \var_k\right\}_{k=1}^K$, it finds the coefficients
 \begin{eqnarray}
-\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
-             &=& \argmax_{\alpha} \, \prod_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
+\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\data_n \given \alpha) \nonumber\\
+             &=& \argmax_{\alpha} \, \prod_n \, \int p(\data_n \given \truth_n) \, p(\truth_n \given \alpha)\, \dd\truth_n
 \label{eq:xdmml}
@@
 
-To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.
+To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\truth \given \alpha)$, using Equation \ref{eq:posterior}.
 We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
@@
 \begin{eqnarray}
-p(\mathbf{Y}_0 \given \mathbf{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
+p(\truth_0 \given \data, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}

### 2017-06-05 paper/paper.tex

@@
 \begin{eqnarray}
-p(\truth_0 \given \data, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
+p(\truth_0 \given \data, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{\alpha}) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}

### 2017-06-05 paper/paper.tex
more math tweaks

@@
 \newcommand{\true}{\mathrm{true}}
+\newcommand{\corr}{\mathrm{c}}
 \renewcommand{\vec}[1]{\boldsymbol{{#1}}}
@@
 \newcommand{\mean}{\vec{\mu}}
-\newcommand{\mat}[1]{{\mathrm{{#1}}}}
+\newcommand{\mat}[1]{{\mathbf{{#1}}}}
 \newcommand{\var}{\mat{V}}
@@
 Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
-Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
-The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\data$.
+Our true, latent values are the $(J-K)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
+The observed data is the dust corrected color $(J-K)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.
 
@@
 \begin{equation}
-y_0 = \varpi\,10^{\frac{1}{5}\,J^c_n} = 10^{\frac{1}{5}\,M^c_{J,n} + 2}
+y_0 = \varpi\,10^{\frac{1}{5}\,J^{\corr}_n} = 10^{\frac{1}{5}\,M^{\corr}_{J,n} + 2}
 \label{eq:transform}
@@
 \begin{aligned}
-\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
-\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
-\mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
-\mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
+\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K)^{\true}_n] \\
+\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J- K)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
+\mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
+\mathrm{and \; color} \quad &(J - K)^{\corr}_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
 \mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
@@
 
-We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood for the $n$th star is
+We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^{\corr}_n$ color likelihood. So our full likelihood for the $n$th star is
 \begin{equation}
@@
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
-0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^C}\right)^2
+0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^{\corr}}\right)^2
 \end{bmatrix}
@@
 Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
-We take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+We take $J_n^{\corr}$ as a point estimate and only propagate the parallax uncertainty.
 
@@
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
-              1 \quad \mathrm{if} \quad -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
+              1 \quad \mathrm{if} \quad -2 < \mathrm{log}_{10} \varpi_{\true} < 2\\
               0 \quad \mathrm{otherwise}
@@
 \end{eqnarray}
-with $f(\varpi_{true})$ being a window function to insure $\varpi_{\true}$ is positive and to put $\varpi_{\true}$ on a similar grid for all the posteriors.
+with $f(\varpi_{\true})$ being a window function to insure $\varpi_{\true}$ is positive and to put $\varpi_{\true}$ on a similar grid for all the posteriors.
 
@@
 The median fractional change is 1.2.
-Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^C$ of about 0.6.
+Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^{\corr}$ of about 0.6.
 The right panel shows the posterior point estimates on the CMD colored by fraction change in variance.

### 2017-06-05 paper/paper.tex
figure 5 caption

--- 2017-06-05 paper/paper.tex
@@
 \newcommand{\given}{\,|\,}
-\newcommand{\dd}{\mathrm{d}}
 \newcommand{\true}{\mathrm{true}}
-\renewcommand{\vec}[1]{\boldsymbol{{#1}}}
-\newcommand{\truth}{\vec{Y}}
-\newcommand{\data}{\vec{y}}
-\newcommand{\mean}{\vec{\mu}}
-\newcommand{\mat}[1]{{\mathrm{{#1}}}}
-\newcommand{\var}{\mat{V}}
-\newcommand{\covar}{\mat{C}}
 
@@
 \begin{eqnarray}
-p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \dd y_{\true, n} \\
+p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \mathrm{d}y_{\true, n} \\
 &=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2) \quad,
@@
 Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
-Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
-The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\data$.
+Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$.
+The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\mathbf{y}$.
 
 We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
-Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
+Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude
 \begin{equation}
@@
 \begin{aligned}
-\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
-\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
+\mathrm{true \, vector} \quad &\mathbf{Y}_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
+\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
 \mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
@@
 \begin{equation}
-p(\data_n \given \truth_n, I_n) = \mathcal{N}(\data_n \given \truth_n, \covar_n)
+p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
 \end{equation}
@@
 \[
-\covar_n = \begin{bmatrix}
+V_n = \begin{bmatrix}
 \sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
@@
 \]
-where $\mathcal{N}(\mean, \mat{\Sigma})$ represents a normal distribution with
-mean vector $\mean$ and covariance matrix $\mat{\Sigma}$.
+where $\mathcal{N}(\boldsymbol{\mu}, \mathbf{\Sigma})$ represents a normal distribution with mean vector $\boldsymbol{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
 Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
@@
 \begin{equation}
-p(\truth_n \given \data_n, I_n) \propto p(\data_n \given \truth_n, I_n) \, p(\truth_n) \label{eq:posterior}
+p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) \propto p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n) \label{eq:posterior}
 \end{equation}
-where $p(\data_n \given \truth_n, I_n)$ is the likelihood of the data vector, and $p(\truth_n)$ is our prior.
+where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
 Before we describe how this can be turned into a posterior belief on the true parallax, we turn our attention to the prior, which will be generated from the data themselves.
@@
 
-To generate the empirical prior $p(\truth_n)$,
-we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd,
-which deconvolves the data $\{ \truth_n \}$ with heterogeneous,
-Gaussian noise variances $\{ \covar_n\}$. \xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
+To generate the empirical prior $p(\mathbf{Y}_n)$, we fit our data in the transformed \cmd\ space (see previous sub-section) using \xd, which deconvolves the data $\{ \mathbf{Y}_n \}$ with heterogeneous, Gaussian noise variances $\{ V_n\}$.
+\xd\ generates an estimate of the underlying distribution from which the uncertain data were drawn, modeled as a mixture of $K$ Gaussians:
 \begin{equation}
-	p\bigl(\ \truth \given  \left\{A_k, \mean_k, \var_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\truth \given \mean_k, \var_k )
+	p\bigl(\ \mathbf{Y} \given  \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K \bigr) = \sum_{k=1}^K A_k \ \mathcal{N}(\mathbf{Y} \given \mathbf{\mu}_k, V_k )
 \end{equation}
-where $\mean_k$ and $\var_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
-The relative amplitudes of the components are given by the scalars $\{ A_k \}$.
+where $\mathbf{\mu}_k$ and $V_k$ are a two dimensional vector and matrix, respectively, which correspond to the location and covariance of the $k$th component (a Gaussian in our transformed \cmd\ space).
+The relative amplitudes between the components are described by the scalars $\{ A_k \}$.
 
-Given the data $\{ \truth_n, \covar_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
-Assuming $\alpha = \left\{A_k, \mean_k, \var_k\right\}_{k=1}^K$, it finds the coefficients
+Given the data $\{ \mathbf{Y}_n, V_n\}$, \xd\ maximizes the marginalized likelihood of the data, marginalizing over the true data.
+Assuming $\alpha = \left\{A_k, \mathbf{\mu}_k, V_k\right\}_{k=1}^K$, it finds the coefficients
 \begin{eqnarray}
-\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\data_n \given \alpha) \nonumber\\
-             &=& \argmax_{\alpha} \, \prod_n \, \int p(\data_n \given \truth_n) \, p(\truth_n \given \alpha)\, \dd\truth_n
+\hat{\alpha} &=& \argmax_{\alpha} \, \prod_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
+             &=& \argmax_{\alpha} \, \prod_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
 \label{eq:xdmml}
@@
 
-To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\truth \given \alpha)$, using Equation \ref{eq:posterior}.
+To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\mathbf{Y} \given \alpha)$, using Equation \ref{eq:posterior}.
 We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
@@
 \begin{eqnarray}
-p(\truth_0 \given \data, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{\alpha}) \, f(\varpi_{\true}) \\
+p(\mathbf{Y}_0 \given \mathbf{y}, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{alpha}) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
@@
   \includegraphics[width=\textwidth]{posterior.png}
-\caption{5 example posteriors showing the priors and likelihoods as well}
+\caption{Five example posterior pdfs over parallax using our method.
+The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
+We chose example points that evenly spanned the color space, but within each color bin, we chose a point at random.
+The remaining panels show the posterior pdf in blue, the prior pdf in green, and likelihood of the data in black.
+Most stars have posterior pdfs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
+Examples 1-4 are not very affected by the prior (green), where good data is good data.
+Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior pdf keeps the star on the main sequence.
+Example 5 shows a significantly more precise posterior pdf compared with the likelihood, and looks very similar to the prior pdf.
+Our prior tends to have the largest affect on stars in the red giant branch.}
 \label{fig:posterior}

### 2017-06-05 paper/paper.tex
figure 5 caption fixes

@@
 Most stars have posterior pdfs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1-4 are not very affected by the prior (green), where good data is good data.
-Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior pdf keeps the star on the main sequence.
+Examples 1-4 are not very effected by the prior (green), where good data is good data.
+Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior pdf, transformed into an absolute magnitude, remains on the main sequence.
 Example 5 shows a significantly more precise posterior pdf compared with the likelihood, and looks very similar to the prior pdf.
-Our prior tends to have the largest affect on stars in the red giant branch.}
+It also exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
 \label{fig:posterior}

### 2017-06-05 paper/paper.tex
beauty

@@
 space.
-This data-driven model contains no knowledge of stellar astrophysics or the
-Milky Way, but rather derives its precision from its generative model of noisy
-parallax measurements and an assumption of stationarity.
+This model contains no knowledge of stellar astrophysics or the
+Milky Way, but is precise (and beautiful) because it is a generative model fit to many noisy
+parallax measurements under an assumption of stationarity.
 We use the Extreme Deconvolution (\xd) algorithm---an approximation to a full

### 2017-06-05 paper/paper.tex
caption 6

--- 2017-06-05 paper/paper.tex
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{The expectation values of the posterior inferences (and error
-  bars for a sub-sample). Notice that some parts of this \cmd\ are
-  tighter than the prior, or noise-deconvolved \cmd. The visual of the
-  prior is a sampling of the Gaussians, but this is the expectation
-  value so this lies much closer to the center of the
-  distribution. Similar to the toy model, where the pdf maximum of the
-  noisier points lies in the center of the distribution.}
+\caption{The distribution of expectation values of the parallax posterior pdfs transformed into an absolute magnitude vs color, the \cmd, in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution.
+The right panel shows a subsample of data, plotting the expectation value and $1\sigma$ uncertainty of posterior pdf projected onto the \cmd.
+Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}
+The visual of the prior is a sampling of the Gaussians, but this is the expectation
+  value, so it lies much closer to the center of the
+  distribution. Similar to the toy model, where the maximum of the
+  posterior pdf of the noisy points lies closer to the center of the true distribution than the true values.}
 \label{fig:posteriorCMD}

### 2017-06-06 paper/paper.tex
dust section and posterior

@@
 
-In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (pdfs) from \gaia\ measurements.
+In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (PDFs) from \gaia\ measurements.
 Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
@@
 simpler toy model with simulated data.
-Instead of inferring parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior pdfs of the value $Y$.
+Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $Y$.
 We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
@@
 
-To infer a posterior pdf of the true $y$ value for each measured $y$ value we need a prior.
+To infer a posterior PDF of the true $y$ value for each measured $y$ value we need a prior.
 We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
@@
 not typical values or typical draws from the posterior; they show the
-modes (and, with error bars, the variances) of the posterior pdfs.
+modes (and, with error bars, the variances) of the posterior PDFs.
 When an original data point has a large variance, the posterior ends
@@
 \end{equation}
-where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
+where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
 parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
@@
 Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
-Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
-Our true, latent values are the $(J-K)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
-The observed data is the dust corrected color $(J-K)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.
+Specifically, we use the $J$ and $K_s$ band photometry from \tmass, as well as the parallax from \gaia.
+Our true, latent values are the $(J-K_s)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
+The observed data is the dust corrected color $(J-K_s)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.
 
@@
 \begin{aligned}
-\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K)^{\true}_n] \\
-\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J- K)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
+\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K_s)^{\true}_n] \\
+\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J-K_s)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
 \mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
-\mathrm{and \; color} \quad &(J - K)^{\corr}_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
+\mathrm{and \; color} \quad &(J-K_s)^{\corr}_n = J_n - K_{s,n} - Q_{JK}\,E(B-V)_n \\
 \mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
@@
 
-We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^{\corr}_n$ color likelihood. So our full likelihood for the $n$th star is
+We assume that the parallax likelihood is Gaussian, as well as the $(J-K_s)^{\corr}_n$ color likelihood. So our full likelihood for the $n$th star is
 \begin{equation}
@@
   \includegraphics[width=\textwidth]{data.png}
-\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K$ color, the $J$ band apparent magnitude, and the \gaia\ parallax.
+\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K_s$ color, the $J$ band apparent magnitude, and the \gaia\ parallax.
 The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution.
@@
 inference. Dust corrections are particularly non trivial for stars
-with poor parallax signal to noise. Stars with poor signal to noise
-have a likelihood that is consistent with infinite distance and can
+with poor parallax signal to noise; these stars have
+a likelihood that is consistent with infinite distance and can
 therefore have severe dust corrections. This is most obvious in the
@@
 discussion). Instead, we correct for dust using a point estimate from
-the 3D dust map \citep{green15} sampled at a distance estimated from
-iteratively inferring and then sampling our posterior pdf over
-$\varpi_{\true}$. In more detail, to apply a dust correction to the
-\tmass\ photometry, we first generate our prior using the observed,
+the 3D dust map \citep{green15}. We determine this dust correction by
+iteratively inferring, and then sampling, our posterior PDF over
+$\varpi_{\true}$. We first generate our prior using the observed,
 attenuated photometry. Using this raw prior, we infer more precise
-parallaxes to all the stars in our sample. We then use this more
+parallaxes to all the stars in our sample. We use this more
 precise parallax posterior to get a measurement of dust in the 3D dust
-map for each star. We then iterate this procedure. This approach minimizes severe
-distances, and their associated severe dust corrections, which would
-occur if using the likelihoods alone.
+map for each star. We then iterate this process 10 times, however
+the dust values seem to converge after about 2-3 iterations. The
+iterative process allows for slight variations in distance and
+corresponding dust to become more consistent with the \cmd.
 
@@
 We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
-correction because over-correcting for dust can push the peak of the likelihood
+correction because over-correcting for dust can push the likelihood
 function into an unphysical, different space of the \cmd\ and thus bias the
-parallax inference.
-For each distance, we sample the median of the
-probabilistic dust model. We apply each dust correction to our
+parallax inference. At the 3D position of each star, we sample the \texttt{mode=median} value of the probabilistic dust map.
+We apply each dust correction to our
 \tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
-= [0.709, 0.302]$ for bands $[J, K]$ respectively (see Table 6 in
+= [0.709, 0.302]$ for bands $[J, K_s]$ respectively (see Table 6 in
 \citealt{schlafly11}).
-We then regenerate our prior using the dust-corrected photometry.
 The dust values we obtain in our final
-iteration, and which we apply to our photometry, are visualized in
+iteration, and which we apply to the photometry in our inference, are visualized in
 galactic coordinates in \figurename~\ref{fig:dust}.
@@
 the dust to our likelihoods, but this is beyond the scope of this
-demonstration of concept. We iterate this process 10 times, however
-the dust values seem to converge after about 2-3 iterations. The
-iterative process allows for slight variations in distance and
-corresponding dust to become more consistent with the \cmd.
+demonstration of concept.
 \begin{figure}
@@
 This shows both the footprint of our analysis as well as the adopted extinction values.
-Each point represents a single star in our cross matched catalog.
+Each point represents a single star in our cross-matched catalog.
 The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
@@
 
-Figure~\ref{fig:posterior} shows the end result of our inference for a few randomly-chosen objects: the posterior pdf on the true parallaxes.
-We also show the likelihood and the prior.
-In all cases, the prior information brought by the \cmd\ is significant, and leads to a compelling decrease in the parallax uncertainty.
+Figure~\ref{fig:posterior} shows the various probability distributions of our inference for a few randomly-chosen objects: the likelihood (black), prior (green), and posterior PDF (blue) on the true parallaxes.
+For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.
+Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$.
+This is visualized by Examples 1-4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
+However, $14\%$ of our parllax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$.
+This larger change in precision is visualized in Example 5, which shows a significantly narrower posterior PDF compared to the likelihood.
 The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact strongly depend on the location of each object in \cmd\ space.
 
-Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax inferences derived with our \cmd\ prior.
+Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax PDFs derived with our \cmd\ prior.
 This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
@@
   \includegraphics[width=\textwidth]{posterior.png}
-\caption{Five example posterior pdfs over parallax using our method.
+\caption{Five example posterior PDFs over parallax using our method.
 The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
 We chose example points that evenly spanned the color space, but within each color bin, we chose a point at random.
-The remaining panels show the posterior pdf in blue, the prior pdf in green, and likelihood of the data in black.
-Most stars have posterior pdfs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
+The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
+Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
 Examples 1-4 are not very effected by the prior (green), where good data is good data.
-Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior pdf, transformed into an absolute magnitude, remains on the main sequence.
-Example 5 shows a significantly more precise posterior pdf compared with the likelihood, and looks very similar to the prior pdf.
+Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior PDF, transformed into an absolute magnitude, remains on the main sequence.
+Example 5 shows a significantly more precise posterior PDF compared with the likelihood, and looks very similar to the prior PDF.
 It also exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{The distribution of expectation values of the parallax posterior pdfs transformed into an absolute magnitude vs color, the \cmd, in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution.
-The right panel shows a subsample of data, plotting the expectation value and $1\sigma$ uncertainty of posterior pdf projected onto the \cmd.
+\caption{The distribution of expectation values of the parallax posterior PDFs transformed into an absolute magnitude vs color, the \cmd, of the full data set is shown in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution shown in black.
+The right panel shows a subsample of data set, plotting the expectation value and $1\sigma$ uncertainty of posterior PDF projected onto the \cmd.
 Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}
@@
   distribution. Similar to the toy model, where the maximum of the
-  posterior pdf of the noisy points lies closer to the center of the true distribution than the true values.}
+  posterior PDF of the noisy points lies closer to the center of the true distribution than the true values.}
 \label{fig:posteriorCMD}
@@
 The median fractional change is 1.2.
-Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^{\corr}$ of about 0.6.
+Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K_s)^{\corr}$ of about 0.6.
 The right panel shows the posterior point estimates on the CMD colored by fraction change in variance.
@@
 
-The method produces a mixture-of-Gaussian parallax posterior pdf for every
+The method produces a mixture-of-Gaussian parallax posterior PDF for every
 star in the intersection of \tgas, \tmass, and the dust map.
@@
 It contains posterior expectations and variances for the stellar parallax.
-It also contains the full posterior pdf for each star, in the form of a set of pdf values,
+It also contains the full posterior PDF for each star, in the form of a set of PDF values,
 evaluated on a common grid in parallax.
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of 3--$5~{\rm Gyr}$.
+M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
 Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
@@
 
-We select sources from our catalog within a $1~{\rm deg}^2$ square window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
+We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
+To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \gaia\ within this window on the sky.
+Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
 With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
@@
 The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
-The posterior pdfs, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster (vertical lines) for far more stars.
-Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
+The posterior PDFs, shown in the bottom row, are more sharply peaked showing an increase in precision.
+The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
+This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{distancesM67.png}
-\caption{Proof of concept: parallax inference to the open cluster M67}
+\caption{Proof of concept: parallax inference to the open cluster M67.
+The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $1~{\rm deg}^2$ window centered on the position of M67.
+Our prior, visualized as a mixture of Gaussians, is shown in green.
+The upper right panel shows the \gaia\ parallax likelihoods in black, with the "true" distance to M67, determined by previous methods, bracketed by the red lines.
+Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
+The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
+The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
+The posterior PDFs are much more clustered and sharply peaked around the "true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
 \label{fig:m67}
@@
 by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
-and using it as a prior pdf for Bayesian inference.
-The posterior pdfs for distance that we obtain are, in general, much
+and using it as a prior PDF for Bayesian inference.
+The posterior PDFs for distance that we obtain are, in general, much
 narrower than the likelihood functions delivered by the
@@
 Because our model is representitive of our full subset catalog, we expect any (sensible) parallax estimates
-we generate from our parallax posterior pdfs to be only weakly biased.
+we generate from our parallax posterior PDFs to be only weakly biased.
 
@@
   \includegraphics[width=\textwidth]{comparePrior.png}
-\caption{Comparison with a common spatial prior: A visualization of the posterior pdf for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
-Left panel shows the posterior pdf using our \cmd\ prior, right panel shows the posterior pdfs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
+\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
+Left panel shows the posterior PDF using our \cmd\ prior, right panel shows the posterior PDFs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
 The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}.
-Using the \cmd\ as a prior generates a posterior pdf that has smaller variance.}
+Using the \cmd\ as a prior generates a posterior PDF that has smaller variance.}
 \label{fig:comparePrior}
@@
 \figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
-Here the parallax posterior pdf expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
+Here the parallax posterior PDF expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
 The left panel visualizes results using our \cmd\ prior, and the right panel visualizes results using the EDSD prior.
@@
   \includegraphics[width=\textwidth]{whatsThatFeature.pdf}
-\caption{Features in the \cmd: The expectation values from the posterior pdfs of the true parallax}
+\caption{Features in the \cmd: The expectation values from the posterior PDFs of the true parallax projected onto the \cmd.
+Here we have noted the features previously found in the \cmd\ that are also highlighted nicely by the \gaia\ + \tmass\ data.
+The upper and lower main sequence, as well as the main sequence turn off.
+There is a plausible binary sequence of stars, as well as a prominent red clump, and straight and narrow red giant branch. }
 \label{fig:wtf}
@@
 star formation history of the local volume also mixed with a Malmquist
-bias.
+bias (\citealt{malmquist22}).
 
@@
 
-The outputs of this project include posterior pdfs for stellar parallaxes.
+The outputs of this project include posterior PDFs for stellar parallaxes.
 It is tempting to treat these outputs as equivalent to some kind of catalog of
@@
 of a likelihood function, whereas our ouptut is a
-representation of a posterior pdf, one per star.
+representation of a posterior PDF, one per star.
 Posterior information used naively can lead to serious statistical
@@
 Multiplicative uses of the data will effectively take the prior
-pdf to a large power.
+PDF to a large power.
 We have (in previous work)
-given examples of correct uses of posterior samplings or pdfs
+given examples of correct uses of posterior samplings or PDFs
 for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we

### 2017-06-06 comparePrior.py

@@ -102,7 +102,7 @@ def comparePrior():
 
     for a in ax:
         a.set_xlim(-0.5, 1.5)
-        a.set_xlabel(r'$(J - K)^C$')
+        a.set_xlabel(r'$(J - K_s)^C$')
     ax[0].set_ylabel(r'$M_J^C$')
     ax[1].set_ylabel(r'$\varpi 10^{0.2\,m_J}$')
     ax[0].set_ylim(6, -6)
@@ -144,7 +144,7 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
         mag1 = 'J'
         mag2 = 'K'
         absmag = 'J'
-        xlabel = r'$(J-K)^C$'
+        xlabel = r'$(J-K_s)^C$'
         ylabel = r'$M_J^C$'
         xlim = [-0.25, 1.25]
         ylim = [6, -6]
@@ -266,7 +266,7 @@ def whatsThatFeature(survey='2MASS', quantile=0.05, dataFilename='All.npz', iter
         mag1 = 'J'
         mag2 = 'K'
         absmag = 'J'
-        xlabel = r'$(J-K)^C$'
+        xlabel = r'$(J-K_s)^C$'
         ylabel = r'$M_J^C$'
         xlim = [-0.25, 1.25]
         ylim = [6, -6]
@@ -412,7 +412,7 @@ def examplePosterior(nexamples=100, postFile='posteriorSimple.npz', dustFile='du
         plt.legend(loc='best')
         plt.tight_layout()
         plt.xlabel('parallax [mas]', fontsize=18)
-        plt.title('J-K: ' + '{0:.1f}'.format(color[i]) + '    M: ' +  '{0:.1f}'.format(testXD.absMagKinda2absMag(absMagKinda[i])))
+        plt.title(r'J-K$_s$: ' + '{0:.1f}'.format(color[i]) + '    M: ' +  '{0:.1f}'.format(testXD.absMagKinda2absMag(absMagKinda[i])))
         plt.savefig('exampleCMDPosteriorLargerVariance_' + str(i) + '.png')
 
 
@@ -423,7 +423,7 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
     absmag = 'J'
     mag1 = 'J'
     mag2 = 'K'
-    xlabel = '$(J-K)^C$'
[…]

### 2017-06-06 paper/paper.tex
noise assumption

--- 2017-06-05 paper/paper.tex
@@
   prior and the use of it for parallax inference has been done under the deep
-  assumption that the \gaia\ noise model is correct. If the
-  uncertainties are over-estimated then we could be over-deconvolving.
+  assumption that the \gaia\ noise model is correct, and an additional
+  assumption that the parallax uncertainty is always significantly larger
+  than the photometric uncertainty in the $J$ band, even after dust correction.
+  In the unlikely event (given these choices) that the
+  uncertainties are over-estimated, we could be over-deconvolving.
   That is, the data-driven model could be producing a tighter or more
@@
   side, wrong uncertainty estimates lead to further biases in the derived photometric parallaxes.
+  Since we are explicitly ignoring the (non-Gaussian, in parallax space) photometric
+  and dust-extinction uncertainties, our noise estimates are certainly biased, and
+  this will translate into small biases in posterior parallax estimates.
 \item[dust] Within our probabilistic framework, we improperly handle

### 2017-06-06 paper/paper.tex
toy model tweaks

--- 2017-06-06 paper/paper.tex
@@
 In the lower-left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
-The blue points represent the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
+The blue points represent the expectation value of the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
 The red points represent the actual true values.
-The lower-right panel shows the posterior belief of the true value for all the points.
+The lower-right panel shows the expectation value, and $1\sigma$ uncertainties, of the posterior belief of the true value for all the points.
 They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior.
@@
 Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
-To ease understanding, here we demonstrate the generaly methodology using a
+To ease understanding, here we demonstrate the general methodology using a
 simpler toy model with simulated data.
@@
 Similarly, noisy samples from this distribution are shown in the upper-right panel.
-Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noise distribution.
+Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noisy distribution.
 
@@
 To infer a posterior PDF of the true $y$ value for each measured $y$ value we need a prior.
-We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
+We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y_n$ data themselves.
 We learn the noise-deconvolved function using marginalized, maximum-likelihood estimation (MMLE), which marginalizes over the true data values and maximizes the likelihood of the data.
@@
 With large data sets, the MMLE or Empirical Bayes approximation is safe,
-in some sense because no individual data point makes a large difference to
+in some sense, because no individual data point makes a large difference to
 the inference.
@@
 \end{eqnarray}
-In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs of the true values $y_{\true,n}$ for the 5 highlighted black points from the upper-right panel.
+In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs (visualized as the expectation value and $1\sigma$ uncertainties) of the true values $y_{\true,n}$ for the 5 highlighted black points from the upper-right panel.
 The red points are the associated true $y$ values.
-In the lower-right panel of \figurename~\ref{fig:toy}, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
+In the lower-right panel of \figurename~\ref{fig:toy}, the blue points represent the posterior beliefs, again visualized as the expectation value and $1\sigma$ uncertainty, for all 1024 $y_n$.
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance, $t^2$, and the measurement variance, $\sigma_n^2$.
@@
 In contrast, more certain measurements remain closer to their measured values, $y_n$.
-We now have better estimates of each true y value by inferring each true y value from the measured y value and using an estimate of the true distribution the measured y value was drawn from as the prior.
+We now have better estimates of each true $y$ value by inferring each true $y$ value from the measured $y$ value using an estimate of the true distribution the measured $y$ value was drawn from as the prior.
 
@@
 
-For the generation of the prior and the evaluation of the likelihood, we need to
+To generate the prior and evaluate the likelihood, we need to
 correct the \tmass\ photometry for dust extinction.
@@
 In detail, we take the $5\%$
-quantile of each distance posterior (the closest part of the
-distance posterior) and query the 3D dust map at that distance.
+quantile of each parallax posterior (the closest part) and query the 3D dust map at that distance.
 We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
 correction because over-correcting for dust can push the likelihood
-function into an unphysical, different space of the \cmd\ and thus bias the
+function into an unphysical, extremely blue part of the \cmd\ and thus bias the
 parallax inference. At the 3D position of each star, we sample the \texttt{mode=median} value of the probabilistic dust map.
@@
 After applying the dust correction, we do not update the uncertainties, which is
-technically wrong and under-represents the true uncertainties; as mentioned
-previously, this leads to a less severe deconvolution.
+technically wrong and under-represents the true uncertainties; this leads to a less severe deconvolution, as mentioned previously.
 In other words, the features of our inferred
@@
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$.
-This is visualized by Examples 1-4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
-However, $14\%$ of our parllax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$.
+Examples of these slight increases in precision are visualized in Examples 1-4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
+However, $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$.
 This larger change in precision is visualized in Example 5, which shows a significantly narrower posterior PDF compared to the likelihood.
-The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact strongly depend on the location of each object in \cmd\ space.
-
-Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax PDFs derived with our \cmd\ prior.
-This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
-Those typically are the regions with the largest photometric and parallax errors, for the faintest objects.
+The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact, and shape, strongly depend on the location of each star in \cmd\ space.
+
+Figure~\ref{fig:posteriorCMD} shows the distribution of the expectation values of the posterior parallax PDFs projected back onto the \cmd\ space.
+The left panel shows the distribution for all 1.3 million stars in our dataset, and the right panel shows a subsample of the dataset but includes the $1\sigma$ uncertainties on the posterior PDFs.
+Compared with the raw data shown in Figure~\ref{fig:data}, it is clear the precision of the posterior PDFs is greater, especially for the red giant branch stars.
+
+This change in precision is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes in the variance.
+The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours over plotted.
+The points are colored by the natural log of their fractional change in variance (the y-axis of the left panel) in both panels to help guide your eye for the right panel.
+The regions of \cmd\ space with the greatest improvements are shown in black, and the regions with the lowest improvements are shown in yellow.
+The giants show obvious large improvements, as well as some main sequence stars.
+Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
 This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
+Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.
+
+Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog.
+This more directly shows what was alluded to in Figure~ref{fig:posterior}.
+For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.
+Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$, shown as the dotted line at $y=0.5$.
+And $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, or have a fraction decrease in variance by a factor of $>4$, shown as the other dotted line at $y=0.14$.
+
 \begin{figure}

### 2017-06-06 paper/paper.tex
abstract tweaks

@@
 parallax measurements under an assumption of stationarity.
-We use the Extreme Deconvolution (\xd) algorithm---an approximation to a full
+We use the Extreme Deconvolution (\xd) algorithm---an Empirical Bayes approximation to a full
 hierarchical model of the true parallax and photometry of every star---to
 construct this prior.
-The \xd-optimized prior is used to infer a precise photometric parallax estimate
+The prior is combined with a \tgas\ likelihood to infer a precise photometric parallax estimate
 and uncertainty (and full posterior) for every star.
@@
 estimates, but appears as a cluster in our posterior parallax estimates.
-Our results, including a posterior parallax and distance sampling for 1.4
+Our results, including a parallax posterior pdf for each of 1.4
 million \tgas\ stars, are available in companion electronic tables.

### 2017-06-06 paper/paper.tex
title
typo

@@
 % start and title
-\shorttitle{improving gaia parallaxes}
+\shorttitle{improving gaia parallax precision}
 \shortauthors{anderson et al.}
 \begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me
-\title{Improving \textsl{Gaia} parallaxes
-       with a data-driven model \\
-       of the color--magnitude diagram}
+\title{Improving \textsl{Gaia} parallax precision \\
+       with a data-driven model of stars}
 
@@
 \gaia\ \tgas\ Catalog and is incredibly informative about stellar distances.
-Here we use \tmass\ colors for $1.2$ million \tgas\ stars to build a
+Here we use \tmass\ colors for $1.4$ million \tgas\ stars to build a
 noise-deconvolved empirical prior distribution for stars in color--magnitude

### 2017-06-06 paper/paper.tex
dashes

@@
 Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $Y$.
-We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
+We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the $y$ direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
 Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
@@
 \item[big data] We assume that we have large numbers of stars, large enough that
-  empirical Bayes, or maximum-marginalized likelihood, is a safe
+  empirical-Bayes, or maximum-marginalized-likelihood, is a safe
   approximation to full Bayesian inference. This assumption will be
@@
 giant stars, which tend to have the lowest signal to noise, and can
-therefore have dust corrections $> 2$ magnitudes. Although the dust
+therefore have dust corrections $> 2$~mag. Although the dust
 model is probabilistic, as well as our distance inference, to move
@@
 precise parallax posterior to get a measurement of dust in the 3D dust
-map for each star. We then iterate this process 10 times, however
-the dust values seem to converge after about 2-3 iterations. The
+map for each star. We then iterate this process ten times, however
+the dust values seem to converge after a few iterations. The
 iterative process allows for slight variations in distance and
@@
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$.
-Examples of these slight increases in precision are visualized in Examples 1-4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
+Examples of these slight increases in precision are visualized in Examples 1 through 4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
 However, $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$.
@@
 The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours over plotted.
-The points are colored by the natural log of their fractional change in variance (the y-axis of the left panel) in both panels to help guide your eye for the right panel.
+The points are colored by the natural log of their fractional change in variance (the $y$ axis of the left panel) in both panels to help guide your eye for the right panel.
 The regions of \cmd\ space with the greatest improvements are shown in black, and the regions with the lowest improvements are shown in yellow.
@@
 Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1-4 are not very effected by the prior (green), where good data is good data.
+Examples 1 through 4 are not very effected by the prior (green), where good data is good data.
 Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior PDF, transformed into an absolute magnitude, remains on the main sequence.
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
+M67 is at a Heliocentric distance of $800$ to $900~{\rm pc}$ with an estimated age of $3$ to $5~{\rm Gyr}$.
 Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
@@
 for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
-encourage power-users to consult those methodological
+encourage power users to consult those methodological
 contributions before engaging with these outputs.

### 2017-06-06 paper/paper.tex
Added full affiliations

--- 2017-06-06 paper/paper.tex
@@
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
+\affil{NASA Einstein Fellow}
 
@@
 
-This project was partially supported by This research was partially supported by
+This research was partially supported by
   the \acronym{NSF} (\acronym{AST-1517237}),
@@
   and the Moore-Sloan Data Science Environment at \acronym{NYU}. 
+  BL was supported by the National Aeronautics and Space Administration through Einstein Postdoctoral Fellowship Award Number PF6-170154.
 It made use of the \acronym{NASA} Astrophysics Data System.

### 2017-06-06 paper/paper.tex
removed (beautiful) from abstract

@@
 \author[0000-0001-5725-9329]{Lauren Anderson}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
 
 \author[0000-0003-2866-9403]{David W. Hogg}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
@@
 This model contains no knowledge of stellar astrophysics or the
-Milky Way, but is precise (and beautiful) because it is a generative model fit to many noisy
+Milky Way, but is precise because it is a generative model fit to many noisy
 parallax measurements under an assumption of stationarity.

### 2017-06-06 paper/paper.tex
Added footnote about deconvolution

@@
 In other words, the features of our inferred
-  \cmd\ are conservatively broad (broader than if we over-estimated
-  the errors which would have lead to excessive deconvolution and
-  unrealistically narrow features). We could add the covariance of
+  \cmd\ are conservatively broad\footnote{This can be understood by considering the following procedure: if one incrementally increase the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution will be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added}.
+We could add the covariance of
 the dust to our likelihoods, but this is beyond the scope of this

### 2017-06-06 paper/paper.tex
typos
repetition

@@
 
-Here we deliver a demonstration-of-concept that shows that a
+Here we present a demonstration-of-concept that shows that a
 data-driven model of the \gaia\ data could in principle deliver very
@@
 $2\sigma$ contours of the distribution.
-The right panel shows a subset of the data with it's associated error bars.
+The right panel shows a subset of the data with it's associated error bars, calculated via subsampling.
 The uncertainties in the colors are fairly well behaved, but the large

### 2017-06-06 paper/paper.tex
typos

@@
 This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
-The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
+The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
 The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
@@
 
-We use stars crossmatched in \tgas\ and \tmass\.
-The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
-\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
+We use stars crossmatched in \tgas\ and \tmass.
+The match was done using a nearest-neighbor algorithm with a search radius of $4"$.\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
 We also required that the stars lie within the

### 2017-06-06 paper/paper.tex
edited footnote

@@
 In other words, the features of our inferred
-  \cmd\ are conservatively broad\footnote{This can be understood by considering the following procedure: if one incrementally increase the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution will be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added}.
+  \cmd\ are conservatively broad.\footnote{This can be understood by considering the following procedure: if one incrementally increased the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution would be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added.
+Thus, under-estimating the uncertainties leads to a conservative under-deconvolution.}
 We could add the covariance of

### 2017-06-06 paper/paper.tex
removed "bad things"!

@@
 M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
-Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
+Previous work \citep{cbj15} has noted that distance estimates directly derived from the inverse of the observed (noisy) parallax are unreliable (significantly biased and uncertain) when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
 With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.

### 2017-06-06 paper/paper.tex
Typos

@@
 Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
-This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
+This is a fairly standard result which we illustrated in the introductory section of this paper: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
 Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.
@@
 Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog.
-This more directly shows what was alluded to in Figure~ref{fig:posterior}.
+This more directly shows what was alluded to in Figure~\ref{fig:posterior}.
 For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.

### 2017-06-06 paper/paper.tex
fixed quotes

@@
 Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
-This is a fairly standard result which we illustrated in the introductory section of this paper: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
+This is a fairly standard result which we illustrated in the toy model above: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
 Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.
@@
 Our prior, visualized as a mixture of Gaussians, is shown in green.
-The upper right panel shows the \gaia\ parallax likelihoods in black, with the "true" distance to M67, determined by previous methods, bracketed by the red lines.
+The upper right panel shows the \gaia\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
 Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
@@
 The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
-The posterior PDFs are much more clustered and sharply peaked around the "true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
+The posterior PDFs are much more clustered and sharply peaked around the ``true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
 \label{fig:m67}

### 2017-06-06 paper/paper.tex
move citation

@@
 with a Malmquist bias of seeing a much larger volume of upper main
-sequence stars. There is a plausible binary sequence of stars with
+sequence stars \citep{malmquist22}. There is a plausible binary sequence of stars with
 about twice the brightness as the main sequence. The helium burning
@@
 star formation history of the local volume also mixed with a Malmquist
-bias (\citealt{malmquist22}).
+bias.
 

### 2017-06-06 paper/paper.tex
comment and typo
added sentences about selection function

@@
   drawn from the same distribution within some feature or patch of the
-  \cmd\ space.
+  \cmd\ space. Thus, this has a small effect on our results, and could be improved by modelling the noise distribution simultaneously with the \cmd.
 \item[selection] The biggest limitation of our method for generating
@@
   Milky Way.  Importantly, because we used \emph{all} of the stars in
-  the \tgas--\tmass\ intersection, the model \emph{is} representive of
+  the \tgas--\tmass\ intersection, the model \emph{is} representative of
   the full intersection, even the distant parts, where all the
@@
   bluer than those included in the \tgas--\tmass\ intersection.
+  A similar argument can be made for \textit{internal} selection effects, such as inhomogeneities of the depth and noise distributions, which we have neglected. 
 \item[big data] The empirical Bayesian methodology is a good

### 2017-06-06 paper/paper.tex
selection function stuff again

@@
   A similar argument can be made for \textit{internal} selection effects, such as inhomogeneities of the depth and noise distributions, which we have neglected. 
+  Those inhomogeneities are unlikely to affect the \cmd results; we effectively model the average distribution of the \tgas--\tmass\ intersection. 
+  By modeling the selection function in detail, one could improve the quality of the inferred \cmb, its effect on individual objects, and its connection to physical models of the Milky Way. 
 \item[big data] The empirical Bayesian methodology is a good
@@
   for that would be to fit not the \cmd\ but rather deviations of the \cmd\ away
-  from a good physics-based model prediction.
+  from a good physics-based model prediction. This would also require modeling the selection function of the data under consideration.
 \end{description}

### 2017-06-06 paper/paper.tex
typo

@@
   Those inhomogeneities are unlikely to affect the \cmd results; we effectively model the average distribution of the \tgas--\tmass\ intersection. 
-  By modeling the selection function in detail, one could improve the quality of the inferred \cmb, its effect on individual objects, and its connection to physical models of the Milky Way. 
+  By modeling the selection function in detail, one could improve the quality of the inferred \cmd, its effect on individual objects, and its connection to physical models of the Milky Way. 
 \item[big data] The empirical Bayesian methodology is a good

### 2017-06-07 plots.py
getting plots together

@@ -1,0 +1,189 @@
+import toy
+import os
+import testXD
+from astroML.plotting import setup_text_plots
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import corner
+from xdgmm import XDGMM
+import drawEllipse
+
+def plotPrior(xdgmm, ax, c='black'):
+    for gg in range(xdgmm.n_components):
+        points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
+        ax.plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=1, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
+
+
+def plot_samples(x, y, xerr, yerr, ind, contourColor='black', rasterized=True, plot_contours=True, dataColor='black', titles=None, xlim=(0,1), ylim=(0,1), xlabel=None, ylabel=None, prior=False, xdgmm=None, pdf=False):
+    setup_text_plots(fontsize=16, usetex=True)
+    plt.clf()
+    alpha = 0.1
+    alpha_points = 0.01
+    fig = plt.figure(figsize=(12, 5.5))
+    fig.subplots_adjust(left=0.1, right=0.95,
+                            bottom=0.15, top=0.95,
+                            wspace=0.1, hspace=0.1)
+    ax1 = fig.add_subplot(121)
+    ax2 = fig.add_subplot(122)
+    levels = 1.0 - np.exp(-0.5 * np.arange(1.0, 2.1, 1.0) ** 2)
+    im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor, rasterized=True, plot_contours=plot_contours, plot_datapoints=False)
+    ax1.scatter(x, y, s=1, lw=0, c=dataColor, alpha=alpha, zorder=0)
+    if prior:
+        plotPrior(xdgmm, ax2, c=dataColor)
+    else:
+        ax2.errorbar(x[ind], y[ind], xerr=xerr[ind], yerr=[yerr[0][ind], yerr[1][ind]], fmt="none", zorder=0, mew=0, ecolor='black', alpha=0.5, elinewidth=0.5)
+    ax = [ax1, ax2]
+    for i, axis in enumerate(ax):
+        axis.set_xlim(xlim)
+        axis.set_ylim(ylim[0], ylim[1]*1.1)
[…]

### 2017-06-07 paper/paper.tex
I don't want my office room number there :-)
fixing figures

--- 2017-06-06 paper/paper.tex
@@
 \author[0000-0001-5725-9329]{Lauren Anderson}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 
 \author[0000-0003-2866-9403]{David W. Hogg}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
@@
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
-\affil{NASA Einstein Fellow}
 
@@
 This model contains no knowledge of stellar astrophysics or the
-Milky Way, but is precise because it is a generative model fit to many noisy
+Milky Way, but is precise (and beautiful) because it is a generative model fit to many noisy
 parallax measurements under an assumption of stationarity.
@@
 
-Here we present a demonstration-of-concept that shows that a
+Here we deliver a demonstration-of-concept that shows that a
 data-driven model of the \gaia\ data could in principle deliver very
@@
 \centering
-  \includegraphics[width=\textwidth]{toy.png}
+  \includegraphics[width=\textwidth]{toy.pdf}
 \caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior.
@@
 simpler toy model with simulated data.
-Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $Y$.
-We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
+Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $y$.
+We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the $y$ direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
 Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
@@
 This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
-The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
+The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
 The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
@@
 
-We use stars crossmatched in \tgas\ and \tmass.
-The match was done using a nearest-neighbor algorithm with a search radius of $4"$.\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
+We use stars crossmatched in \tgas\ and \tmass\.
+The match was done using a nearest-neighbor algorithm with a search radius of $4"$.
+\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
 We also required that the stars lie within the
@@
 $2\sigma$ contours of the distribution.
-The right panel shows a subset of the data with it's associated error bars, calculated via subsampling.
+The right panel shows a subset of the data with it's associated error bars.
 The uncertainties in the colors are fairly well behaved, but the large
@@
 In other words, the features of our inferred
-  \cmd\ are conservatively broad.\footnote{This can be understood by considering the following procedure: if one incrementally increased the noise levels (leaving the data unchanged; just increasing the declared uncertainties), the deconvolved distribution would be increasingly narrower, since it targets the distribution that would generate the noiseless observations, as well as the observed data once noise is added.
-Thus, under-estimating the uncertainties leads to a conservative under-deconvolution.}
-We could add the covariance of
+  \cmd\ are conservatively broad (broader than if we over-estimated
+  the errors which would have lead to excessive deconvolution and
+  unrealistically narrow features). We could add the covariance of
 the dust to our likelihoods, but this is beyond the scope of this
@@
 Regions of large improvement are those with the largest photometric and parallax errors, which are the faintest observed objects.
-This is a fairly standard result which we illustrated in the toy model above: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
+This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
 Stars with a color $(J-K_s)^C \sim 0.6$ show a decrease in precision due to the width of the prior there; the more vertical structure in the red giant branch of the \cmd.
@@
 Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog.
-This more directly shows what was alluded to in Figure~\ref{fig:posterior}.
+This more directly shows what was alluded to in Figure~ref{fig:posterior}.
 For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
-Previous work \citep{cbj15} has noted that distance estimates directly derived from the inverse of the observed (noisy) parallax are unreliable (significantly biased and uncertain) when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
+M67 is at a Heliocentric distance of $800$ to $900~{\rm pc}$ with an estimated age of $3$ to $5~{\rm Gyr}$.
+Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
 With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
@@
 Our prior, visualized as a mixture of Gaussians, is shown in green.
-The upper right panel shows the \gaia\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
+The upper right panel shows the \gaia\ parallax likelihoods in black, with the "true" distance to M67, determined by previous methods, bracketed by the red lines.
 Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
@@
 The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
-The posterior PDFs are much more clustered and sharply peaked around the ``true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
+The posterior PDFs are much more clustered and sharply peaked around the "true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
 \label{fig:m67}
@@
 with a Malmquist bias of seeing a much larger volume of upper main
-sequence stars \citep{malmquist22}. There is a plausible binary sequence of stars with
+sequence stars. There is a plausible binary sequence of stars with
 about twice the brightness as the main sequence. The helium burning
@@
   drawn from the same distribution within some feature or patch of the
-  \cmd\ space. Thus, this has a small effect on our results, and could be improved by modelling the noise distribution simultaneously with the \cmd.
+  \cmd\ space.
 \item[selection] The biggest limitation of our method for generating
@@
   Milky Way.  Importantly, because we used \emph{all} of the stars in
-  the \tgas--\tmass\ intersection, the model \emph{is} representative of
+  the \tgas--\tmass\ intersection, the model \emph{is} representive of
   the full intersection, even the distant parts, where all the
@@
   bluer than those included in the \tgas--\tmass\ intersection.
-  A similar argument can be made for \textit{internal} selection effects, such as inhomogeneities of the depth and noise distributions, which we have neglected. 
 \item[big data] The empirical Bayesian methodology is a good
@@
 
-This research was partially supported by
+This project was partially supported by This research was partially supported by
   the \acronym{NSF} (\acronym{AST-1517237}),
@@
   and the Moore-Sloan Data Science Environment at \acronym{NYU}.
-  BL was supported by the National Aeronautics and Space Administration through Einstein Postdoctoral Fellowship Award Number PF6-170154.
 It made use of the \acronym{NASA} Astrophysics Data System.

### 2017-06-07 plots.py

@@ -28,11 +28,11 @@ def plot_samples(x, y, xerr, yerr, ind, contourColor='black', rasterized=True, p
     ax2 = fig.add_subplot(122)
     levels = 1.0 - np.exp(-0.5 * np.arange(1.0, 2.1, 1.0) ** 2)
     im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor, rasterized=True, plot_contours=plot_contours, plot_datapoints=False)
-    ax1.scatter(x, y, s=1, lw=0, c=dataColor, alpha=alpha, zorder=0)
+    ax1.scatter(x, y, s=1, lw=0, c=dataColor, alpha=alpha, zorder=0, rasterized=True)
     if prior:
         plotPrior(xdgmm, ax2, c=dataColor)
     else:
-        ax2.errorbar(x[ind], y[ind], xerr=xerr[ind], yerr=[yerr[0][ind], yerr[1][ind]], fmt="none", zorder=0, mew=0, ecolor='black', alpha=0.5, elinewidth=0.5)
+        ax2.errorbar(x[ind], y[ind], xerr=xerr[ind], yerr=[yerr[0][ind], yerr[1][ind]], fmt="none", zorder=0, mew=0, ecolor=dataColor, alpha=0.5, elinewidth=0.5)
     ax = [ax1, ax2]
     for i, axis in enumerate(ax):
         axis.set_xlim(xlim)
@@ -44,7 +44,7 @@ def plot_samples(x, y, xerr, yerr, ind, contourColor='black', rasterized=True, p
             axis.yaxis.set_major_formatter(plt.NullFormatter())
         else:
             axis.set_ylabel(ylabel, fontsize = 18)
-    if pdf: fig.savefig('plot_sample.pdf', rasterized=True)
+    if pdf: fig.savefig('plot_sample.pdf', dpi=400)
     fig.savefig('plot_sample.png')
 
 
@@ -68,7 +68,7 @@ def sampleXDGMM(xdgmm, Nsamples):
     return samplex, sampley
 
 #    for label, style in zip(['paper', 'talk'],['seaborn-paper', 'seaborn-talk']):
-pdf = False
+pdf = True
 style = 'seaborn-paper'
 plt.style.use(style)
 mpl.rcParams['xtick.labelsize'] = 18
@@ -123,7 +123,7 @@ absMag_err = absMagError(tgas['parallax'][positive], tgas['parallax_error'][posi
[…]

### 2017-06-07 plots.py
made some figures pdf

@@ -184,6 +184,19 @@ absMag = testXD.absMagKinda2absMag(mean[positive]*10.**(0.2*apparentMagnitude[po
 absMag_err = absMagError(mean[positive], sigma[positive], apparentMagnitude[positive], absMag)
 titles = ["De-noised Expectation Values", "Posterior Distributions"]
 plot_samples(color, absMag, color_err, absMag_err, ind, contourColor='black', rasterized=True, plot_contours=True, dataColor=posteriorColor, titles=titles, xlim=xlim_cmd, ylim=ylim_cmd, xlabel=xlabel_cmd, ylabel=ylabel_cmd, pdf=pdf)
-if pdf: os.rename('plot_sample.pdf', 'paper/posterior.pdf')
+if pdf: os.rename('plot_sample.pdf', 'paper/posteriorCMD.pdf')
 os.rename('plot_sample.png', 'posterior.png')
 #-------------------------------------------------------
+
+
+#dust plot
+
+#posterior example plot
+
+#delta plot
+
+#delta cdf plot
+
+#M67 plot
+
+#what's that feature plot 

### 2017-06-07 paper/paper.tex
made some figures pdf

--- 2017-06-06 paper/paper.tex
@@
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 \affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
@@
 \author[0000-0002-3962-9274]{Boris Leistedt}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, room 1005, New York, NY 10003, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 \affil{NASA Einstein Fellow}
@@
 \centering
-  \includegraphics[width=\textwidth]{data.png}
+  \includegraphics[width=\textwidth]{data.pdf}
 \caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K_s$ color, the $J$ band apparent magnitude, and the \gaia\ parallax.
@@
 \centering
-  \includegraphics[width=\textwidth]{prior_ngauss128.png}
+  \includegraphics[width=\textwidth]{prior.pdf}
 \caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by
@@
 \centering
-  \includegraphics[width=\textwidth]{posteriorCMD.png}
+  \includegraphics[width=\textwidth]{posteriorCMD.pdf}
 \caption{The distribution of expectation values of the parallax posterior PDFs transformed into an absolute magnitude vs color, the \cmd, of the full data set is shown in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution shown in black.
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
+
 M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
 Previous work \citep{cbj15} has noted that distance estimates directly derived from the inverse of the observed (noisy) parallax are unreliable (significantly biased and uncertain) when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
+
 With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
@@
 \centering
-  \includegraphics[width=\textwidth]{comparePrior.png}
+  \includegraphics[width=\textwidth]{comparePrior.pdf}
 \caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.

### 2017-06-08 paper/paper.tex
smaller

--- 2017-06-07 paper/paper.tex
@@
 \centering
-  \includegraphics[width=\textwidth]{toy.pdf}
+  \includegraphics[width=\textwidth]{toy.png}
 \caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior.
@@
 simpler toy model with simulated data.
-Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $y$.
+Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $Y$.
 We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the $y$ direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
@@
 \centering
-\includegraphics[width=\textwidth]{deltaCDF.png}
+\includegraphics[width=0.75\textwidth]{deltaCDF.png}
 \caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.

### 2017-06-08 comparePrior.py
plot tweaks;

@@ -111,23 +111,23 @@ def comparePrior():
     plt.tight_layout()
     fig.savefig('priorNgaussComparison.png')
 
-def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz'):
+def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz', ax=None, tgas=None):
 
-    tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+    if tgas is None: tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
     dustFile = dustFilename(ngauss, quantile, iter, survey, dataFilename)
     data = np.load(dustFile)
     dust = data['ebv']
-    fig, ax = plt.subplots(figsize=(12,7))
+    if ax is None: fig, ax = plt.subplots(figsize=(12,7))
     norm = mpl.colors.PowerNorm(gamma=1/2.)#(vmin=-0.5, vmax=1)
-    im = ax.scatter(tgas['l'], tgas['b'], c=dust, lw=0, cmap='Greys', s=1, vmin=0, vmax=1, norm=norm)
+    im = ax.scatter(tgas['l'], tgas['b'], c=dust, lw=0, cmap='Greys', s=1, vmin=0, vmax=1, norm=norm, rasterized=True)
     ax.set_xlim(0, 360)
     ax.set_ylim(-90, 90)
-    ax.set_xlabel('$\mathscr{l}$ [deg]')
-    ax.set_ylabel('$\mathscr{b}$ [deg]')
+    ax.set_xlabel(r'$\mathscr{l}$ [deg]')
+    ax.set_ylabel(r'$\mathscr{b}$ [deg]')
     cb = plt.colorbar(im, ax=ax)
     cb.set_clim(-0.1, 1)
     cb.set_label(r'E($B-V$ )')
-    fig.savefig('dustViz.dQ' + str(quantile) + '.png')
+    if ax is None: fig.savefig('dustViz.dQ' + str(quantile) + '.png')
 
 def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz', sdss5=False, whatsThatFeature=False):
 
@@ -595,7 +595,7 @@ def paperComparePrior(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
[…]

### 2017-06-08 plots.py

@@ -8,11 +8,17 @@ import matplotlib as mpl
 import corner
 from xdgmm import XDGMM
 import drawEllipse
+import comparePrior
+import scipy.interpolate
 
-def plotPrior(xdgmm, ax, c='black'):
+def gaussian(mean, sigma, array, amplitude=1.0):
+    return amplitude/np.sqrt(2.*np.pi*sigma**2.)*np.exp(-(array - mean)**2./(2.*sigma**2.))
+
+
+def plotPrior(xdgmm, ax, c='black', lw=1):
     for gg in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
-        ax.plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=1, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
+        ax.plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=lw, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
 
 
 def plot_samples(x, y, xerr, yerr, ind, contourColor='black', rasterized=True, plot_contours=True, dataColor='black', titles=None, xlim=(0,1), ylim=(0,1), xlabel=None, ylabel=None, prior=False, xdgmm=None, pdf=False):
@@ -46,7 +52,7 @@ def plot_samples(x, y, xerr, yerr, ind, contourColor='black', rasterized=True, p
             axis.set_ylabel(ylabel, fontsize = 18)
     if pdf: fig.savefig('plot_sample.pdf', dpi=400)
     fig.savefig('plot_sample.png')
-
+    plt.close(fig)
 
 
 def absMagError(parallax, parallax_err, apparentMag, absMag):
@@ -66,6 +72,21 @@ def sampleXDGMM(xdgmm, Nsamples):
     samplex = sample[:,0]
     sampley = testXD.absMagKinda2absMag(sample[:,1])
     return samplex, sampley
+
+def likePriorPost(color, absMagKinda, color_err, absMagKinda_err, apparentMagnitude, xdgmm, ndim=2, nPosteriorPoints=1000, projectedDimension=1):
+    meanData, covData = testXD.matrixize(color, absMagKinda, color_err, absMagKinda_err)
+    meanPrior, covPrior = testXD.matrixize(color, absMagKinda, color_err, 1e5)
+    meanData = meanData[0]
+    covData = covData[0]
+    meanPrior = meanPrior[0]
[…]

### 2017-06-08 testXD.py

@@ -229,7 +229,7 @@ def posterior2d(means, amps, covs, xbins, ybins, nperGauss=100000., plot=False):
     Z, xedges, yedges = np.histogram2d(samplesX, absMagKinda2absMag(samplesY), bins=[xbins, ybins], normed=True)
     return xedges, yedges, Z, samplesY
 
-def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike, covLike, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=None, Y=None, Z=None, plot2D=False, samplesY=None, axAll=None, dataColor='black', posteriorColor='blue', truthColor='red', likeLabel=None, postLabel=None):
+def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike, covLike, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=None, Y=None, Z=None, plot2D=False, samplesY=None, axAll=None, dataColor='black', posteriorColor='blue', truthColor='red', likeLabel=None, postLabel=None, xlim=[0,1], ylim=[0,1], xlabel=None, ylabel=None, like2D_lw=1):
     """
     plot the posterior of each star into example.png and feed it axAll to put it on another plot
     """
@@ -265,8 +265,8 @@ def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike,
     pointsData = drawEllipse.plotvector(meanLike, covLike)
     ax[0].plot(pointsData[0, :], absMagKinda2absMag(pointsData[1,:]), 'g', lw=2)
     if axAll is not None:
-        axAll[0].plot(pointsData[0, :], absMagKinda2absMag(pointsData[1,:]), dataColor, lw=2, label=likeLabel)
-        axAll[2].plot(pointsData[0, :], absMagKinda2absMag(pointsData[1,:]), dataColor, lw=2, label=likeLabel)
+        axAll[0].plot(pointsData[0, :], absMagKinda2absMag(pointsData[1,:]), dataColor, lw=like2D_lw, label=likeLabel)
+        axAll[2].plot(pointsData[0, :], absMagKinda2absMag(pointsData[1,:]), dataColor, lw=like2D_lw, label=likeLabel)
 
     #plot the posterior in parallax space
     parallaxLikelihood = st.gaussian(meanLike[1], np.sqrt(covLike[1,1]), xabsMagKinda)
[…]

### 2017-06-12 paper/paper.tex
paper figure updates

--- 2017-06-07 paper/paper.tex
@@
 
-%@archiver{posteriorCMD.png}
+%@archiver{posteriorCMD.pdf}
 
@@
 These spectrophotometric models, in turn, along with \gaia's on-board
-low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
+low-resolution $BP/RP$ spectrophotometry, are used to provide
 parallax estimates to less precise stars.
@@
 where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
-The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain (Gaussian uncertainties are required to use \xd, see below).
+The transformation of the dust corrected absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
@@
 \covar_n = \begin{bmatrix}
-\sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
-0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^{\corr}}\right)^2
+\left(\sigma_{\varpi}\,10^{0.2\,J_n^{\corr}}\right)^2 & 0 \\
+0 & \sigma_{J,n}^2 + \sigma_{K,n}^2
 \end{bmatrix}
@@
 We use stars crossmatched in \tgas\ and \tmass.
-The match was done using a nearest-neighbor algorithm with a search radius of $4"$.\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}
+The match was done using a nearest-neighbor algorithm with a search radius of $4$ arcsec\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}.
 We also required that the stars lie within the
@@
 $2\sigma$ contours of the distribution.
-The right panel shows a subset of the data with it's associated error bars, calculated via subsampling.
+The right panel shows a subset of the data with its associated error bars, calculated via subsampling.
 The uncertainties in the colors are fairly well behaved, but the large
@@
 \centering
-  \includegraphics[width=\textwidth]{dust.png}
+  \includegraphics[width=\textwidth]{dust.pdf}
 \caption{The converged dust values at the 5\% distance quantile in Galactic coordinates.
@@
 \centering
-\includegraphics[width=\textwidth]{delta.png}
+\includegraphics[width=\textwidth]{delta.pdf}
 \caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color.
@@
 \centering
-\includegraphics[width=\textwidth]{deltaCDF.png}
+\includegraphics[width=\textwidth]{deltaParallax.pdf}
+\caption{No bias in posterior pdf parallax expectation values.}
+\label{fig:delta}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth]{deltaCDF.pdf}
 \caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.
@@
 \centering
-  \includegraphics[width=\textwidth]{distancesM67.png}
+  \includegraphics[width=\textwidth]{m67.pdf}
 \caption{Proof of concept: parallax inference to the open cluster M67.

### 2017-06-12 plots.py

@@ -11,6 +11,76 @@ import drawEllipse
 import comparePrior
 import scipy.interpolate
 
+def figsize_and_margins(plotsize,subplots=(1,1),**absolute_margins):
+    '''Determine figure size and margins from plot size and absolute margins
+
+       Parameters:
+         plotsize: (width, height) of plot area
+         subplots: (nrows, ncols) of subplots
+         left, right, top, bottom: absolute margins around plot area
+         wspace, hspace: width and height spacing between subplots
+       Returns:
+         size: figure size for figsize argument of figure()
+         margins: relative margins dict suitable for subplots_adjust()
+
+       Example: making 2x2 grid of 3" square plots with specific spacings:
+
+       sz, rm = figsize_and_margins((3,3), (2,2), left=1, right=.5,
+                                                  top=.5, bottom=1,
+                                                  wspace=.5, hspace=.5)
+       figure(figsize=sz)
+       subplots_adjust(**rm)
+       subplot(221); subplot(222)
+       subplot(223); subplot(224)
+    '''
+    #from matplotlib import rcParams
+    pw,ph = plotsize
+    nr,nc = subplots
+    amarg = absolute_margins
+    #dictionary for relative margins
+    # initialize from rcParams with margins not in amarg
+    rmarg = dict((m, mpl.rcParams['figure.subplot.' + m])
+                for m in ('left','right','top','bottom','wspace','hspace')
+                if m not in amarg
+            )
+    #subplots_adjust wants wspace and hspace relative to plotsize:
+    if 'wspace' in amarg: rmarg['wspace'] = float(amarg['wspace']) / pw
+    if 'hspace' in amarg: rmarg['hspace'] = float(amarg['hspace']) / ph
+    #in terms of the relative margins:
+    #width  * (right - left)
+    #    = ncols * plot_width  + (ncols - 1) * wspace * plot_width
+    #height * (top - bottom)
+    #    = nrows * plot_height + (nrows - 1) * hspace * plot_height
[…]

### 2017-06-12 testXD.py

@@ -29,7 +29,7 @@ def convert2gal(ra, dec):
     """
     return SkyCoord([ra, dec], unit=(units.hourangle, units.deg))
 
-def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963'):
+def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963', twoMass=None):
     """
     return the indices of tgas which fall within dl and db of l and b.
     default is M67
@@ -39,6 +39,11 @@ def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963'):
             (tgas['b'] > np.float(b) - db) & \
             (tgas['l'] < np.float(l) + dl) & \
             (tgas['l'] > np.float(l) - dl)
+
+    if twoMass is not None:
+        print twoMass['j_cmsig'][index], twoMass['k_cmsig'][index]
+        goodError = (twoMass['j_cmsig'] < 2.) & (twoMass['k_cmsig'] < 2.)
+        index = index & goodError
     if plot:
         fig, ax = plt.subplots()
         ax.scatter(tgas['l'][index], tgas['b'][index], alpha=0.5, lw=0)
@@ -229,7 +234,7 @@ def posterior2d(means, amps, covs, xbins, ybins, nperGauss=100000., plot=False):
     Z, xedges, yedges = np.histogram2d(samplesX, absMagKinda2absMag(samplesY), bins=[xbins, ybins], normed=True)
     return xedges, yedges, Z, samplesY
 
-def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike, covLike, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=None, Y=None, Z=None, plot2D=False, samplesY=None, axAll=None, dataColor='black', posteriorColor='blue', truthColor='red', likeLabel=None, postLabel=None, xlim=[0,1], ylim=[0,1], xlabel=None, ylabel=None, like2D_lw=1):
[…]

### 2017-06-12 toy/toy.py

@@ -42,47 +42,49 @@ def denoise_one_datum(xn, yn, sigman, m, b, t):
     return (yn * s2inv + (m * xn + b) * t2inv) / (s2inv + t2inv), \
         np.sqrt(1. / (s2inv + t2inv))
 
-def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5, dataColor='k', priorColor='g', posteriorColor='b', trueColor='r', posteriorMap = 'Blues'):
+def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5, dataColor='k', priorColor='g', posteriorColor='b', trueColor='r', posteriorMap = 'Blues', markersize=20):
 
     alpha_all = 0.05
     alpha_chosen = 1.0
     xlim = (-1, 1)
     ylim = (-5, 5)
-
+    markersize=markersize
     for ax in axes[1:-1]:
-        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color=dataColor, alpha=alpha_all ,mew=0)
-        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
-
+        ax.errorbar(xns, yns, yerr=sigmans, fmt=None, color=dataColor, alpha=alpha_all ,mew=0)
+        ax.scatter(xns, yns, s=markersize, alpha=alpha_all, color=dataColor)
+        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="none", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
+    axes[1].scatter(xns[0:nexamples], yns[0:nexamples], s=markersize, color=dataColor, alpha=alpha_chosen)
+    axes[2].scatter(xns[0:nexamples], yns[0:nexamples], s=2.*markersize, color=dataColor, alpha=alpha_chosen)
     xp = np.array(xlim)
     axes[0].plot(xp, mtrue*xp + btrue + ttrue, color=trueColor, linewidth=2, alpha=0.75, label=r'$y=m_{true}\,x+b_{true}\pm t$')
[…]

### 2017-06-13 plots.py
updated example plot

@@ -175,17 +175,17 @@ def likePriorPost(color, absMagKinda, color_err, absMagKinda_err, apparentMagnit
 
 #    for label, style in zip(['paper', 'talk'],['seaborn-paper', 'seaborn-talk']):
 pdf = True
-plot_data = True
-plot_dust = True
-plot_prior = True
-plot_m67 = True
-plot_compare = True
-plot_expectation = True
+plot_data = False
+plot_dust = False
+plot_prior = False
+plot_m67 = False
+plot_compare = False
+plot_expectation = False
 plot_examples = True
-plot_delta = True
-plot_deltacdf = True
-plot_nobias = True
-plot_wtf = True
+plot_delta = False
+plot_deltacdf = False
+plot_nobias = False
+plot_wtf = False
 
 #figsize2x1 = (12, 5.5)
 #figsize2x2 = (12, 11)
@@ -248,6 +248,7 @@ if pdf:
 tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
 posterior = np.load(posteriorFile)
 mean = posterior['mean']
+sigma = np.sqrt(posterior['var'])
 positive = (tgas['parallax'] > 0.) & (mean > 0.)
 ind = np.random.randint(0, len(tgas[positive]), nsubsamples)
 
@@ -394,7 +395,7 @@ if plot_examples:
 
     #plot likelihood and posterior in each axes
     for iteration in np.arange(20, 40):
-        fig, ax = makeFigureInstance(x=3, y=2) #, figsize=figsize3x2)
+        fig, ax = makeFigureInstance(x=3, y=2, hspace=0.75, figureSize=(2,2)) #, figsize=figsize3x2)
         #fig, ax = plt.subplots(2, 3, figsize=figsize3x2)
         #ax = ax.flatten()
         #fig.subplots_adjust(left=0.1, right=0.9,
@@ -411,16 +412,23 @@ if plot_examples:
         for i in range(np.max(digit)):
             currentInd = np.where((digit == i))[0]
             index = currentInd[np.random.randint(0, high=len(currentInd))]
-            ax[0].scatter(color[index], absMag_dust[index], c=dataColor)
+            ax[0].scatter(color[index], absMag_dust[index], c=dataColor, s=20)
[…]

### 2017-06-13 paper/paper.tex
dwh put in distance prior note
updated example plot
some of bovys comments

@@
 It is possible to build a stellar photometric model from the data themselves,
-because there are stars at all luminosities with useful parallax information.
+because there are stars at a wide range of luminosities with useful parallax information, and this
+range of stellar luminosities with useful parallax information will grow with future \gaia\ data releases.
 One challenge is that different stars are observed at different levels of parallax precision,
@@
 They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior.
-In contrast, more certain measurements tend to lie closer to their measured value.
-The blue points are better than the black for doing stuff.}
+In contrast, more certain measurements tend to lie closer to their measured value.}
 \label{fig:toy}
@@
   higher signal-to-noise.
-  In detail, this assumption is wrong, (for example, distant halo stars are different than local
-  disk stars), but the stationarity assumption is fairly weak:
+  In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
+  and also will be different than local
+  disk stars, in age and metallicity. The stationarity assumption is fairly weak:
   The requirement is that, locally in the \cmd, there is
@@
   work.
-\item[selection] We have no model for the selection function of \tgas,
+\item[selection] We have no model for the selection function of \tgas (but see \cite{bovy17}),
   nor selection volumes for any kinds of stars.
@@
   properties of the stars in the Milky Way (nor any volume-limited component or region
-  thereof).
+  thereof). For this reason, our \cmd\ prior might not be appropriate to use with other surveys. 
 \item[big data] We assume that we have large numbers of stars, large enough that
@@
 
+DWH: write something about how our prior is a prior on distance
+
 \subsection{Back to the Inference}
@@
 \centering
-  \includegraphics[width=\textwidth]{posterior.png}
+  \includegraphics[width=\textwidth]{posterior.pdf}
 \caption{Five example posterior PDFs over parallax using our method.

### 2017-06-13 paper/paper.tex
the rest of bovy comments

@@
 where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
-The transformation of the dust corrected absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
+The transformation of the (dust corrected) absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
@@
 To infer $\varpi_{\true}$ once the \cmd\ prior is constructed, we first do the inference in the 2D space of the \cmd\ prior $p(\truth \given \alpha)$, using Equation \ref{eq:posterior}.
-We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
+We then project the 2D posterior into the 1D $\varpi_{\true}$ space by multiply by the Jacobian to calculate the posterior PDF over $\varpi_{\true}$
 
 \begin{eqnarray}
-p(\truth_0 \given \data, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{\alpha}) \, f(\varpi_{\true}) \\
+p(\varpi_{\true} \given \data, \hat{\alpha}) \, &\propto& \, \left|\frac{\dd\truth}{\dd\varpi}\right| \, p(\truth \given \data, \hat{\alpha}) \, f(\varpi_{\true}) \\
+%p(\truth_0 \given \data, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{\alpha}) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
@@
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$, shown as the dotted line at $y=0.5$.
-And $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, or have a fraction decrease in variance by a factor of $>4$, shown as the other dotted line at $y=0.14$.
+The precision of our posterior PDFs is greater than or equal to the precision of the \gaia\ catalogue for $87\%$ of the stars, shown as the dotted line at $x=0$,
+and $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, which corresponds to a fraction decrease in variance by a factor of $>4$, shown as the dotted line at $y=0.14$.
 
@@
 The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
-The posterior PDFs are much more clustered and sharply peaked around the ``true" distance to M67 than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
+The posterior PDFs are much more narrow and clustered than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
 \label{fig:m67}
@@
 cluster are much more clustered than likelihood-based distance
-estimates.
+estimates, and they cluster around a sensible value.
+Accuracy is demonstrated by this test because each star in the cluster is
+subject to a different, unique distance prior (see \figurename~\ref{fig:posterior})
+and yet the posterior PDFs are all pulled to the same parallax value.
 
@@
   \includegraphics[width=\textwidth]{comparePrior.pdf}
-\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and the sqrt of the variance as the uncertainty.
+\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and the square root of the variance as the uncertainty.
 Left panel shows the posterior PDF using our \cmd\ prior, right panel shows the posterior PDFs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
@@
 It is a fairly weak prior.
-It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
 \figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
@@
 The lower main sequence is very narrow while the upper main sequence
-is much wider. This may be an age variation on the main sequence mixed
+is much wider. This may be an age and metallicity variation on the main sequence mixed
 with a Malmquist bias of seeing a much larger volume of upper main
@@
   drawn from the same distribution within some feature or patch of the
-  \cmd\ space. Thus, this has a small effect on our results, and could be improved by modelling the noise distribution simultaneously with the \cmd.
+  \cmd\ space. Thus, this has a small effect on our results, and could be improved by modeling the noise distribution simultaneously with the \cmd.
 \item[selection] The biggest limitation of our method for generating
@@
   inverse-selection-volume corrections.  This model is a model of the
-  \tgas\ Catalog (or really the \tgas--\tmass\ intersection), not of
+  \tgas\ Catalog (or really the \tgas--\tmass--\psone\ intersection), not of
   the stars in the Milky Way, nor any volume-limited subsample of the

### 2017-06-14 paper/paper.tex
some of my corrections

@@
 
-%\author{Jo Bovy}
-%\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
-%\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
+\author{Jo Bovy}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
+\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
 
@@
 This model contains no knowledge of stellar astrophysics or the
-Milky Way, but is precise because it is a generative model fit to many noisy
+Milky Way, but is precise because it is a generative-model fit to many noisy
 parallax measurements under an assumption of stationarity.
@@
 precise stars, and then calibrate spectrophotometric models.
-These spectrophotometric models, in turn, along with \gaia's on-board
+These spectrophotometric models, along with \gaia's on-board
 low-resolution $BP/RP$ spectrophotometry, are used to provide
@@
 The full stack required for these parallax inferences is complex.
-It involves modeling the stars, as well as the dust in the Milky Way,
+It involves modeling the stars, as well as the dust, in the Milky Way,
 and the response of the telescope itself.
@@
 However, the \gaia\ data set is ideal for thinking about purely statistical
-models of stars:
+models of stars.
 The different stars observed by \gaia\ are measured at very different
-signal-to-noise ratios.
-A data-driven model of the stellar color--magnitude distribution can capitalize
+signal-to-noise ratios,
+and a data-driven model of the stellar color--magnitude distribution can capitalize
 on the precise information coming from the high signal-to-noise stars
-and deliver valuable information about the low signal-to-noise stars,
-which in turn will deliver precise parallax (or distance or absolute magnitude) inferences.
+and deliver valuable information about the low signal-to-noise stars.
+This information can then deliver precise parallax (or distance or absolute magnitude) inferences.
 
@@
 radiative transfer, nuclear reactions, and atomic and molecular transitions.
-Because there are small issues with each of these components, the models are inaccurate in detail, and
-the physical models therefore produce parallax (and hence distance) estimates that are biased.
+Because there are small issues with each of these components, the physical models are inaccurate in detail, and
+therefore produce parallax (and distance) estimates that are biased.
 In addition, they build a long list of physical assumptions (about
@@
 
-We then use the novel \xd\ output---the deconvolved \cmd---as
-a prior for use in inference of individual-star parallaxes.
+We then use the \xd\ output---the deconvolved \cmd---as
+a prior for use in inference of individual stellar parallaxes.
 These inferences, one per star, provide much more precise parallax, distance,
@@
 about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
-These papers use the \gaia\ measurement to construct a parallax likelihood,
+These papers use a \gaia\ measurement to construct a parallax likelihood,
 and combine this with a distance prior.
@@
 \centering
-\includegraphics[width=\textwidth]{deltaParallax.pdf}
-\caption{No bias in posterior pdf parallax expectation values.}
-\label{fig:delta}
-\end{figure}
-
-\begin{figure}
-\centering
 \includegraphics[width=0.75\textwidth]{deltaCDF.pdf}
@@
 \label{fig:deltaCDF}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{deltaParallax.pdf}
+\caption{The difference between the expectation value of our parallax posteriors and the \gaia\ parallax measurement as a function of the \gaia\ uncertainty.
+The points are colored by the fractional change in the variance of the posterior PDF compared with the \gaia\ likelihood, the same as Figure \ref{fig:delta}.
+The points are evenly distributed around $y=0$ showing that there is little bias in our inferred parallax PDFs.
+The distribution gets wider with larger \gaia\ uncertainties due to those lower signal to noise measurements being more influenced by the prior.
+They also tend to have the largest fractional change in their variance (shown as darker points)}
+\label{fig:deltaParallax}
 \end{figure}
@@
   Ana Bonaca (Harvard),
-  Jo Bovy (Toronto),
   Andy Casey (Monash),

### 2017-06-14 paper/paper.tex
some of my corrections

@@
 We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the $y$ direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
-Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
-The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel.
-Similarly, noisy samples from this distribution are shown in the upper-right panel.
+Here the thickness of the toy distribution mimics the true thickness, or intrinsic scatter, of the \cmd\ due to age, mass, and metallicity.
+The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel in red.
+Similarly, noisy samples from this distribution are shown in the upper-right panel in grey and black.
 Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noisy distribution.
@@
 The MMLE gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the green lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
@@
 \end{eqnarray}
-In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs (visualized as the expectation value and $1\sigma$ uncertainties) of the true values $y_{\true,n}$ for the 5 highlighted black points from the upper-right panel.
+In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs (visualized as the expectation value and $1\sigma$ uncertainties) of the true values $y_{\true,n}$ for the 5 highlighted (black) points from the upper-right panel.
 The red points are the associated true $y$ values.
@@
 modes (and, with error bars, the variances) of the posterior PDFs.
-When an original data point has a large variance, the posterior ends
+When a measured value has a large variance, the posterior ends
 up close to the prior, which has a mode at the center line; the
@@
 We can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
+For the remainder of the paper, we will also maintain the color scheme presented here (where possible): truth will be red, data will be black, prior information will be green, and posterior information will be blue.
 
@@
   In detail, this assumption is wrong. For example, more distant stars will have lower signal to noise ratio measurements
-  and also will be different than local
-  disk stars, in age and metallicity. The stationarity assumption is fairly weak:
-  The requirement is that, locally in the \cmd, there is
+  and will also be different than local
+  disk stars in age and metallicity. However, the stationarity assumption is fairly weak:
+  the requirement is that, locally in the \cmd, there is
   support among the high signal-to-noise stars for the types
@@
   work.
-\item[selection] We have no model for the selection function of \tgas (but see \cite{bovy17}),
+\item[selection] We have no model for the selection function of \tgas\ (but see \cite{bovy17}),
   nor selection volumes for any kinds of stars.
@@
   (the \texttt{parallax{\_}error} entry in the \gaia\ catalog), are correct.
-  Here we are making use of the fact that, for
-  de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
-In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than than if we over-estimated the noise.
+  Here we are making use of the fact that when
+  de-noising data, under estimating measurement noise is conservative, and leads to under-deconvolution.
+In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise.
 \item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
@@
 Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude.
-We take $J_n^{\corr}$ as a point estimate and only propagate the parallax uncertainty.
+In detail, we take $J_n^{\corr}$ as a point estimate and only propagate the parallax uncertainty.
 
@@
 The left panel shows the point estimates of the color and absolute magnitude,
-using the point estimate of the parallax from \gaia, and the $1\sigma$ and
-$2\sigma$ contours of the distribution.
+using the point estimate of the parallax from \gaia. The $1\sigma$ and
+$2\sigma$ contours of the distribution are shown in grey.
 The right panel shows a subset of the data with its associated error bars, calculated via subsampling.

### 2017-06-14 paper/paper.tex
some of my corrections

@@
 $2\sigma$ contours of the distribution are shown in grey.
-The right panel shows a subset of the data with its associated error bars, calculated via subsampling.
+The right panel shows a subset of the data with the associated error bar for each star.
 The uncertainties in the colors are fairly well behaved, but the large
@@
 prior, shown in \figurename~\ref{fig:prior}.
-The left panel shows 1.3 million samples from the prior distribution, and the
+The left panel shows 1.4 million samples from the prior distribution, and the
 right panel shows $1\sigma$ contours for each component in the underlying
@@
 Figure~\ref{fig:posterior} shows the various probability distributions of our inference for a few randomly-chosen objects: the likelihood (black), prior (green), and posterior PDF (blue) on the true parallaxes.
-For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.
+For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ likelihood.
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$.
@@
 Figure~\ref{fig:posteriorCMD} shows the distribution of the expectation values of the posterior parallax PDFs projected back onto the \cmd\ space.
-The left panel shows the distribution for all 1.3 million stars in our dataset, and the right panel shows a subsample of the dataset but includes the $1\sigma$ uncertainties on the posterior PDFs.
+The left panel shows the distribution for all 1.4 million stars in our dataset, and the right panel shows a subsample of the dataset and includes the $1\sigma$ uncertainties on the posterior PDFs.
 Compared with the raw data shown in Figure~\ref{fig:data}, it is clear the precision of the posterior PDFs is greater, especially for the red giant branch stars.
@@
 This change in precision is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes in the variance.
-The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours over plotted.
+The left panel shows the natural log of the fractional change in variance as a function of $(J-K_s)^C$ color, with the $1\sigma$ and $2\sigma$ contours of the distribution over plotted.
 The points are colored by the natural log of their fractional change in variance (the $y$ axis of the left panel) in both panels to help guide your eye for the right panel.
@@
 
-Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog.
+Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog (the y-axis of Figure~\ref{fig:delta}).
 This more directly shows what was alluded to in Figure~\ref{fig:posterior}.
@@
 and $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, which corresponds to a fraction decrease in variance by a factor of $>4$, shown as the dotted line at $y=0.14$.
+
+LMA: describe Figure \ref{fig:deltaParallax}
 
@@
 As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
-
 M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
@@
 \figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
-The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
-The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not at all obvious.
-The posterior PDFs, shown in the bottom row, are more sharply peaked showing an increase in precision.
+The vertical red lines bracket the previously measured parallaxes of, or distances to, the cluster.
+The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
+The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
 The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
@@
 \figurename~\ref{fig:comparePrior} shows a comparison of the EDSD prior with our \cmd\ prior.
-Here the parallax posterior PDF expectation value and square root of the variance is converted into an absolute magnitude and visualized as the \cmd.
-The left panel visualizes results using our \cmd\ prior, and the right panel visualizes results using the EDSD prior.
+Here the parallax posterior PDF expectation value, and $1\sigma$ uncertainty, for a subset of stars in our dataset is converted into an absolute magnitude and visualized as the \cmd.
+The left panel visualizes results using the EDSD prior, and the right panel visualizes results using our \cmd\ prior.
 The variance for the \cmd\ prior is significantly smaller than the variance for the EDSD prior.
-In general, it is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
-It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
+In general, EDSD is a fairly broad prior so the median \gaia\ star has a posterior variance equal to the likelihood variance.
+EDSD has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
 
@@
   Milky Way.  Importantly, because we used \emph{all} of the stars in
-  the \tgas--\tmass\ intersection, the model \emph{is} representative of
+  the \tgas--\tmass--\psone\ intersection, the model \emph{is} representative of
   the full intersection, even the distant parts, where all the
@@
   parallaxes.  However, our use of the entire
-  \tgas--\tmass\ intersection without any selection function restricts
+  \tgas--\tmass--\psone\ intersection without any selection function restricts
   the use of our \cmd\ model as a prior to inferences within that same
-  \tgas--\tmass\ intersection.  It would be biasing to use this same
+  \tgas--\tmass--\psone\ intersection.  It would be biasing to use this same
   prior to infer distances for the full billion-star catalog released
   in \gaia\ DR1 alongside \tgas, or for stars fainter or brighter or
-  bluer than those included in the \tgas--\tmass\ intersection.
+  bluer than those included in the \tgas--\tmass--\psone\ intersection.
   A similar argument can be made for \textit{internal} selection effects, such as inhomogeneities of the depth and noise distributions, which we have neglected.
-  Those inhomogeneities are unlikely to affect the \cmd results; we effectively model the average distribution of the \tgas--\tmass\ intersection.
+  Those inhomogeneities are unlikely to affect the \cmd\ results; we effectively model the average distribution of the \tgas--\tmass--\psone\ intersection.
   By modeling the selection function in detail, one could improve the quality of the inferred \cmd, its effect on individual objects, and its connection to physical models of the Milky Way.
@@
   after some experimentation as being able to capture features but still
-  be easily optimized. We performed some experiments
+  easily optimized. We performed some experiments
   that suggest that larger $K$ would lead to better models, even in the
-  conservative sense of cross-validation score. Choosing an optimized or
+  conservative sense of a cross-validation score. Choosing an optimized or
   optimal number for this mixture is a natural extension of this work.

### 2017-06-14 paper/paper.tex
add bovy orcid
adding citations
abstract tweak

@@
 
-\author{Jo Bovy}
+\author[0000-0001-6855-442X]{Jo Bovy}
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA\hspace*{-1cm}}
@@
 This model contains no knowledge of stellar astrophysics or the
-Milky Way, but is precise because it is a generative-model fit to many noisy
-parallax measurements under an assumption of stationarity.
+Milky Way, but is precise because it generates accurately a large number of noisy
+parallax measurements under an assumption of stationarity; that is,
+it combines the information from many stars.
 We use the Extreme Deconvolution (\xd) algorithm---an Empirical Bayes approximation to a full
@@
 the photometry (colors and apparent magnitude or magnitudes) of a
-star.
+star. This idea was pioneered by \cite{juric08} using the Sloan Digital Sky Survey (\citealt{sdss}).
 Fundamentally, photometric parallaxes capitalize on the fact that the

### 2017-06-14 paper/paper.tex
slight ammendment to abstract

@@
 This model contains no knowledge of stellar astrophysics or the
-Milky Way, but is precise because it generates accurately a large number of noisy
+Milky Way, but is precise because it accurately generates a large number of noisy
 parallax measurements under an assumption of stationarity; that is,
-it combines the information from many stars.
+it is capable of combining the information from many stars.
 We use the Extreme Deconvolution (\xd) algorithm---an Empirical Bayes approximation to a full

### 2017-06-14 paper/paper.tex
intro tweak

--- 2017-06-14 paper/paper.tex
@@
 In general, data-driven models contain fewer assumptions than
-physics-driven models, but they are also usually less interpretable.
-However, the \gaia\ data set is ideal for thinking about purely statistical
+physics-driven models, but they are also usually less interpretable;
+the model presented here has these same properties.
+The \gaia\ data set is ideal for thinking about these kinds of purely statistical
 models of stars.
@@
 and deliver valuable information about the low signal-to-noise stars.
-This information can then deliver precise parallax (or distance or absolute magnitude) inferences.
+This information can then deliver precise parallax (or distance or absolute magnitude) inferences for all of the stars.
 

### 2017-06-14 paper/paper.tex
bias

@@
 
-LMA: describe Figure \ref{fig:deltaParallax}
+We have shown that the posterior parallax estimates are more precise
+than the likelihoods, both on average and especially for low
+signal-to-noise stars.
+It is also interesting to ask whether our data-driven prior leads to
+biased inferences.
+In Figure~\ref{fig:deltaParallax}, we show the shifts of the parallax
+estimates from likelihood \tgas\ catalog entry to posterior
+expectation.
+This shows that while many parallax estimates change substantially,
+they do not change on average, even at low signal-to-noise ratios.
+Our conclusion is that the prior developed here does not introduce additional
+bias into the \tgas\ parallax measurements.
 
@@
 The points are colored by the fractional change in the variance of the posterior PDF compared with the \gaia\ likelihood, the same as Figure \ref{fig:delta}.
-The points are evenly distributed around $y=0$ showing that there is little bias in our inferred parallax PDFs.
+The points are evenly distributed around $y=0$ showing that there is little bias introduced by inferred parallax PDFs.
 The distribution gets wider with larger \gaia\ uncertainties due to those lower signal to noise measurements being more influenced by the prior.

### 2017-06-14 paper/paper.tex
figuresize
apsis

@@
 Only a small fraction (but large number) of these
-measurements will be both precise and purely astrometric.
+measurements will be precise and purely astrometric.
 \gaia\ will use astrometric parallaxes to determine the distances of the more
-precise stars, and then calibrate spectrophotometric models.
+precise stars, and then calibrate spectrophotometric models (\citealt{apsis}).
 These spectrophotometric models, along with \gaia's on-board
-low-resolution $BP/RP$ spectrophotometry, are used to provide
-parallax estimates to less precise stars.
+low-resolution $BP/RP$ spectrophotometry, can be used to provide
+more precise parallax estimates for stars with poor astrometric parallax data.
 The full stack required for these parallax inferences is complex.
@@
 \centering
-\includegraphics[width=0.75\textwidth]{deltaCDF.pdf}
+\includegraphics[width=0.5\textwidth]{deltaCDF.pdf}
 \caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.
@@
 \centering
-\includegraphics[width=\textwidth]{deltaParallax.pdf}
+\includegraphics[width=0.5\textwidth]{deltaParallax.pdf}
 \caption{The difference between the expectation value of our parallax posteriors and the \gaia\ parallax measurement as a function of the \gaia\ uncertainty.

### 2017-06-14 plots.py
bibtex

@@ -182,9 +182,9 @@ plot_m67 = False
 plot_compare = False
 plot_expectation = False
 plot_examples = True
-plot_delta = False
+plot_delta = True
 plot_deltacdf = False
-plot_nobias = False
+plot_nobias = True
 plot_wtf = False
 
 #figsize2x1 = (12, 5.5)
@@ -418,7 +418,7 @@ if plot_examples:
             #print len(color), len(absMagKinda_dust), len(color_err), len(absMagKinda_dust_err), len(apparentMagnitude)
             likeParallax, priorParallax, posteriorParallax = likePriorPost(color[index], absMagKinda_dust[index], color_err[index], absMagKinda_dust_err[index], apparentMagnitude[index], xdgmm, ndim=2, nPosteriorPoints=1000, projectedDimension=1)
 
-            l1, = ax[i+1].plot(xparallaxMAS, likeParallax*np.max(posteriorParallax)/np.max(likeParallax), lw=2, color=dataColor)
+            l1, = ax[i+1].plot(xparallaxMAS, likeParallax*np.max(posteriorParallax)/np.max(likeParallax), lw=1, color=dataColor, zorder=100)
             l2, = ax[i+1].plot(xparallaxMAS, priorParallax*np.max(posteriorParallax)/np.max(priorParallax), lw=0.5, color=priorColor)
             l3, = ax[i+1].plot(xparallaxMAS, posteriorParallax, lw=2, color=posteriorColor)
             maxInd = posteriorParallax == np.max(posteriorParallax)
@@ -468,7 +468,7 @@ if plot_delta:
     norm = plt.matplotlib.colors.Normalize(vmin=-1.5, vmax=1)
     cmap = 'inferno'
     ax[0].scatter(x[notnans], y[notnans], c=y[notnans], s=1, lw=0, alpha=0.05, norm=norm, cmap=cmap, rasterized=True)
-    corner.hist2d(x[notnans], y[notnans], bins=200, ax=ax[0], levels=levels, no_fill_contours=True, plot_density=False, plot_data=False, color=contourColor, rasterized=True)
+    #corner.hist2d(x[notnans], y[notnans], bins=200, ax=ax[0], levels=levels, no_fill_contours=True, plot_density=False, plot_data=False, color=contourColor, rasterized=True)
     ax[0].set_xlabel(xlabel_cmd)
     ax[0].set_ylim(-6, 2)
[…]

### 2017-06-14 paper/paper.tex
final edits

@@
 The black points are 5 randomly chosen data to highlight the method.
-In the lower-left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
+In the lower-left panel, the green lines represent the noise deconvolved, best estimate for the underlying distribution.
 The blue points represent the expectation value of the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
@@
   correctly known variances. In detail, we assume that the parallax uncertainty estimates,
-  (the \texttt{parallax{\_}error} entry in the \gaia\ catalog), are correct.
+  (the \texttt{parallax{\_}error} entry in the \tgas\ catalog), are correct.
   Here we are making use of the fact that when
@@
 Equation~\ref{eq:bayes} does not fully specify this procedure.
-Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
-Specifically, we use the $J$ and $K_s$ band photometry from \tmass, as well as the parallax from \gaia.
+Because our prior comes from the \cmd, our inference will include not only the parallax from \tgas, but also the photometry of the stars.
+Specifically, we use the $J$ and $K_s$ band photometry from \tmass, as well as the parallax from \tgas.
 Our true, latent values are the $(J-K_s)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$.
-The observed data is the dust corrected color $(J-K_s)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.
+The observed data is the dust corrected color $(J-K_s)^{\corr}_n$, the \tgas\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.
 
@@
   \includegraphics[width=\textwidth]{data.pdf}
-\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K_s$ color, the $J$ band apparent magnitude, and the \gaia\ parallax.
-The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution.
+\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K_s$ color, the $J$ band apparent magnitude, and the \tgas\ parallax.
+The grey lines represent the $1\sigma$ and $2\sigma$ contours for the distribution.
 To give a sense of the uncertainties, in the right panel we subsample the dataset and include the $1\sigma$ uncertainties.
@@
 The left panel shows the point estimates of the color and absolute magnitude,
-using the point estimate of the parallax from \gaia. The $1\sigma$ and
+using the point estimate of the parallax from \tgas. The $1\sigma$ and
 $2\sigma$ contours of the distribution are shown in grey.
@@
   \includegraphics[width=\textwidth]{prior.pdf}
-\caption{The \cmd\ prior, modeled as a Gaussian mixture inferred by
+\caption{The \cmd\ prior, modeled as a Gaussian mixture, inferred by
   running XD on all of the data. The left panel shows a sampling of
-  the prior, with the blue lines showing the $1\sigma$ and $2\sigma$
+  the prior, with the black lines showing the $1\sigma$ and $2\sigma$
   contours. The right panel shows the 1-$\sigma$ contours of the
@@
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$.
-Examples of these slight increases in precision are visualized in Examples 1 through 4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
+Examples of these slight increases in precision are visualized in Examples 1, 3, and 4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
 However, $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$.
@@
 
-Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \gaia\ catalog (the y-axis of Figure~\ref{fig:delta}).
+Figure~\ref{fig:deltaCDF} shows the cumulative distribution function of the natural log of the fractional change in variance of the parallax posterior pdf relative to the \tgas\ catalog (the y-axis of Figure~\ref{fig:delta}).
 This more directly shows what was alluded to in Figure~\ref{fig:posterior}.
-For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance.
+For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \tgas\ variance.
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$, shown as the dotted line at $y=0.5$.
-The precision of our posterior PDFs is greater than or equal to the precision of the \gaia\ catalogue for $87\%$ of the stars, shown as the dotted line at $x=0$,
+The precision of our posterior PDFs is greater than or equal to the precision of the \tgas\ catalogue for $87\%$ of the stars, shown as the dotted line at $x=0$,
 and $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, which corresponds to a fraction decrease in variance by a factor of $>4$, shown as the dotted line at $y=0.14$.
@@
 Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1 through 4 are not very effected by the prior (green), where good data is good data.
-Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior PDF, transformed into an absolute magnitude, remains on the main sequence.
-Example 5 shows a significantly more precise posterior PDF compared with the likelihood, and looks very similar to the prior PDF.
-It also exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
+Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
+Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
+Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
 \label{fig:posterior}
@@
   \includegraphics[width=\textwidth]{posteriorCMD.pdf}
-\caption{The distribution of expectation values of the parallax posterior PDFs transformed into an absolute magnitude vs color, the \cmd, of the full data set is shown in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution shown in black.
+\caption{Posterior Expectation Values: The distribution of expectation values of the parallax posterior PDFs transformed into an absolute magnitude vs color, the \cmd, of the full data set is shown in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution shown in black.
 The right panel shows a subsample of data set, plotting the expectation value and $1\sigma$ uncertainty of posterior PDF projected onto the \cmd.
-Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}
+Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}.
 The visual of the prior is a sampling of the Gaussians, but this is the expectation
@@
 \includegraphics[width=\textwidth]{delta.pdf}
-\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color.
-The coloring of the points in both panels is by the fractional change in variance to guide the coloring of the right panel.
-The blue lines are the $1\sigma$ and $2\sigma$ contours.
+\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior PDF relative to the likelihood as a function of the stellar color.
+The coloring of the points in both panels is the fractional change in variance (the y-axis in the left panel) to guide the interpretation of the right panel.
 The median fractional change is 1.2.
-Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K_s)^{\corr}$ of about 0.6.
+Some posteriors have significantly smaller variances shown in black, but some increase in variance, shown as the yellow bump at $(J-K_s)^{\corr}$ of about 0.6.
 The right panel shows the posterior point estimates on the CMD colored by fraction change in variance.
@@
 \includegraphics[width=0.5\textwidth]{deltaCDF.pdf}
-\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \gaia\ catalogue.
-Negative numbers represent stars which have a posterior that is narrower than the \gaia\ catalogue, which is true for a large fraction of the stars.
-We've drawn two dashed lines.
+\caption{Fractional Change in Variance: The cumulative distribution function of the natural log of the fractional variance change when comparing our posterior to the \tgas\ catalogue.
+Negative numbers represent stars which have a posterior that is narrower than the \tgas\ catalogue, which is true for a large fraction of the stars.
+We've drawn three dashed lines.
 One represents the median star $(y=0.5)$ which has a fractional variation of $x=-0.3=ln(1/1.2^2)$, corresponding to an increase in precision by a factor of 1.2.
-Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.}
+Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.
+The final line, $x=0$, represents the fraction of stars with an increased precision, $87\%$.}
 \label{fig:deltaCDF}
@@
 \includegraphics[width=0.5\textwidth]{deltaParallax.pdf}
-\caption{The difference between the expectation value of our parallax posteriors and the \gaia\ parallax measurement as a function of the \gaia\ uncertainty.
-The points are colored by the fractional change in the variance of the posterior PDF compared with the \gaia\ likelihood, the same as Figure \ref{fig:delta}.
+\caption{Minimal bias introduced: The difference between the expectation value of our parallax posteriors and the \tgas\ parallax measurement as a function of the \tgas\ uncertainty.
+The points are colored by the fractional change in the variance of the posterior PDF compared with the \tgas\ likelihood, the same as Figure \ref{fig:delta}.
 The points are evenly distributed around $y=0$ showing that there is little bias introduced by inferred parallax PDFs.
-The distribution gets wider with larger \gaia\ uncertainties due to those lower signal to noise measurements being more influenced by the prior.
+The distribution gets wider with larger \tgas\ uncertainties due to those lower signal to noise measurements being more influenced by the prior.
 They also tend to have the largest fractional change in their variance (shown as darker points)}
@@
 star in the intersection of \tgas, \tmass, and the dust map.
-Associated with this paper is an electronic table containing our results, and everything
+Associated with this paper is an electronic table containing our results\footnote{\url{http://voms.simonsfoundation.org:50013/8kM7XXPCJleK2M02B9E7YIYmvu5l2rh/ServedFiles/}}, and everything
 that is necessary to use them responsibly.
-It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
+It includes the \tgas\ identifier, sky position, and dust-corrected
 photometry.
@@
 
-With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
+With \tgas\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
 M67 is therefore a great test case for comparing parallax inferences with or without a prior.
@@
 We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
-To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \gaia\ within this window on the sky.
+To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
 Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
@@
 Our prior, visualized as a mixture of Gaussians, is shown in green.
-The upper right panel shows the \gaia\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
+The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
 Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
@@
 The lower right panel shows the projection of the 2D posterior PDF onto parallax space.
-The posterior PDFs are much more narrow and clustered than the \gaia\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
+The posterior PDFs are much more narrow and clustered than the \tgas\ likelihoods, showing that our prior generates posterior PDFs that are accurate and more precise than \gaia\ alone.}
 \label{fig:m67}
@@
   \includegraphics[width=\textwidth]{comparePrior.pdf}
-\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and the square root of the variance as the uncertainty.
-Left panel shows the posterior PDF using our \cmd\ prior, right panel shows the posterior PDFs when using the exponentially declining spatial density, a common space-density prior used when inferring distances.
+\caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and $1\sigma$ uncertainty.
+The left panel shows the posterior PDF using the exponentially declining spatial density, a common space-density prior used when inferring distances.
+The right panel shows the posterior PDF using our \cmd\ prior.
 The scale of the exponentially declining spatial density prior is $1.35~{\rm kpc}$, the optimal value found in \cite{astraatmadja16a}.
@@
 Here we have noted the features previously found in the \cmd\ that are also highlighted nicely by the \gaia\ + \tmass\ data.
-The upper and lower main sequence, as well as the main sequence turn off.
+The upper and lower main sequence, as well as the main sequence turn off are easily seen.
 There is a plausible binary sequence of stars, as well as a prominent red clump, and straight and narrow red giant branch. }

### 2017-06-14 plots.py

@@ -181,7 +181,7 @@ plot_prior = False
 plot_m67 = False
 plot_compare = False
 plot_expectation = False
-plot_examples = True
+plot_examples = False
 plot_delta = True
 plot_deltacdf = False
 plot_nobias = True
@@ -525,7 +525,7 @@ if plot_nobias:
     ax.scatter(x[notnans], y[notnans], c=colorDeltaVar[notnans], s=1, lw=0, alpha=0.05, norm=norm, cmap=cmap, rasterized=True)
     #ax.scatter(x[good], y[good], c=sigma[good], s=1, lw=0, alpha=0.05, norm=norm, cmap=cmap, rasterized=True)
     #ax.scatter(x[good], y[good], c=np.sqrt(sigma[good]), s=1, rasterized=True, zorder=0, alpha=0.1, cmap=cmap, norm=norm)
-    ax.plot([0, 1.1], [0,0], 'k--')
+    ax.plot([0, 1.1], [0,0], 'k--', lw=1)
     ax.set_xlim(0.15, 1.05)
     ax.set_ylim(-2.5, 2.5)
     ylabel = r'$\mathrm{Posterior \, Expectation \, Value} - \varpi_n$'

### 2017-06-14 paper/paper.tex
last tweaks

@@
-% This project is part of the stellarTwins project.
+% This project is part of the photoParallax project.
 % Copyright 2017 the authors.
@@
 the model presented here has these same properties.
-The \gaia\ data set is ideal for thinking about these kinds of purely statistical
+Yet, the \gaia\ data set is ideal for thinking about these kinds of purely statistical
 models of stars.
@@
 p(\varpi_{\true} \given \data, \hat{\alpha}) \, &\propto& \, \left|\frac{\dd\truth}{\dd\varpi}\right| \, p(\truth \given \data, \hat{\alpha}) \, f(\varpi_{\true}) \\
-%p(\truth_0 \given \data, \hat{\alpha}) &\rightarrow& p(\varpi_{\true} \given y, \hat{\alpha}) \, f(\varpi_{\true}) \\
 \mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
@@
 Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$, shown as the dotted line at $y=0.5$.
-The precision of our posterior PDFs is greater than or equal to the precision of the \tgas\ catalogue for $87\%$ of the stars, shown as the dotted line at $x=0$,
+The precision of our posterior PDFs is greater than or equal to the precision of the \tgas\ catalogue for $83\%$ of the stars, shown as the dotted line at $x=0$,
 and $14\%$ of our parallax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$, which corresponds to a fraction decrease in variance by a factor of $>4$, shown as the dotted line at $y=0.14$.
@@
 Another represents the fraction of stars $(y=0.14)$ which have a fractional variance greater than $x=-1.4=ln(1/2^2)$, corresponding to an increase in precision by a factor of 2.
-The final line, $x=0$, represents the fraction of stars with an increased precision, $87\%$.}
+The final line, $x=0$, represents the fraction of stars with an increased precision, $83\%$.}
 \label{fig:deltaCDF}

### 2017-06-14 plots.py

@@ -182,10 +182,10 @@ plot_m67 = False
 plot_compare = False
 plot_expectation = False
 plot_examples = False
-plot_delta = True
+plot_delta = False
 plot_deltacdf = False
-plot_nobias = True
-plot_wtf = False
+plot_nobias = False
+plot_wtf = True
 
 #figsize2x1 = (12, 5.5)
 #figsize2x2 = (12, 11)
@@ -347,14 +347,14 @@ if plot_compare:
     #ax2 = fig.add_subplot(122)
     #ax = [ax1, ax2]
     titles =  ['Exp Dec Sp \nDen Prior', 'CMD Prior']
-    for i, file in enumerate(['posteriorSimple.npz', posteriorFile]):
-        data = np.load(file)
-        posterior = data['posterior']
-        sigma = np.sqrt(data['var'])
-        mean = data['mean']
-        absMag = testXD.absMagKinda2absMag(mean[positive]*10.**(0.2*apparentMagnitude))
-        absMag_err = absMagError(mean[positive], sigma[positive], apparentMagnitude, absMag)
-        #ax[i].scatter(color[ind], absMag[ind], c=posteriorColor, s=1, lw=0, alpha=alpha, zorder=0)
+for i, file in enumerate(['posteriorSimple.npz', posteriorFile]):
+    data = np.load(file)
+    posterior = data['posterior']
+    sigma = np.sqrt(data['var'])
+    mean = data['mean']
+    absMag = testXD.absMagKinda2absMag(mean[positive]*10.**(0.2*apparentMagnitude))
+    absMag_err = absMagError(mean[positive], sigma[positive], apparentMagnitude, absMag)
+    if plot_compare:    #ax[i].scatter(color[ind], absMag[ind], c=posteriorColor, s=1, lw=0, alpha=alpha, zorder=0)
         ax[i].errorbar(color[ind], absMag[ind], xerr=color_err[ind], yerr=[absMag_err[0][ind], absMag_err[1][ind]], fmt="none", zorder=0, mew=0, ecolor=posteriorColor, alpha=0.5, elinewidth=0.5, color=posteriorColor)
         ax[i].set_xlim(xlim_cmd)
         ax[i].set_ylim(ylim_cmd[0], ylim_cmd[1]*1.1)
@@ -364,6 +364,7 @@ if plot_compare:
[…]

### 2018-02-16 paper/paper.tex
last tweaks
arxiver
first round of edits

@@
 
-%@archiver{posteriorCMD.pdf}
-
-\documentclass[modern]{aastex61}
+%@arxiver{posteriorCMD.pdf}
+
+\documentclass[trackchanges]{aastex61}
 \usepackage{amsmath}
@@
 \affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
+\affil{Alfred P. Sloan Fellow}
 
@@
 The transformation of the (dust corrected) absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
+\added{This transformation also allows our method to handle negative parallaxes from the \tgas\ dataset naturally.
+A negative parallax simply refers to a distant object that is affected by large errors, and therefore intrinsically bright, and our transformation places these objects in the bright portion of the \cmd.}
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
@@
 
-DWH: write something about how our prior is a prior on distance
-
 \subsection{Back to the Inference}
@@
 The match was done using a nearest-neighbor algorithm with a search radius of $4$ arcsec\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}.
+\added{There is also a \tgas\-\tmass\ cross-match as part of the official \gaia\ product, but this project began before that catalogue was available.}
 We also required that the stars lie within the

### 2018-02-20 paper/paper.tex
updated to improve paper plots

@@
 \begin{abstract}\noindent % trust me
-Converting a noisy parallax measurement into a posterior
+{\bf \textcolor{red}{Stephen Feeeeeeeneeeeeey is the best.}} Converting a noisy parallax measurement into a posterior
 belief over distance requires inference with a prior.

### 2018-02-20 comparePrior.py

@@ -111,23 +111,23 @@ def comparePrior():
     plt.tight_layout()
     fig.savefig('priorNgaussComparison.png')
 
-def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz', ax=None, tgas=None):
+def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz'):
 
-    if tgas is None: tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
+    tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
     dustFile = dustFilename(ngauss, quantile, iter, survey, dataFilename)
     data = np.load(dustFile)
     dust = data['ebv']
-    if ax is None: fig, ax = plt.subplots(figsize=(12,7))
+    fig, ax = plt.subplots(figsize=(12,7))
     norm = mpl.colors.PowerNorm(gamma=1/2.)#(vmin=-0.5, vmax=1)
-    im = ax.scatter(tgas['l'], tgas['b'], c=dust, lw=0, cmap='Greys', s=1, vmin=0, vmax=1, norm=norm, rasterized=True)
+    im = ax.scatter(tgas['l'], tgas['b'], c=dust, lw=0, cmap='Greys', s=1, vmin=0, vmax=1, norm=norm)
     ax.set_xlim(0, 360)
     ax.set_ylim(-90, 90)
-    ax.set_xlabel(r'$\mathscr{l}$ [deg]')
-    ax.set_ylabel(r'$\mathscr{b}$ [deg]')
+    ax.set_xlabel('$\mathscr{l}$ [deg]')
+    ax.set_ylabel('$\mathscr{b}$ [deg]')
     cb = plt.colorbar(im, ax=ax)
     cb.set_clim(-0.1, 1)
     cb.set_label(r'E($B-V$ )')
-    if ax is None: fig.savefig('dustViz.dQ' + str(quantile) + '.png')
+    fig.savefig('dustViz.dQ' + str(quantile) + '.png')
 
 def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz', sdss5=False, whatsThatFeature=False):
 
@@ -386,7 +386,7 @@ def examplePosterior(nexamples=100, postFile='posteriorSimple.npz', dustFile='du
     mean = data['mean']
     var = data['var']
[…]

### 2018-02-20 plots.py

@@ -52,7 +52,7 @@ def figsize_and_margins(plotsize,subplots=(1,1),**absolute_margins):
     #height * (top - bottom)
     #    = nrows * plot_height + (nrows - 1) * hspace * plot_height
     #solve for width and height, using absolute margins as necessary:
-    print nc, rmarg['wspace'], pw, amarg.get('lef', 0), amarg.get('right', 0), rmarg.get('right', 1), rmarg.get('left', 0)
+    print(nc, rmarg['wspace'], pw, amarg.get('lef', 0), amarg.get('right', 0), rmarg.get('right', 1), rmarg.get('left', 0))
     width  = float((nc + (nc - 1) * rmarg['wspace']) * pw        \
                    + amarg.get('left',0) + amarg.get('right',0)) \
              / (rmarg.get('right',1) - rmarg.get('left',0))
@@ -173,390 +173,469 @@ def likePriorPost(color, absMagKinda, color_err, absMagKinda_err, apparentMagnit
     likeParallax = gaussian(absMagKinda/10.**(0.2*apparentMagnitude), absMagKinda_err/10.**(0.2*apparentMagnitude), xparallaxMAS)
     return likeParallax, priorParallax, posteriorParallax
 
-#    for label, style in zip(['paper', 'talk'],['seaborn-paper', 'seaborn-talk']):
-pdf = True
-plot_data = False
-plot_dust = False
-plot_prior = False
-plot_m67 = False
-plot_compare = False
-plot_expectation = False
-plot_examples = False
-plot_delta = False
-plot_deltacdf = False
-plot_nobias = False
-plot_wtf = True
+def main():
+    #    for label, style in zip(['paper', 'talk'],['seaborn-paper', 'seaborn-talk']):
+    pdf = True
+    plot_data = False
+    plot_dust = False
+    plot_prior = False
+    plot_m67 = True
+    plot_compare = False
+    plot_expectation = False
+    plot_odd_examples = False
+    plot_examples = False
+    plot_delta = False
+    plot_deltacdf = False
+    plot_nobias = False
+    plot_wtf = False
+    plot_toy = False
 
-#figsize2x1 = (12, 5.5)
-#figsize2x2 = (12, 11)
-#figsize3x2 = (18, 11)
[…]

### 2018-02-20 stellarTwins.py

@@ -7,7 +7,6 @@ import itertools
 import matplotlib as mpl
 from astroML.density_estimation import XDGMM
 from astroML.plotting.tools import draw_ellipse
-from extreme_deconvolution import extreme_deconvolution as ed
 from matplotlib.patches import Ellipse
 import pdb
 from dustmaps.sfd import SFDQuery
@@ -39,11 +38,11 @@ def distMetric(sourceInd, matchedInd, apassMagnitudes, varMuMatched, p=False):
     absMagSigs = np.log(apassMagnitudes['e_vmag'][sourceInd]**2. + apassMagnitudes['e_vmag'][matchedInd]**2. + varMuMatched[sourceInd] + varMuMatched[matchedInd])
     totChiSq = colorChiSq + absMagChiSq + colorSigs + absMagSigs
     if p:
-        print 'the total chi2: ',totChiSq[0:5]
-        print 'the color chi2: ',colorChiSq[0:5]
-        print 'the color sigs: ', colorSigs[0:5]
-        print 'the absmag chi2:',absMagChiSq[0:5]
-        print 'the absmag sigs:',absMagSigs[0:5]
+        print( 'the total chi2: ',totChiSq[0:5])
+        print( 'the color chi2: ',colorChiSq[0:5])
+        print( 'the color sigs: ', colorSigs[0:5])
+        print( 'the absmag chi2:',absMagChiSq[0:5])
+        print( 'the absmag sigs:',absMagSigs[0:5])
 
     return totChiSq
 
@@ -381,9 +380,9 @@ def observationsCutMatched(SNthreshold=1., filename='cutMatchedArrays.npz'):
     raveInd = np.in1d(tgasRave['source_id'], tgasMatched['source_id'])
     raveMatched = tgasRave[raveInd]
 
-    print 'Number of tgas stars: ', len(tgas)
-    print 'Number of matched stars: ', np.sum(matched)
-    print 'Percent matched = ', 100 - (len(tgas) - np.sum(matched))/np.float(len(tgas))*100., '%'
+    print( 'Number of tgas stars: ', len(tgas))
+    print( 'Number of matched stars: ', np.sum(matched))
+    print( 'Percent matched = ', 100 - (len(tgas) - np.sum(matched))/np.float(len(tgas))*100., '%')
 
[…]

### 2018-02-20 testXD.py

@@ -4,7 +4,6 @@ import matplotlib.pyplot as plt
 import os
 import sys
 import stellarTwins as st
-from extreme_deconvolution import extreme_deconvolution as ed
 import astropy.units as units
 import numpy as np
 import pdb
@@ -29,7 +28,7 @@ def convert2gal(ra, dec):
     """
     return SkyCoord([ra, dec], unit=(units.hourangle, units.deg))
 
-def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963', twoMass=None):
+def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963'):
     """
     return the indices of tgas which fall within dl and db of l and b.
     default is M67
@@ -39,11 +38,6 @@ def m67indices(tgas, plot=False, dl=0.1, db=0.1, l='215.6960', b='31.8963', twoM
             (tgas['b'] > np.float(b) - db) & \
             (tgas['l'] < np.float(l) + dl) & \
             (tgas['l'] > np.float(l) - dl)
-
-    if twoMass is not None:
-        print twoMass['j_cmsig'][index], twoMass['k_cmsig'][index]
-        goodError = (twoMass['j_cmsig'] < 2.) & (twoMass['k_cmsig'] < 2.)
-        index = index & goodError
     if plot:
         fig, ax = plt.subplots()
         ax.scatter(tgas['l'][index], tgas['b'][index], alpha=0.5, lw=0)
@@ -91,7 +85,7 @@ def scalability(X, Xerr, numberOfStars = [1024, 2048, 4096, 8192]):
     totTime = np.zeros(4)
     for i, ns in enumerate(numberOfStars):
         totalTime, numStar = timing(X, Xerr, nstars=ns, ngauss=64)
-        print totalTime, numStar
+        print(totalTime, numStar)
         totTime[i] = totalTime
         plt.plot(numData, totTime)
         plt.savefig('timing64Gaussians.png')
@@ -234,7 +228,7 @@ def posterior2d(means, amps, covs, xbins, ybins, nperGauss=100000., plot=False):
     Z, xedges, yedges = np.histogram2d(samplesX, absMagKinda2absMag(samplesY), bins=[xbins, ybins], normed=True)
     return xedges, yedges, Z, samplesY
 
[…]

### 2018-02-20 toy/toy.py

@@ -42,49 +42,47 @@ def denoise_one_datum(xn, yn, sigman, m, b, t):
     return (yn * s2inv + (m * xn + b) * t2inv) / (s2inv + t2inv), \
         np.sqrt(1. / (s2inv + t2inv))
 
-def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5, dataColor='k', priorColor='g', posteriorColor='b', trueColor='r', posteriorMap = 'Blues', markersize=20):
+def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5, dataColor='k', priorColor='g', posteriorColor='b', trueColor='r', posteriorMap = 'Blues'):
 
     alpha_all = 0.05
     alpha_chosen = 1.0
     xlim = (-1, 1)
     ylim = (-5, 5)
-    markersize=markersize
+
     for ax in axes[1:-1]:
-        ax.errorbar(xns, yns, yerr=sigmans, fmt=None, color=dataColor, alpha=alpha_all ,mew=0)
-        ax.scatter(xns, yns, s=markersize, alpha=alpha_all, color=dataColor)
-        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="none", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
-    axes[1].scatter(xns[0:nexamples], yns[0:nexamples], s=markersize, color=dataColor, alpha=alpha_chosen)
-    axes[2].scatter(xns[0:nexamples], yns[0:nexamples], s=2.*markersize, color=dataColor, alpha=alpha_chosen)
+        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color=dataColor, alpha=alpha_all ,mew=0)
+        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
+
     xp = np.array(xlim)
     axes[0].plot(xp, mtrue*xp + btrue + ttrue, color=trueColor, linewidth=2, alpha=0.75, label=r'$y=m_{true}\,x+b_{true}\pm t$')
[…]

### 2018-03-04 paper/paper.tex
tweaks to the paper

@@
 These papers use a \gaia\ measurement to construct a parallax likelihood,
-and combine this with a distance prior.
+and combine this with a parallax prior.
 While this is sensible and correct, all present methods proposed along
 these lines build in informative (and known-to-be-wrong) assumptions
-about the line-of-sight distance distribution to \gaia\ stars.
+about the line-of-sight parallax distribution to \gaia\ stars.
 That is, they build in strong and informative assumptions about the Milky
@@
   effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
-  dust correction at its updated, inferred distance. Our assumption is that these corrections are
+  dust correction at its updated, inferred parallax. Our assumption is that these corrections are
   good enough, and uncertainties in these are not dominant, for neither
@@
 The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
-The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
+The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \added{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
 We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
@@
 therefore have dust corrections $> 2$~mag. Although the dust
-model is probabilistic, as well as our distance inference, to move
+model is probabilistic, as well as our parallax inference, to move
 forward, we break from a fully Bayesian framework in which we might
-infer the distance and the dust simultaneously (more about this in the
+infer the parallax and the dust simultaneously (more about this in the
 discussion). Instead, we correct for dust using a point estimate from
@@
 the dust values seem to converge after a few iterations. The
-iterative process allows for slight variations in distance and
+iterative process allows for slight variations in parallax and
 corresponding dust to become more consistent with the \cmd.
@@
   \includegraphics[width=\textwidth]{dust.pdf}
-\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates.
+\caption{The converged dust values at the 5\% parallax quantile in Galactic coordinates.
 This shows both the footprint of our analysis as well as the adopted extinction values.
@@
 \centering
+  \includegraphics[width=\textwidth]{posterior_odd.pdf}
+\caption{Five example posterior PDFs over parallax using our method for more odd objects.
+The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
+In blue, we show the posterior expectation value and $1 \sigma$ uncertainties.
+We chose example points that lie outside the well populated portions of the \cmd to illustrate how our method works for rare \tgas objects. 
+The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
+
+Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
+Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
+Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
+Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
+\label{fig:posterior}
+\end{figure}
+
+\begin{figure}
+\centering
   \includegraphics[width=\textwidth]{posteriorCMD.pdf}
@@
 
-We select sources from our catalog within a $1~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
+We select sources from our catalog within a $0.75~{\rm deg}^2$ window centered on the position of M67; the cluster is visible as an over-density of stars in the sky positions alone.
 To be clear, we do no prior selection on previously known members of M67, but include all stars observed by \tgas\ within this window on the sky.
@@
 \figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
-The vertical red lines bracket the previously measured parallaxes of, or distances to, the cluster.
+The vertical red dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
 The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
 The posterior PDFs, shown in the bottom row in blue, are more sharply peaked showing an increase in precision.
-The posterior PDFs also have modes that are more similar to the previously determined distance to the cluster (vertical lines) for far more stars.
+The posterior PDFs also have modes that are more similar to the previously determined parallax of the cluster (vertical lines) for far more stars.
 This shows that our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
@@
 \caption{Proof of concept: parallax inference to the open cluster M67.
-The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $1~{\rm deg}^2$ window centered on the position of M67.
+The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $0.75~{\rm deg}^2$ window centered on the position of M67.
 Our prior, visualized as a mixture of Gaussians, is shown in green.
-The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" distance to M67, determined by previous methods, bracketed by the red lines.
+The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the red dotted lines.
 Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
@@
 and using it as a prior PDF for Bayesian inference.
-The posterior PDFs for distance that we obtain are, in general, much
+The posterior PDFs for parallax that we obtain are, in general, much
 narrower than the likelihood functions delivered by the
@@
 least some aspects of \emph{accuracy} have been improved as well:
-The posterior distance estimates to stars in (at least one) stellar
-cluster are much more clustered than likelihood-based distance
+The posterior parallax estimates to stars in (at least one) stellar
+cluster are much more clustered than likelihood-based parallax
 estimates, and they cluster around a sensible value.
 Accuracy is demonstrated by this test because each star in the cluster is
-subject to a different, unique distance prior (see \figurename~\ref{fig:posterior})
+subject to a different, unique parallax prior (see \figurename~\ref{fig:posterior})
 and yet the posterior PDFs are all pulled to the same parallax value.
@@
 \caption{Comparison with a common spatial prior: A visualization of the posterior PDF for a subsection of \tgas\ stars, shown as the expectation value and $1\sigma$ uncertainty.
-The left panel shows the posterior PDF using the exponentially declining spatial density, a common space-density prior used when inferring distances.
+The left panel shows the posterior PDF using the exponentially declining spatial density, a common space-density prior used when inferring parallax.
 The right panel shows the posterior PDF using our \cmd\ prior.
@@
 
-There are other options for prior information to include when inferring distances to stars.
+There are other options for prior information to include when inferring the parallax of a star.
 One option is the exponentially declining spatial density (EDSD) of stars \citep{astraatmadja16a}.
@@
   \tgas--\tmass--\psone\ intersection.  It would be biasing to use this same
-  prior to infer distances for the full billion-star catalog released
+  prior to infer parallaxes for the full billion-star catalog released
   in \gaia\ DR1 alongside \tgas, or for stars fainter or brighter or
@@
   dust. We are taking a point estimate of the probabilistic dust model
-  at a quantile of our distance posterior, correcting our photometry
+  at a quantile of our parallax posterior, correcting our photometry
   using this point estimate, and then re-deriving the prior. A full,

### 2018-03-04 plots.py

@@ -10,6 +10,7 @@ from xdgmm import XDGMM
 import drawEllipse
 import comparePrior
 import scipy.interpolate
+import astropy.visualization as av
 
 def figsize_and_margins(plotsize,subplots=(1,1),**absolute_margins):
     '''Determine figure size and margins from plot size and absolute margins
@@ -52,7 +53,7 @@ def figsize_and_margins(plotsize,subplots=(1,1),**absolute_margins):
     #height * (top - bottom)
     #    = nrows * plot_height + (nrows - 1) * hspace * plot_height
     #solve for width and height, using absolute margins as necessary:
-    print(nc, rmarg['wspace'], pw, amarg.get('lef', 0), amarg.get('right', 0), rmarg.get('right', 1), rmarg.get('left', 0))
+    #print(nc, rmarg['wspace'], pw, amarg.get('lef', 0), amarg.get('right', 0), rmarg.get('right', 1), rmarg.get('left', 0))
     width  = float((nc + (nc - 1) * rmarg['wspace']) * pw        \
                    + amarg.get('left',0) + amarg.get('right',0)) \
              / (rmarg.get('right',1) - rmarg.get('left',0))
@@ -85,10 +86,14 @@ def gaussian(mean, sigma, array, amplitude=1.0):
     return amplitude/np.sqrt(2.*np.pi*sigma**2.)*np.exp(-(array - mean)**2./(2.*sigma**2.))
 
 
-def plotPrior(xdgmm, ax, c='black', lw=1):
+def plotPrior(xdgmm, ax, c='black', lw=1, stretch=False):
     for gg in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
-        ax.plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=lw, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
+        if stretch:
+            Stretch = av.PowerStretch(1./5)
+            alpha=np.power(xdgmm.weights[gg]/np.max(xdgmm.weights), 1./10)
+        else: alpha=xdgmm.weights[gg]/np.max(xdgmm.weights)
+        ax.plot(points[0,:],testXD.absMagKinda2absMag(points[1,:]), c, lw=lw, alpha=alpha)
 
[…]

### 2018-03-05 paper/paper.tex
took out unmade figure

@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{posterior_odd.pdf}
-\caption{Five example posterior PDFs over parallax using our method for more odd objects.
-The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
+%\begin{figure}
+%\centering
+%  \includegraphics[width=\textwidth]{posterior_odd.pdf}
+%\caption{Five example posterior PDFs over parallax using our method for more odd objects.
+%The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
 In blue, we show the posterior expectation value and $1 \sigma$ uncertainties.
-We chose example points that lie outside the well populated portions of the \cmd to illustrate how our method works for rare \tgas objects. 
-The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
-
-Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
-Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
-Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
-\label{fig:posterior}
-\end{figure}
+%We chose example points that lie outside the well populated portions of the \cmd to illustrate how our method works for rare \tgas objects. 
+%The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
+
+%Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
+%Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
+%Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
+%Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
+%\label{fig:posterior}
+%\end{figure}
 

### 2018-03-05 paper/paper.tex
copyright
hogg stuff

@@
 % This project is part of the photoParallax project.
-% Copyright 2017 the authors.
+% Copyright 2017, 2018 the authors.
 
@@
 Only a small fraction (but large number) of these
-measurements will be precise and purely astrometric.
-\gaia\ will use astrometric parallaxes to determine the distances of the more
-precise stars, and then calibrate spectrophotometric models (\citealt{apsis}).
+measurements will be precise and purely astrometric
+\added{The long-term complete distance information coming from \gaia\ will
+involve astrometric, photometric, and spectral information, and the balance
+of these contributions will be a strong function of stellar brightness (\citealt{apsis}).}
 These spectrophotometric models, along with \gaia's on-board
@@
 and handles in a principled way, heteroskedastic data.
-Its disadvantage, for the present purposes, is that it requires
+\added{It also handles incomplete data, in the sense of there being missing
+data values (these are like data values with arbitrarily large uncertainties);
+this aspect of XD is not so relevant to this project,
+where we have an essentially complete catalog.
+(There is another sense for the word ``incomplete'' in this area, which has to do with
+the completeness or selection function of the survey as a whole. We address this
+kind of completeness in our assumption called ``selection'' above and in the Discussion.)}
+XD's disadvantage, for the present purposes, is that it requires
 or implies a strictly Gaussian noise model.
@@
 Way.
-The strongest new assumption made in this work is that for every star
+The \added{fundamental and critical} assumption made in this work is that for every star
 in \tgas, there are other, photometrically (and bolometrically)

### 2018-03-05 plots.py
fixing color scheme

@@ -223,8 +223,9 @@ def main():
     np.random.seed(0)
 
     trueColor='darkred'
-    priorColor='darkgreen'
-    posteriorColor='royalblue'
+    priorColor= '#9ecae1' #'royalblue'
+    cmap_prior = 'Blues'
+    posteriorColor= '#3182bd' #'darkblue'
     dataColor='black'
     posteriorMapColor = 'Blues'
 
@@ -489,7 +490,7 @@ def main():
 
 
             samplex, sampley = sampleXDGMM(xdgmm, len(tgas)*10)
-            ax[0].hist2d(samplex, sampley, bins=100, norm=mpl.colors.LogNorm(), cmap=mpl.colors.Greens())
+            ax[0].hist2d(samplex, sampley, bins=500, norm=mpl.colors.LogNorm(), cmap=plt.get_cmap(cmap_prior))
             #plotPrior(xdgmm, ax[0], c=priorColor, lw=1, stretch=True)
             ax[0].set_ylim(15, -10)
             #ax[0].set_xlim(xlim_cmd)
@@ -530,8 +531,8 @@ def main():
                 if np.isnan(yminus): yminus = 10.
 
 
-                l1, = ax[i+1].plot(xparallaxMAS, likeParallax*np.max(posteriorParallax)/np.max(likeParallax), lw=1, color=dataColor, zorder=100)
-                l2, = ax[i+1].plot(xparallaxMAS, priorParallax*np.max(posteriorParallax)/np.max(priorParallax), lw=0.5, color=priorColor)
+                l1, = ax[i+1].plot(xparallaxMAS, likeParallax*np.max(posteriorParallax)/np.max(likeParallax), lw=2, color=dataColor, zorder=100)
+                l2, = ax[i+1].plot(xparallaxMAS, priorParallax*np.max(posteriorParallax)/np.max(priorParallax), lw=2, color=priorColor, linestyle='--')
                 l3, = ax[i+1].plot(xparallaxMAS, posteriorParallax, lw=2, color=posteriorColor)
                 ax[0].scatter(meanPosteriorColor, absMagPost, c=posteriorColor, s=20)
                 ax[0].errorbar(meanPosteriorColor, absMagPost, xerr=[[np.sqrt(varPosteriorColor)], [np.sqrt(varPosteriorColor)]], yerr=[[yplus], [yminus]], fmt="none", zorder=0, lw=2.0, mew=0, alpha=1.0, color=posteriorColor, ecolor=posteriorColor)

### 2018-04-03 paper/paper.tex
updated paper with asteroseismic comparison

@@
 \newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\apass}{\project{\acronym{APASS}}}
 \newcommand{\tgas}{\project{\acronym{TGAS}}}
@@
 
-%\begin{figure}
-%\centering
-%  \includegraphics[width=\textwidth]{posterior_odd.pdf}
-%\caption{Five example posterior PDFs over parallax using our method for more odd objects.
-%The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{posterior_odd.pdf}
+\caption{\added{Five example posterior PDFs over parallax using our method for more rare objects.
+The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented as a 2D histogram.
 In blue, we show the posterior expectation value and $1 \sigma$ uncertainties.
-%We chose example points that lie outside the well populated portions of the \cmd to illustrate how our method works for rare \tgas objects. 
-%The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
-
-%Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-%Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
-%Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
-%Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
-%\label{fig:posterior}
-%\end{figure}
+We chose example points that lie outside the well populated portions of the \cmd\ to illustrate how our method works for rare \tgas\ objects. 
+The remaining panels show the posterior PDF in blue, the prior PDF as shaded in green, and likelihood of the data in black. 
+XXX talk about specific examples}}
+\label{fig:posterior_odd}
+\end{figure}
 
@@
 \label{fig:m67}
+\end{figure}
+
+\subsection{Distances from Asteroseismology}
+\added{As a second test of accuracy of these photometric parallax inferences, we compared to the inferred distances to the stars from asteroseismology.
+Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars.
+There are about 2200 stars that overlap our data set here and the asteroseimic data from Huber at all.
+We compared the estimates in the parallax space (since the distance estimates are more precise, this is a safer/better comparison space) and show that our parallax estimates are unbiased relative to asteroseismic estimates, and are also more similar to the asteroseimic estimates than tgas parallax likelihoods.
+}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{photoVSseismic.pdf}
+\caption{Proof of concept: parallax estimate comparisons.}
+
+\label{fig:seismic}
 \end{figure}
@@
 we generate from our parallax posterior PDFs to be only weakly biased.
+
+Our model has similarities in spirit to previous work we have done on modeling the \gaia\ color-magnitude diagram (\citealt{leistedtHogg2017}; LH17).
+In LH17, the motivation was---like ours---to de-noise the distance information coming from \gaia\ using a data-driven model of the absolute photometric properties of stars.
+The results of LH17 are similar to the results here in terms of the reduction in variance in the distance (or parallax) estimates.
+The principal differences between LH17 and this work are that LH17 uses \apass\ photometry rather than \tmass\ photometry, and LH17 does a full sampling and marginalization of the color-magnitude-diagram model, not an optimized point estimate as we obtain and use here.
+There are several comments to make about the comparison between this work and LH17: For one, the \tmass\ photometry is very well calibrated and consistent across the sky, which is part of the reason that our results show more clearly the narrowness of the lower main sequence and the red clump than the LH17 results, and probably part of the reason that we resolve better the binary sequence than LH17.
+For two, the LH17 model is more principled in that it involves a full marginalization over the color-magnitude-diagram parameters.
+That technical feat is accomplished in part by making use of a coarser and less flexible model for the color-magnitude-diagram morphology.
+That is, the probabilistic correctness comes at the expense of model flexibility and therefore resolution of fine color-magnitude details.
+Overall, results are similar, however, and these two projects jointly point at the great value of hierarchical modeling for de-noising \gaia\ data, and indeed any data from large astronomy projects; they also show that there are multiple different but reasonable options for setting trade-offs between probabilistic correctness and model flexibility.
 

### 2018-04-17 paper/paper.tex
updated comparison to asteroseismology

@@
 The upper-left panel visualizes the simplified toy distribution.
-The red lines are the $1\sigma$ contours of the distribution, and the red points are 1024 samples from the distribution.
+The orange lines are the $1\sigma$ contours of the distribution, and the orange points are 1024 samples from the distribution.
 These same $1\sigma$ contours are drawn in the bottom left panel.
@@
 The black points are 5 randomly chosen data to highlight the method.
-In the lower-left panel, the green lines represent the noise deconvolved, best estimate for the underlying distribution.
+In the lower-left panel, the teal lines represent the noise deconvolved, best estimate for the underlying distribution.
 The blue points represent the expectation value of the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
-The red points represent the actual true values.
+The orange points represent the actual true values.
 The lower-right panel shows the expectation value, and $1\sigma$ uncertainties, of the posterior belief of the true value for all the points.
@@
 Here the thickness of the toy distribution mimics the true thickness, or intrinsic scatter, of the \cmd\ due to age, mass, and metallicity.
-The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel in red.
+The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel in orange.
 Similarly, noisy samples from this distribution are shown in the upper-right panel in grey and black.
@@
 The MMLE gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the green lines.
+In \figurename~\ref{fig:toy}, lower-left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the orange lines, and the best fit MMLE functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the teal lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
@@
 In \figurename~\ref{fig:toy} lower-left panel, the blue points represent the posterior beliefs (visualized as the expectation value and $1\sigma$ uncertainties) of the true values $y_{\true,n}$ for the 5 highlighted (black) points from the upper-right panel.
-The red points are the associated true $y$ values.
+The orange points are the associated true $y$ values.
 In the lower-right panel of \figurename~\ref{fig:toy}, the blue points represent the posterior beliefs, again visualized as the expectation value and $1\sigma$ uncertainty, for all 1024 $y_n$.
@@
 We can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
-For the remainder of the paper, we will also maintain the color scheme presented here (where possible): truth will be red, data will be black, prior information will be green, and posterior information will be blue.
+For the remainder of the paper, we will also maintain the color scheme presented here (where possible): truth will be orange, data will be black, prior information will be teal, and posterior information will be blue.
 
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise.
+\added{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior.
+This also requires us to ignore the star to star covariances in the \gaia\ data.}
 \item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
@@
 
-Figure~\ref{fig:posterior} shows the various probability distributions of our inference for a few randomly-chosen objects: the likelihood (black), prior (green), and posterior PDF (blue) on the true parallaxes.
+Figure~\ref{fig:posterior} shows the various probability distributions of our inference for a few randomly-chosen objects: the likelihood (black), prior (teal), and posterior PDF (blue) on the true parallaxes.
 For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ likelihood.
@@
 \caption{Five example posterior PDFs over parallax using our method.
-The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
+The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in teal.
 We chose example points that evenly spanned the color space, but within each color bin, we chose a point at random.
-The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
+The remaining panels show the posterior PDF in blue, the prior PDF in teal, and likelihood of the data in black.
 Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1, 3, and 4 are not very effected by the prior (green), where good data is good data.
+Examples 1, 3, and 4 are not very effected by the prior (teal), where good data is good data.
 Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
@@
 We chose example points that lie outside the well populated portions of the \cmd\ to illustrate how our method works for rare \tgas\ objects. 
-The remaining panels show the posterior PDF in blue, the prior PDF as shaded in green, and likelihood of the data in black. 
+The remaining panels show the posterior PDF in blue, the prior PDF as shaded in teal, and likelihood of the data in black. 
 XXX talk about specific examples}}
@@
 \figurename~\ref{fig:m67} shows the comparison of the likelihoods and posteriors.
-The vertical red dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
+The vertical orange dotted lines bracket the previously measured parallaxes of, or distances to, the cluster.
 The likelihoods of the observed parallaxes, shown in the top row in black, are broad and the cluster is not at all obvious.
@@
 The upper left panel visualizes the \gaia\ + \tmass\ likelihoods of the data, shown in black and projected onto the \cmd, for all the stars within our $0.75~{\rm deg}^2$ window centered on the position of M67.
-Our prior, visualized as a mixture of Gaussians, is shown in green.
-The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the red dotted lines.
-Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in green, as well as the posterior PDFs in blue.
+Our prior, visualized as a mixture of Gaussians, is shown in teal.
+The upper right panel shows the \tgas\ parallax likelihoods in black, with the ``true" parallax to M67, determined by previous methods, bracketed by the orange dotted lines.
+Similar to the upper left panel, the lower left panel shows the likelihoods of the data in black and our prior in teal, as well as the posterior PDFs in blue.
 The posterior PDFs are a mixture of Gaussians as well, each a product of the likelihood with a Gaussian in our prior.
@@
 \subsection{Distances from Asteroseismology}
-\added{As a second test of accuracy of these photometric parallax inferences, we compared to the inferred distances to the stars from asteroseismology.
-Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars.
-There are about 2200 stars that overlap our data set here and the asteroseimic data from Huber at all.
-We compared the estimates in the parallax space (since the distance estimates are more precise, this is a safer/better comparison space) and show that our parallax estimates are unbiased relative to asteroseismic estimates, and are also more similar to the asteroseimic estimates than tgas parallax likelihoods.
-}
+\added{As a second comparison of these photometric parallax inferences, we compare to the inferred distances to the stars from asteroseismology.
+Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars than the photometric parallax method.
+There are about $2200$ stars that overlap our data set here and the asteroseimic data from \citealt{huber17}.
+We compared the estimates in the parallax space (since the asteroseismic distance estimates are more precise, this is a safer comparison space since the inverted pdfs will be less skewed).
+We compare the cumulative difference in the parallax estimates in units of mas; the same units as the precision of \tgas.
+The comparison shows that our photo parallax estimates are more similar to the asteroseismic estimates than the \tgas\ likelihoods.
+The left panel and the middle panel show the one to one comparison of asteroseismic parallax estimates to the \tgas\ likelihoods and the photo parallax posterior pdf respectively for all $2200$ stars.
+The right panel shows the cumulative distribution function of the differences between the estimates for each star.
+The comparison to the photo parallax estimates, represented by the blue line, shows a distribution with smaller differences, and therefore more similar to the asteroseismic estimates.}
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{photoVSseismic.pdf}
-\caption{Proof of concept: parallax estimate comparisons.}
+\caption{\added{Proof of concept: parallax estimate comparisons.
+All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate or our photo parallax estimate.
+And the figure is meant to guide visual comparisons between \tgas\ and photo parallax, and therefore highlight differences in the estimates.
+The first panel shows the one to one comparison of the \tgas\ parallax estimates and the inverted distance estimates from \citealt{huber17}, and the associated $1\sigma$ uncertainties.
+The orange line shows the one to one line.
+The second panel is similar, but instead of the \tgas\ parallax estimates, it visualizes our photo parallax estimates.
+This shows the shrinkage of the precision of the parallax estimates, and that the expectation values are more in agreement with the asteroseismic estimates.
+The third panel panel visualizes the difference between the asteroseismic estimate and a \gaia\ estimate, \tgas\ in grey and photo parallax in blue, as a cumulative distribution function of the difference in units of mas.
+The photo parallax estimates are more similar to the asteroseismic estimates than \tgas\ likelihoods alone.}}
 
@@
 
-Our model has similarities in spirit to previous work we have done on modeling the \gaia\ color-magnitude diagram (\citealt{leistedtHogg2017}; LH17).
+\added{Our model has similarities in spirit to previous work we have done on modeling the \gaia\ color-magnitude diagram (\citealt{leistedtHogg2017}; LH17).
 In LH17, the motivation was---like ours---to de-noise the distance information coming from \gaia\ using a data-driven model of the absolute photometric properties of stars.
@@
 That is, the probabilistic correctness comes at the expense of model flexibility and therefore resolution of fine color-magnitude details.
-Overall, results are similar, however, and these two projects jointly point at the great value of hierarchical modeling for de-noising \gaia\ data, and indeed any data from large astronomy projects; they also show that there are multiple different but reasonable options for setting trade-offs between probabilistic correctness and model flexibility.
+Overall, results are similar, however, and these two projects jointly point at the great value of hierarchical modeling for de-noising \gaia\ data, and indeed any data from large astronomy projects; they also show that there are multiple different but reasonable options for setting trade-offs between probabilistic correctness and model flexibility.}
 

### 2018-04-18 paper/paper.tex
last edits to paper

@@
 \begin{abstract}\noindent % trust me
-{\bf \textcolor{red}{Stephen Feeeeeeeneeeeeey is the best.}} Converting a noisy parallax measurement into a posterior
+%{\bf \textcolor{red}{Stephen Feeeeeeeneeeeeey is the best.}} 
+Converting a noisy parallax measurement into a posterior
 belief over distance requires inference with a prior.
@@
 \added{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior.
-This also requires us to ignore the star to star covariances in the \gaia\ data.}
+We also ignore the star to star covariances because they are not available in the \tgas\ dataset.}
 \item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
@@
 Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
-Examples 1, 3, and 4 are not very effected by the prior (teal), where good data is good data.
-Examples 2 and 5 show a more precise posterior PDF compared with the likelihood.
-Example 5 looks very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
+Examples 1, 2, and 3 are not very effected by the prior (teal), where good data is good data.
+Examples 4 and 5 show a more precise posterior PDF compared with the likelihood.
+For these examples, the posterior PDF is very similar to the prior PDF and exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
 \label{fig:posterior}
@@
 The remaining panels show the posterior PDF in blue, the prior PDF as shaded in teal, and likelihood of the data in black. 
-XXX talk about specific examples}}
+Examples 2 and 4 are not very affected by the prior (teal), where good data is good data.
+Examples 1, 3, and 5 show a more precise posterior PDF compared with the likelihood.
+For these examples, the posterior PDF is very similar to the prior PDF and might be biasing the posterior PDF estimate due to a lack of prior support at those values.
+These might be places where our stationarity assumption breaks down.}}
 \label{fig:posterior_odd}
@@
 There are about $2200$ stars that overlap our data set here and the asteroseimic data from \citealt{huber17}.
-We compared the estimates in the parallax space (since the asteroseismic distance estimates are more precise, this is a safer comparison space since the inverted pdfs will be less skewed).
+In \figurename~\ref{fig:seismic} we compare the estimates in the parallax space.
+We chose the parallax space, rather than the distance space, since the asteroseismic distance estimates are more precise, making it a less skewed inverted pdf.
 We compare the cumulative difference in the parallax estimates in units of mas; the same units as the precision of \tgas.
 The comparison shows that our photo parallax estimates are more similar to the asteroseismic estimates than the \tgas\ likelihoods.
-The left panel and the middle panel show the one to one comparison of asteroseismic parallax estimates to the \tgas\ likelihoods and the photo parallax posterior pdf respectively for all $2200$ stars.
-The right panel shows the cumulative distribution function of the differences between the estimates for each star.
-The comparison to the photo parallax estimates, represented by the blue line, shows a distribution with smaller differences, and therefore more similar to the asteroseismic estimates.}
+The left panel and the middle panel show the one to one comparison of asteroseismic parallax estimates to the \tgas\ likelihoods (in black) and the photo parallax posterior pdf (in blue) respectively for all $2200$ stars.
+The right panel shows the cumulative distribution function of the differences between the asteroseismic estimates and the \tgas\ likelihoods (in black) or the photoparallax estimates (in blue) for each star.
+The comparison to the photo parallax estimates, represented by the blue line, shows a distribution with smaller differences, and therefore represents estimates that are more similar to the asteroseismic estimates.}
 \begin{figure}
@@
 \caption{\added{Proof of concept: parallax estimate comparisons.
-All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate or our photo parallax estimate.
-And the figure is meant to guide visual comparisons between \tgas\ and photo parallax, and therefore highlight differences in the estimates.
-The first panel shows the one to one comparison of the \tgas\ parallax estimates and the inverted distance estimates from \citealt{huber17}, and the associated $1\sigma$ uncertainties.
+All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate (in black) or our photo parallax estimate (in blue).
+And the figure is meant to guide visual comparisons between \tgas\ and photo parallax, and therefore highlight differences in these estimates.
+The first panel shows the one to one comparison of the \tgas\ parallax estimates and the inverted distance estimates from \citealt{huber17} using asteroseismology, and the associated $1\sigma$ uncertainties.
 The orange line shows the one to one line.
@@
 The third panel panel visualizes the difference between the asteroseismic estimate and a \gaia\ estimate, \tgas\ in grey and photo parallax in blue, as a cumulative distribution function of the difference in units of mas.
-The photo parallax estimates are more similar to the asteroseismic estimates than \tgas\ likelihoods alone.}}
+The CDF of the photo parallax estimates is shifted to smaller differences, so the photoparallax estimates are therefore more similar to the asteroseismic estimates than \tgas\ likelihoods alone.}}
 

### 2018-04-18 plots.py

@@ -11,6 +11,7 @@ import drawEllipse
 import comparePrior
 import scipy.interpolate
 import astropy.visualization as av
+from matplotlib.colors import LinearSegmentedColormap
 
 def figsize_and_margins(plotsize,subplots=(1,1),**absolute_margins):
     '''Determine figure size and margins from plot size and absolute margins
@@ -183,13 +184,13 @@ def likePriorPost(color, absMagKinda, color_err, absMagKinda_err, apparentMagnit
 def main():
     #    for label, style in zip(['paper', 'talk'],['seaborn-paper', 'seaborn-talk']):
     pdf = True
-    plot_data = False
+    plot_data = True
     plot_dust = False
     plot_prior = False
-    plot_m67 = False
+    plot_m67 = True
     plot_compare = False
-    plot_expectation = False
-    plot_odd_examples = True
+    plot_expectation = True
+    plot_odd_examples = False
     plot_examples = False
     plot_delta = False
     plot_deltacdf = False
@@ -222,12 +223,24 @@ def main():
     nsubsamples = 1024
     np.random.seed(0)
 
-    trueColor='darkred'
-    priorColor= '#6baed6' #'#9ebcda' #'#9ecae1' #'royalblue'
+    trueColor='#FF8D28'
+    priorColor= '#7bccc4'#'#6baed6' #'#9ebcda' #'#9ecae1' #'royalblue'
+    priorColor = '#6FB8B0'
     cmap_prior = 'Blues'
-    posteriorColor= '#984ea3' #'#7a0177' #'#8856a7' #'#810f7c' #'#08519c' #'darkblue'
+    posteriorColor= '#0977C4' #'#0868ac' #'#984ea3' #'#7a0177' #'#8856a7' #'#810f7c' #'#08519c' #'darkblue'
     dataColor='black'
     posteriorMapColor = 'Blues'
+    annotationColor = '#FF2412'
+    color1 = np.array((240,249,232))/255.
+    color1 = np.array((255,255,255))/255.
+    #color2 = np.array((123,204,196))/255.
+    color2 = np.array((112, 186, 179))/255.
+    colors = [color1, color2]
+    cm = LinearSegmentedColormap.from_list('my_color', colors, N=100)
+    cmap_prior = LinearSegmentedColormap.from_list('my_color', colors, N=100)
+    color2 = np.array((6, 82, 135))/255.
[…]

### 2018-04-18 testXD.py

@@ -402,12 +402,12 @@ def distanceTest(tgas, xdgmm, nPosteriorPoints, data1, data2, err1, err2, xlim, 
     np.save('summedPosteriorM67', summedPosterior)
 
 
-    axDist[0].set_xlabel(xlabel, fontsize=18)
-    axDist[0].set_ylabel(ylabel, fontsize=18)
+    axDist[0].set_xlabel(xlabel)
+    axDist[0].set_ylabel(ylabel)
     axDist[0].set_xlim(xlim)
     axDist[0].set_ylim(ylim)
-    axDist[2].set_xlabel(xlabel, fontsize=18)
-    axDist[2].set_ylabel(ylabel, fontsize=18)
+    axDist[2].set_xlabel(xlabel)
+    axDist[2].set_ylabel(ylabel)
     axDist[2].set_xlim(xlim)
     axDist[2].set_ylim(ylim)
 
@@ -420,10 +420,10 @@ def distanceTest(tgas, xdgmm, nPosteriorPoints, data1, data2, err1, err2, xlim, 
 
     axDist[1].set_xlabel('Parallax [mas]')
     axDist[3].set_xlabel('Parallax [mas]')
-    axDist[1].set_ylabel(ylabel_likelihood_parallax, fontsize=18)
-    axDist[3].set_ylabel(ylabel_posterior_parallax, fontsize=18)
-    axDist[1].set_xlim(-1, 3)
-    axDist[3].set_xlim(-1, 3)
+    axDist[1].set_ylabel(ylabel_likelihood_parallax)
+    axDist[3].set_ylabel(ylabel_posterior_parallax)
+    axDist[1].set_xlim(0, 2)
+    axDist[3].set_xlim(0, 2)
 
     axDist[0].legend(loc='best')
     axDist[2].legend(loc='best')

### 2018-04-18 toy/toy.py

@@ -48,37 +48,38 @@ def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mt
     alpha_chosen = 1.0
     xlim = (-1, 1)
     ylim = (-5, 5)
-
+    markersize = 8
+    size = 25
     for ax in axes[1:-1]:
-        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color=dataColor, alpha=alpha_all ,mew=0)
-        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color=dataColor, zorder=37, alpha=alpha_chosen, mew=0)
+        ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color=dataColor, alpha=alpha_all ,mew=0, zorder=0)
+        ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color=dataColor, zorder=7, alpha=alpha_chosen, mew=0, markersize=markersize)
 
     xp = np.array(xlim)
     axes[0].plot(xp, mtrue*xp + btrue + ttrue, color=trueColor, linewidth=2, alpha=0.75, label=r'$y=m_{true}\,x+b_{true}\pm t$')
     axes[0].plot(xp, mtrue*xp + btrue - ttrue, color=trueColor, linewidth=2, alpha=0.75)
     axes[0].scatter(xns, yntrues, c=trueColor, lw=0, alpha=0.5, label=r'$y_{true,n}$')
-    axes[0].legend(loc='best', fontsize=15)
+    axes[0].legend(loc='best')#, fontsize=15)
 
-    axes[2].plot(xp, m * xp + b + t, color=priorColor)
-    axes[2].plot(xp, m * xp + b - t, color=priorColor)
-    axes[2].scatter(xns[0:nexamples], yntrues[0:nexamples], c=trueColor, lw=2, zorder=36, alpha=alpha_chosen, facecolors='None')
-    axes[2].plot(xp, mtrue*xp + btrue + ttrue, color=trueColor, zorder=35)
-    axes[2].plot(xp, mtrue*xp + btrue - ttrue, color=trueColor, zorder=34)
+    axes[2].plot(xp, m * xp + b + t, color=priorColor, lw=3, zorder=4)
+    axes[2].plot(xp, m * xp + b - t, color=priorColor, lw=3, zorder=5)
[…]

### 2018-07-30 paper/paper.tex
final referee report ;

@@
 Only a small fraction (but large number) of these
-measurements will be precise and purely astrometric
-\added{The long-term complete distance information coming from \gaia\ will
+measurements will be precise and purely astrometric.
+\edit1{The long-term complete distance information coming from \gaia\ will
 involve astrometric, photometric, and spectral information, and the balance
@@
 and handles in a principled way, heteroskedastic data.
-\added{It also handles incomplete data, in the sense of there being missing
+\edit1{It also handles incomplete data, in the sense of there being missing
 data values (these are like data values with arbitrarily large uncertainties);
@@
 Way.
-The \added{fundamental and critical} assumption made in this work is that for every star
+The \edit1{fundamental and critical} assumption made in this work is that for every star
 in \tgas, there are other, photometrically (and bolometrically)
@@
 In other words, the features of the inferred \cmd\ will typically be broader and more conservatively estimated than if we over-estimated the noise.
-\added{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior.
-We also ignore the star to star covariances because they are not available in the \tgas\ dataset.}
+\edit1{We need to make these assumptions, that the noise model is Gaussian for both the parallaxes and the colors (including dust corrections), to be able to use the tool \xd, which requires Gaussian uncertainties, to deconvolve the data with it's noise for generating the prior.
+We also ignore the star to star covariances.
+\deleted{because they are not available in the \tgas\ dataset}.}
 \item[dust] We treat the \psone-based three-dimensional dust maps (\citealt{green15})
@@
 The intrinsic temperatures and luminosities of stars tightly correlate, leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
-The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \added{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
+The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains \edit1{parallax information $\varpi$ via the fundamental relation $m=M+10-5\log(\varpi[\mathrm{mas}])$.}
 We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
@@
 The transformation of the (dust corrected) absolute magnitude is required to keep the uncertainties Gaussian, and Gaussian uncertainties are required to use \xd, see below.
-\added{This transformation also allows our method to handle negative parallaxes from the \tgas\ dataset naturally.
+\edit1{This transformation also allows our method to handle negative parallaxes from the \tgas\ dataset naturally.
 A negative parallax simply refers to a distant object that is affected by large errors, and therefore intrinsically bright, and our transformation places these objects in the bright portion of the \cmd.}
@@
 The match was done using a nearest-neighbor algorithm with a search radius of $4$ arcsec\footnote{\url{http://portal.nersc.gov/project/cosmo/temp/dstn/gaia/tgas-matched-2mass.fits.gz}}.
-\added{There is also a \tgas\-\tmass\ cross-match as part of the official \gaia\ product, but this project began before that catalogue was available.}
+\edit1{There is also a \tgas\-\tmass\ cross-match as part of the official \gaia\ product \cite{Marrese17}, but this project began before that catalogue was available.}
 We also required that the stars lie within the
@@
   \includegraphics[width=\textwidth]{posterior_odd.pdf}
-\caption{\added{Five example posterior PDFs over parallax using our method for more rare objects.
+\caption{\edit1{Five example posterior PDFs over parallax using our method for more rare objects.
 The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented as a 2D histogram.
@@
 Keep in mind, the (low) number of M67 members seen here is constrained by the (unknown) selection effects of the \tgas\ dataset.
+\edit2{Due to the small number of M67 members in the \tgas\ dataset, the reliability of this test is slightly suspect, but still interesting.}
 With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
@@
 \subsection{Distances from Asteroseismology}
-\added{As a second comparison of these photometric parallax inferences, we compare to the inferred distances to the stars from asteroseismology.
+\edit1{As a second comparison of these photometric parallax inferences, we compare to the inferred distances to the stars from asteroseismology.
 Asteroseimology can make much more precise distance, and therefore parallax, estimates to stars than the photometric parallax method.
@@
   \includegraphics[width=\textwidth]{photoVSseismic.pdf}
-\caption{\added{Proof of concept: parallax estimate comparisons.
+\caption{\edit1{Proof of concept: parallax estimate comparisons.
 All quantified comparisons in this figure are made between a parallax estimate from asteroseismology and a parallax estimate from \gaia, either the \tgas\ estimate (in black) or our photo parallax estimate (in blue).
@@
 
-\section{Discussion}
+\section{Discussion and Conclusions}
 
@@
 
-\added{Our model has similarities in spirit to previous work we have done on modeling the \gaia\ color-magnitude diagram (\citealt{leistedtHogg2017}; LH17).
+\edit1{Our model has similarities in spirit to previous work we have done on modeling the \gaia\ color-magnitude diagram (\citealt{leistedtHogg2017}; LH17).
 In LH17, the motivation was---like ours---to de-noise the distance information coming from \gaia\ using a data-driven model of the absolute photometric properties of stars.
@@
 
-\section{Conclusion}
-We forgo an explicit conclusion; we attempt to summarize the conclusions in the Abstract above.
+\deleted{ Conclusion -- 
+We forgo an explicit conclusion; we attempt to summarize the conclusions in the Abstract above.}
 
